### 2022-09-15 paper.tex
First commit

\documentclass[times,sort&compress]{elsarticle}
\usepackage{ycviu}
\usepackage{framed}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{url}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}

\graphicspath{{images/}}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\NewDocumentCommand{\codeword}{v}{\texttt{\textcolor{orange}{#1}}}

\definecolor{newcolor}{rgb}{.8,.349,.1}
\definecolor{codegreen}{rgb}{0.7,0.7,0.7}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{1,1,1}
\lstdefinestyle{mystyle}{ backgroundcolor=\color{backcolour},   
commentstyle=\color{codegreen}, keywordstyle=\color{magenta},
numberstyle=\tiny\color{codegray}, stringstyle=\color{codepurple},
basicstyle=\ttfamily\footnotesize, breakatwhitespace=false,         
breaklines=true,                 
captionpos=b,                    
keepspaces=true,                 
numbersep=5pt,                  
showspaces=false,                
showstringspaces=false, showtabs=false,                  
tabsize=2 }
\lstset{style=mystyle}

\journal{Computer Vision and Image Understanding}

\begin{document}

\thispagestyle{empty}

\begin{table*}[!th]

\begin{minipage}{.9\textwidth}
\baselineskip12pt
\ifpreprint
  \vspace*{1pc}
\else
  \vspace*{-6pc}
\fi

\noindent {\LARGE\itshape Computer Vision and Image Understanding}
\vskip6pt

\noindent {\Large\bfseries Authorship Confirmation}

\vskip1pc

{\bf Please save a copy of this file, complete and upload as the ``Confirmation of
Authorship'' file.}

\vskip1pc

As corresponding author I, \underline{\hphantom{\hspace*{7cm}}}, hereby confirm on
behalf of all authors that:

\vskip1pc

\begin{enumerate}
\itemsep=3pt
\item This manuscript, or a large part of it, \underline {has not been published,  was
not, and is not being submitted to} any other journal.

\item If \underline {presented} at or \underline {submitted} to or \underline
{published }at a conference(s), the conference(s) is (are) identified and  substantial
\underline {justification for re-publication} is presented  below. A \underline {copy of
conference paper(s) }is(are) uploaded with the  manuscript.

\item If the manuscript appears as a preprint anywhere on the web, e.g. arXiv,  etc., it
is identified below. The \underline {preprint should include a  statement that the paper
is under consideration at Computer Vision and Image Understanding}.

\item All text and graphics, except for those marked with sources, are \underline
{original works} of the authors, and all necessary permissions for  publication were
secured prior to submission of the manuscript.

\item All authors each made a significant contribution to the research reported  and
have \underline {read} and \underline {approved} the submitted  manuscript. 
\end{enumerate}

Signature\underline{\hphantom{\hspace*{7cm}}} Date\underline{\hphantom{\hspace*{4cm}}}
\vskip1pc

\rule{\textwidth}{2pt}
\vskip1pc

{\bf List any pre-prints:} \vskip5pc

\rule{\textwidth}{2pt}
\vskip1pc

{\bf Relevant Conference publication(s) (submitted, accepted, or published):} \vskip5pc

{\bf Justification for re-publication:}

\end{minipage}
\end{table*}

\clearpage
\thispagestyle{empty}
\ifpreprint
  \vspace*{-1pc}
\fi

\begin{table*}[!th]
\ifpreprint\else\vspace*{-5pc}\fi

\section*{Graphical Abstract (Optional)}

\vskip1pc \fbox{
\begin{tabular}{p{.4\textwidth}p{.5\textwidth}} \bf Altering Backward Pass Gradients
improves Convergence  \\
Bishshoy Das, Milton Mondal\\ Brejesh Lall, Shiv Dutt Joshi, Sumantra Dutta Roy \\[1pc]
& In standard neural network training, the gradients in the backward pass are determined
by the forward pass. As a result, the two stages are coupled. This is how most neural
networks are trained currently. However, gradient modification in the backward pass has
seldom been studied in the literature. In this paper we explore decoupled training,
where we alter the gradients in the backward pass. We propose a simple yet powerful
method called PowerGrad Transform, that alters the gradients before the weight update in
the backward pass and significantly enhances the predictive performance of the neural
network. PowerGrad Transform trains the network to arrive at a better optima at
convergence. It is computationally extremely efficient, virtually adding no additional
cost to either memory or compute, but results in improved final accuracies on both the
training and test sets. PowerGrad Transform is easy to integrate into existing training
routines, requiring just a few lines of code. PowerGrad Transform makes it possible for
the network to better fit the training data. With decoupled training, PowerGrad
Transform improves baseline accuracies for ResNet-50 by 0.73\%, for SE-ResNet-50 by
0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network on the ImageNet
classification task.
\end{tabular}
}

\end{table*}

\clearpage
\thispagestyle{empty}

\ifpreprint
  \vspace*{-1pc}
\else
%  \vspace*{-6pc}
\fi

\begin{table*}[!t]
\ifpreprint\else\vspace*{-15pc}\fi

\section*{Research Highlights (Required)}

To create your highlights, please type the highlights against each \verb+\item+ command.

\vskip1pc

\fboxsep=6pt \fbox{
\begin{minipage}{.95\textwidth}
It should be short collection of bullet points that convey the core findings of the
article. It should  include 3 to 5 bullet points (maximum 85 characters, including
spaces, per bullet point.)  
\vskip1pc
\begin{itemize}

\item Altering gradients in the backward pass can significantly alter a CNN's
performance.

\item PowerGrad Transform is proposed that enhances the backward pass gradients.

\item Theoretical analysis of the impact of PowerGrad transformation is conducted.

\item Efficacy of PowerGrad Transform is demonstrated through experiments.

\end{itemize}
\vskip1pc
\end{minipage}
}

\end{table*}

\clearpage

\ifpreprint
  \setcounter{page}{1}
\else
  \setcounter{page}{1}
\fi

\begin{frontmatter}

\title{Altering Backward Pass Gradients improves Convergence}

\author[1]{Bishshoy Das\corref{cor1}}
\ead{bishshoy.das@ee.iitd.ac.in}

\author[1]{Milton Mondal\corref{cor2}}
\ead{milton.mondal@ee.iitd.ac.in}

\author[1]{\\ Brejesh Lall}
\ead{brejesh@ee.iitd.ac.in}

\author[1]{Shiv Dutt Joshi}
\ead{sdjoshi@ee.iitd.ac.in}

\author[1]{Sumantra Dutta Roy}
\ead{sumantra@ee.iitd.ac.in}

\cortext[cor1]{Corresponding author}
\cortext[cor2]{Equal contribution}

\address[1]{Electrical Engineering Department, Indian Institute of Technology Delhi, Hauz Khas, New Delhi - 110016, India}

\received{1 May 2013}
\finalform{10 May 2013}
\accepted{13 May 2013}
\availableonline{15 May 2013}
\communicated{S. Sarkar}

\begin{abstract}
In standard neural network training, the gradients in the backward pass are determined
by the forward pass. As a result, the two stages are coupled. This is how most neural
networks are trained currently. However, gradient modification in the backward pass has
seldom been studied in the literature. In this paper we explore decoupled training,
where we alter the gradients in the backward pass. We propose a simple yet powerful
method called PowerGrad Transform, that alters the gradients before the weight update in
the backward pass and significantly enhances the predictive performance of the neural
network. PowerGrad Transform trains the network to arrive at a better optima at
convergence. It is computationally extremely efficient, virtually adding no additional
cost to either memory or compute, but results in improved final accuracies on both the
training and test sets. PowerGrad Transform is easy to integrate into existing training
routines, requiring just a few lines of code. PowerGrad Transform makes it possible for
the network to better fit the training data. With decoupled training, PowerGrad
Transform improves baseline accuracies for ResNet-50 by 0.73\%, for SE-ResNet-50 by
0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network on the ImageNet
classification task.
\end{abstract}

\begin{keyword}
\KWD backpropagation \sep gradients \sep neural networks \sep softmax \sep clipping
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{sec:Intr}

Backpropagation is traditionally used to train deep neural networks
\cite{lillicrap2020backpropagation}. Gradients are computed using basic calculus
principles to adjust the weights during backpropagation \cite{lecun1988theoretical}.
Alternatives to traditional gradients has rarely been studied in the literature
hitherto. In normal training procedures, gradients are computed immediately in the
backward pass utilizing the values obtained in the forward pass. This makes the backward
pass coupled with the forward pass. However, decoupling the backward pass from the
forward pass by modifying the gradients to improve training efficiency and final
convergent accuracy has hardly been explored. In this paper we explore the landscape of
decoupling the forward pass from the backward pass by altering the gradients and
subsequently updating the network's parameters with the modified gradients. There are
several ways to achieve gradient modification in the backward pass. We discuss a few
techniques in Fig. \ref{fig:gradient_altering}. One could modify the gradients at either
\textbf{(I)} multiple points in the backward computation graph or \textbf{(II)} at a
point very early in the backward graph and allow this changed gradient to also affect
the rest of the network's gradients as the it backpropagates through the backward
computation graph.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{PGT_Types}
\caption{ Different ways of altering gradients in the backward pass. Blue blocks denote
a CNN's different layers (convolutional, linear, batchnorm, softmax, etc.). Orange
blocks indicate the backward graph with unmodified gradients. Green blocks represent
transformation functions, while red blocks indicate transformed gradients. (a)
\textbf{Type 0:} Backpropagation without any gradient modification. (b) \textbf{Type I:}
Altering gradients after they are all computed. (c) \textbf{Type II:} Altering gradients
at an earlier stage in the backward computation graph, subsequently leading to
alteration of all other gradients. }
\label{fig:gradient_altering}
\end{figure}

\textbf{Type 0:} No modification: In this method, we calculate the gradients using the
standard calculus rules and use the chain rule to calculate the gradients of the rest of
the network's parameters, also known as backpropagation (Fig.
\ref{fig:gradient_altering}a). The network is then updated with the gradient descent
equation:

\begin{equation}
\hspace{4cm}
W_i^{t+1} = W_i^t - \lambda\ \nabla_{W_i}(L)
\hspace{2cm}
i=D,D-1,\ldots,1
\end{equation}

\textbf{Type I:} Gradient modification at multiple points independently: In this method,
the gradients are first computed using standard procedure and then individually altered
(Fig. \ref{fig:gradient_altering}b). Gradient clipping \cite{pascanu2013difficulty} and
Adaptive gradient clipping \cite{brock2021high} are examples of such modifications. In
both of these methods, the gradients are first computed using standard rules and then
they are modified using some function. It can be described as:

\begin{equation}
\hspace{4cm}
W_i^{t+1} = W_i^t - \lambda\ f(\nabla_{W_i}(L))
\hspace{2cm}
i=D,D-1,\ldots,1
\end{equation}

where the gradients $\nabla_{W_i}(L)$ are transformed using the transformation function
`$f$' before the weight update.

\textbf{Type II:} Gradient modification at a point very early in the backward graph: In
this type of modification, the gradient is altered at a very early stage in the backward
computation graph and then all subsequent gradients are generated using the values
obtained with the modified gradients (Fig. \ref{fig:gradient_altering}c). Because of the
chain rule, network parameters whose gradients are connected to the point of alteration
the computation graph also gets subsequently altered. It can be described as:

\begin{align}
\hspace{4cm}
W_D^{t+1} &= W_D^t - \lambda\ f(\nabla_{W_D}(L)) & \\
W_i^{t+1} &= W_i^t - \lambda\ \nabla_{W_i}(L) &\hspace{-3cm} i=D-1,\ldots,1
\end{align}

where first the gradient $\nabla_{W_D}(L)$ is transformed using the transformation
function `$f$' and then this transformed gradient is propagated through the rest of the
backward graph and all other gradient vectors are computed as is. Because of the early
injection of the transformed gradients, all other gradient vectors that are connected to
the transformed gradient through the chain rule, gets subsequently altered.

Type I modification is computationally more expensive than Type II modification as it
requires altering the gradients of each and every parameter individually. Type II
modification recomputes gradients at each and every location through the natural flow of
backpropagation. We propose PowerGrad Transform (PGT) which is a type II gradient
modification method. With virtually no additional compute or memory requirement,
PowerGrad Transform enhances classification performance of a convolutional neural
network, by leading the network to arrive at a higher quality optima at convergence, at
which both training and test accuracies improve.

PowerGrad Transform modifies the gradients at the softmax layer, which is the earliest
part of most convolutional neural network's computation graph. By changing the gradient
at this stage, all other gradients for all other parameters in the network such as the
linear layer, batchnorm statistics and convolutional layer's filter parameters are
affected. We analyze and explore the behaviour of such alteration both mathematically
and through empirical experiments conducted on networks without batch-normalization.
Experimentally, we see that PGT improved training and test accuracies for both
normalized and non-normalized networks. PGT improves a network's learning capacity and
results in a better fit of the training data.

The following are the major contributions of this paper:

\begin{enumerate}

\item We introduce PowerGrad Transform, which decouples the backward and the forward
passes of neural network training and enables a considerably better fit to the dataset,
as assessed by both training and test accuracy measures. PGT is a gradient enhancement
method that alters the gradients in the backward pass before the update step leading to
significant boost in network's predictive performance.

\item We perform theoretical analysis of the properties of the PowerGrad transformation
(section \ref{sec:pgt_prop}) and explore its effect on weight parameters and gradients
(section \ref{sec:pgt_weight_gradients}), logits (section
\ref{sec:pgt_logit_derivative}), (section \ref{sec:Effe}), loss values (section
\ref{sec:Effe}) and class separation (section \ref{sec:pgt_class_separation}).

\item We study experimentally the degenerate behaviour of non-BN networks that is often
seen during the normal training procedure, particularly at larger batch sizes. With PGT,
we see improved gradient behaviour and a decreased likelihood of the weights attaining
degenerate states.

\item We provide complete results from a variety of models (non-BN and BN ResNets,
SE-ResNets) using the ImageNet dataset. It helps the network to improve its learning
capabilities by locating a more optimum convergence point. Additionally, we conduct an
ablation study and compare its impacts to those of regularization approaches such as
label smoothing.

\end{enumerate}

\section{Related Works}
\label{sec:Rela}

We examine techniques that are related to gradient modification and emphasize the
distinction between them and our proposed method.

Gradient Clipping (GC): Gradient Clipping \cite{pascanu2013difficulty}, often used in
natural language processing methods \cite{merity2017regularizing}, is a technique that
involves changing or clipping the gradients with respect to a predefined threshold value
during backward propagation through the network and updating the weights using the
clipped gradients \cite{zhang2019gradient, smith2020generalization}. By rescaling the
gradients, the weight updates are likewise rescaled, significantly reducing the risk of
an overflow or underflow \cite{pascanu2012understanding}. GC can be used for training
networks without batch-normalization.

The formulation is as follows: If $G$ is a gradient vector which is $\partial
L / \partial \theta$ (the gradient of the loss $L$ with respect to
the parameters $\theta$). In clip by value, the gradients are modified as:

\begin{equation}
G \rightarrow
\begin{cases}
\lambda \frac{G}{\norm{G}} & \text{if } \norm{G} > \lambda, \\
G & \text{otherwise}
\end{cases}
\end{equation}

At larger batch sizes, the clipping threshold in GC becomes highly sensitive and
requires extensive finetuning for various models, batch sizes, and learning rates. As we
demonstrate later in our studies, GC is not as effective in improving the performance of
non-normalized networks. AGC performs better than GC in non-normalized networks.
However, we show that PGT outperforms both in such networks.

Adaptive Gradient Clipping (AGC): Adaptive Gradient Clipping \cite{brock2021high} is
developed to further enhance backward pass gradients than what is performed by GC. It
takes into account the fact that the ratio of the gradient norm to the weight norm can
provide an indication of the expected change in a single step of optimization. Here the
normalized gradient is clipped by $\lambda$, which means that large weights can have
large gradients and still the normalized gradient can be within $\lambda$. If it is not
within the threshold, then their gradient is normalized by the ratio of norm of the
gradient and norm of the weight in order to avoid gradient explosion. If $W^l$ is the
weight matrix of the $l^\textit{th}$ layer, and $G^l$ is its gradient in the backward
pass, then the following equation is used to modify the gradient of the $i^\textit{th}$
filter before its update. In most cases, AGC is used for training non-BN networks. Our
experiments demonstrate that PGT outperforms AGC, when used independently. When used
together, it can further improve a network's performance.

Label Smoothing: Label smoothing, introduced by Szegedy et al.
\cite{szegedy2016rethinking}, utilizes smoothing of the ground truth labels as a method
to impose regularization on the logits and the weights of the neural network.

The formulation is as follows. If $q_i$ is the value at the $i^{th}$ index of the one
hot encoded vector of the ground truth label, then the transformed distribution of the
labels $q_i'$ is: \begin{equation} q_i' = \begin{cases} 1-\epsilon & \text{if } i = y,
\\ \frac{\epsilon}{K-1} & otherwise \end{cases} \end{equation} where $\epsilon$ is a
hyperparameter with a value in $[0,1]$ and $y$ denotes the index of the correct class.

Müller et al. investigates how label smoothing work in \cite{DBLP:conf/nips/MullerKH19}.
Using visualizations of penultimate layer activations, they demonstrate that label
smoothing calibrates the networks predictions and aligns them with the accuracies of the
predictions. Parallel works in label smoothing include Pairwise Confusion
\cite{dubey2018pairwise}, combatting Label Noise \cite{reed2014training}, achieving
regularization through intentional label modification \cite{xie2016disturblabel}.

PowerGrad Transform works in a much different way. PGT does not smooth the ground truth
labels, rather it modifies the gradient by smoothing the predicted probability vector.
As we show in section \ref{sec:Math}, our proposed transformation has the opposite
effect of label smoothing. From our studies, we show that PGT in fact leads to larger
generalization gap between the training and test dataset. Coupled with PGT, label
smoothing can narrow the generalization gap.

Knowledge Distillation: The formulation of PowerGrad Transform is akin to temperature
based logit suppression of \cite{hinton2015distilling}. Knowledge distillation
\cite{DBLP:conf/nips/BaC14} is a process in which two networks are trained. First a
teacher network is trained on a given dataset and then the soft-labels (the predicted
probabilities) of the teacher network are used to train another network, called the
student network. In similar style of label smoothing, knowledge distillation aims to
generate smooth gradient behaviour by forcing the network not to overpredict on any
single image. As in case of label smoothing, the student network's weights are
automatically penalized if the network assigns a probability value higher than the soft
labels generated by the teacher network. Variants of knowledge distillation include
self-distillation \cite{zhang2019your, wang2021memory}, identical student network
distillation \cite{furlanello2018born}, channel distillation \cite{ge2019distilling},
regularizing wrong predictions \cite{yun2020regularizing} and representation or
embedding based knowledge distillation \cite{aguilar2020knowledge, yao2018deep,
passalis2018unsupervised}. Distillation is applied to various tasks such as model
compression \cite{wang2020real}, face recognition \cite{ge2018low}, enhancing ensemble
performance \cite{zhang2018deep}, interpreting neural networks \cite{liu2018improving}.
Efficacy of distillation is extensivly studied in \cite{cho2019efficacy,
yuan2020revisiting}. Applications of logit suppression include: deep metric learning
\cite{zhai2018classification}, non-parametric neighborhood component analysis
\cite{wu2018improving}, sequence-to-sequence modelling \cite{chorowski2016towards},
reinforcement learning \cite{he2018determining} and face recognition
\cite{liu2017sphereface, wang2018cosface}.

The key difference between PGT and distillation methods is that in the latter, the
transformation is applied in the forward pass, while PGT is a backward pass modification
only. Also, in distillation settings the temperature parameter is a part of the
network's computation graph. In the case of PowerGrad Transform, we directly tamper the
gradients without introducing any change in the forward pass. It is a more generalized
way of arriving at altered gradients in the backward pass. Moreover we show that
modifying the gradients in the backward pass can improve performance in many cases. The
rise in training accuracy indicates that PGT enables the network to learn more from the
training data.

Other techniques of weight update include slot machines \cite{aladago2021slot}, where
random weights are generated and carefully selected during backpropagation in a way that
minimizes the loss function.

\section{PowerGrad Transform}
\label{sec:Powe}

\begin{figure}[t]
\centering
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=0.96\columnwidth]{prob_transform_unordered}
\caption{Unordered distribution}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=0.96\columnwidth]{prob_transform_ordered}
\caption{Rank ordered distribution}
\end{subfigure}
\caption{ Illustration of the impact of PowerGrad Transform on probability values of an
arbitrary distribution (number of classes, $C=10$). $\alpha=1$ indicates the unmodified
distribution. As $\alpha$ reduces from $1$ to $0$, the probability distributions (left:
unordered or right: ordered) becomes flatter and flatter, approaching the uniform
distribution of $1/C$ for every index. }
\label{fig:prob_plots}
\end{figure}

The PowerGrad Transform technique is described in this section. It is a general
technique and can be applied to any neural network. A neural network with parameters $W$
generates $C$ logits denoted by $z$ for every input vector $x$. $z$ is given as $z=Wx$.
Then a set of probability values $p_i$ are generated from the logits using a softmax
function which is defined as $p_i=\frac{e^{z_i}}{\sum _{j=1}^C e^{z_j}}$. $p_i$ and
$z_i$ represent the predicted probability values and the logits for the $i^{th}$ class
respectively. Following this step, the loss function is invoked and the loss between the
predicted probability values the ground truth labels (which is also a probability
distribution) is calculated. If the loss function is cross-entropy loss, then the value
of the loss is given as $L=-\sum _{i=1}^C q_i \log \left(p_i\right)$ where
$q_i$ is the ground truth label of the $i^{th}$ class for a particular training example.
By standard gradient update rule, we can calculate the gradient of the loss with respect
to the logits which takes the form $\frac{\partial L}{\partial z_i}=p_i-q_i$.

The PowerGrad Transform technique is now described. We introduce a hyperparameter
$\alpha$, which takes a value between $[0, 1]$ and regulates the degree of gradient
modification. Fig. \ref{fig:prob_plots} shows the effects of the transform under
different values of $\alpha$ for different distributions. The PowerGrad Transform method
modifies the predicted probability values in the backward pass as follows:

\begin{equation} \hspace*{3.5cm} p_i'=\frac{p_i^{\alpha }}{\sum _{j=1}^C
p_j^\alpha} \ \ \ \ \ \ \ \ i=1,\dots,C\ \ \ \ \ 0 \leq \alpha\leq 1
\label{transformed_probabilities} \end{equation}

The above transformation changes the gradient of the loss with respect to the logits as
follows:

\begin{equation} \widehat{\frac{\partial L}{\partial z_i}}=p_i'-q_i
\label{PGT_logit_derivative}
\end{equation}

The rest of the backward pass proceeds as usual. We denote the original probability
distribution as $P$ (with values $p_i$ at the $i^{th}$ index) and the transformed
distribution as $P'$ (with values $p_i'$ at the $i^{th}$ index).

A code snippet of the algorithm implemented using PyTorch \cite{NEURIPS2019_9015} is
given in the appendix. In PyTorch, the available method to modify gradients in the
backward pass is using the \codeword{register_hook} function which is called in the
\codeword{forward} function. We explore the effect of this change from a theoretical
standpoint in section \ref{sec:Math}. In section \ref{sec:Expe}, we offer empirical data
and experiments to support the proposed method.

Two other methods that also changes the gradients in the backward pass are \textbf{i)}
Gradient Clipping \cite{pascanu2013difficulty} \textbf{ii)} Adaptive Gradient Clipping
\cite{brock2021high}. We introduce these two methods before delving into the inner
workings of PowerGrad Transform.

\begin{equation}
G_{i}^{l}\rightarrow
\begin{cases}
\lambda \frac{\norm{W_{i}^{l} }_{F}^{*}}{\norm{G_{i}^{l} }_{F}} G_{i}^{l} & if\ \frac{\norm{G_{i}^{l} }_{F}}{\norm{W_{i}^{l} }_{F}^{*}} \  >\ \lambda \\
G_{i}^{l} & otherwise
\end{cases}
\end{equation}

The hyperparameter $\lambda$ of AGC is less sensitive to changes in batch size and depth
as observed in \cite{brock2021high}. However, when applied to vanilla residual networks
(ResNet-18), the performance is still afar from normalized variants as demonstrated by
our experiments.

\section{Analysis of the PowerGrad transformation}
\label{sec:Math}

In this section, we first describe the properties of the  PowerGrad Transform (PGT) and
then we highlight how these suitable properties of PGT helps the neural network to
improve the performance. We use the same setup as described in section \ref{sec:Powe}.
To explore the properties of PGT, we start by investigating the effect of the transform
on the softmax probabilities.

\subsection{Properties of PGT}
\label{sec:pgt_prop}

\textbf{Lemma 1.} For any arbitrary probability distribution $P$ with probability values
given by $p_i$ for $i=1,\dots,C$, the corresponding transformed probability values
$p_i'$ given by [Eq. \ref{transformed_probabilities}] has a threshold $\Big(\sum
_{j=1}^C p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}}$ and

\begin{equation} \begin{split} p_i' \geq p_i & \text{,\ \ if } p_i \leq
\Big(\sum _{j=1}^C p_j^{\alpha }\Big)^{\frac{1}{\alpha-1}} \\ \hspace*{1cm}
p_i' < p_i & \text{,\ \ if } p_i > \Big(\sum _{j=1}^C
p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}} \end{split} \label{eqn:threshold}
\end{equation}

\textit{Proof.} The proof follows directly from the definition of $p_i'$. We notice that
if $p_i' < p_i$, we get:

\begin{proof}[\unskip\nopunct] \begin{alignat}{2} &\ \ \frac{p_i^{\alpha }}{\sum
_{j=1}^C p_j^{\alpha}} < p_i \\ \Rightarrow &\ p_i^{\alpha -1}<\sum _{j=1}^C p_j^{\alpha
} \\ \Rightarrow &\ p_i > \Big(\sum _{j=1}^C p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}}
\end{alignat} \end{proof}

We call this threshold, the \textit{stationary threshold}. The stationary threshold is
that value of $p_i$ that does not change after the transformation. Therefore, when $p_i$
is greater than the \textit{stationary threshold}, $p_i' < p_i$.

\textbf{Proposition 1.} At $\alpha=0$, the stationary threshold equals $1/C$ and all
values of the transformed distribution $p_i'$ reduces reduces to the uniform
distribution for $i=1,\dots,C$,.

\textit{Proof.} From Eq. (\ref{eqn:threshold}), we see that the stationary threshold at
$\alpha=0$ is $1/C$. Also, following from the definition of the transformed
probabilities (Eq. \ref{transformed_probabilities}) we conclude that if $\alpha=0$, then
all values of $p_i'$ are $1/C$. Therefore the transformed distribution at $\alpha=0$ is
a uniform distribution.

Since we have established that values of $p_i$ which are greater than the stationary
threshold decreases and move down towards the stationary threshold, and values in $p_i$
lower than the stationary threshold moves up towards the stationary threshold, this
transformation makes the distribution more uniform (i.e. it smooths out the actual
distribution) as $\alpha$ is decreased from $1$ and down towards $0$. This final
observation we prove in the following theorem.

\textbf{Theorem 1.} For any arbitrary probability distribution $P$ with probability
values $p_i$ for $i=1,\dots,C$, the stationary threshold of the transformed distribution
$P'$ with probability values $p_i'=\frac{p_i^{\alpha}}{\sum _{j=1}^C p_j^\alpha}, 0 \leq
\alpha \leq 1$ is a monotonically non-decreasing function with respect to $\alpha$.

\textit{Proof.} To prove monotonicity, we first compute the gradient of the stationary
threshold with respect to the variable in concern, $\alpha$.

\begin{alignat}{2} & \frac{\partial }{\partial \alpha} \left(\sum _{j=1}^c
p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ = \left(\sum _{j=1}^c
p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left(\frac{\sum _{j=1}^c p_j^{\alpha } \log
\left(p_j\right)}{(\alpha -1) \sum _{j=1}^c p_j^{\alpha}}-\frac{\log \left(\sum _{j=1}^c
p_j^{\alpha}\right)}{(\alpha -1)^2}\right) \\
& = \frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c p_j^{\alpha
}\right){}^{\frac{1}{\alpha -1}} \left(\frac{(\alpha -1) \sum _{j=1}^C p_j^{\alpha }
\log \left(p_j^{\alpha}\right)}{\sum _{j=1}^c p_j^{\alpha }}-\alpha \log \left(\sum
_{j=1}^c p_j^{\alpha }\right)\right) \label{derivative} \end{alignat}

If $a_1,\dots ,a_n$ and $b_1,\dots ,b_n$ are non-negative numbers, then using the log
sum inequality, \\ we get $\sum _{j=1}^n a_j \log \left(\frac{a_j}{b_j}\right)\geq
\left(\sum _{j=1}^n a_j\right) \log \left(\frac{\sum _{j=1}^n a_j}{\sum _{j=1}^n
b_j}\right)$. Setting $a_j=p_j^\alpha$ and $b_j=1$, we get the following lower bound
\begin{equation} \sum _{j=1}^C p_j^{\alpha } \log
\left(p_j^{\alpha }\right)\geq \left(\sum _{j=1}^C
p_j^{\alpha}\right) \log \left(\frac{1}{C} \sum _{j=1}^C
p_j^{\alpha}\right) \label{lower_bound1} \end{equation}

Substituting (\ref{lower_bound1}) in (\ref{derivative}), we get:

\begin{equation} \frac{\partial }{\partial \alpha} \left(\sum
_{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq
\frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c
p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left((1-\alpha )
\log (C)-\log \left(\sum _{j=1}^C p_j^{\alpha }\right)\right)
\label{lb_subs} \end{equation}

$p^{\alpha}$ is concave, and so by Jensen's inequality we get the following upper bound
for the second term:

\begin{equation} \left(\frac{1}{C}{\sum _{j=1}^C
p_j}{}\right){}^{\alpha }\ \ \geq\ \ \frac{1}{C}{\sum _{j=1}^C
p_j^{\alpha }}{} \end{equation}

\begin{equation}\Rightarrow\ \ \log \left(\sum _{j=1}^C p_j^{\alpha}\right)\leq
(1-\alpha ) \log (C) \label{upper_bound} \end{equation}

Substituting (\ref{upper_bound}) in (\ref{lb_subs}),

\begin{proof}[\unskip\nopunct] \begin{equation} \frac{\partial}{\partial \alpha
}\left(\sum _{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq \ \ 0
\end{equation} \end{proof}

We conclude from the analysis that the stationary threshold is a monotonic
non-decreasing function with respect to $\alpha$. Also the derivative of PGT function
with respect to the true probabilities is non-negative which in turn means that the
transformation is an order-preserving map. All values greater than the threshold move
towards the threshold after transformation and all values below the threshold also move
towards the threshold, and the threshold itself moves monotonically towards $1/C$ as
$\alpha$ is decreased from $1$ to $0$. This concludes that the transformation smooths
the original distribution.

\subsection{PGT restricts the partial derivative of the loss w.r.t. each logit from becoming too small or too high}
\label{sec:pgt_logit_derivative}

We know that neural networks uses the softmax function to generate prediction
probabilities from logits in multi-class classification tasks. If we use cross-entropy
loss to train the network then the partial derivative of the loss w.r.t. $i^{th}$ logit
is dependent on the value of the predicted probability and the value of the class label
(either 0 or 1) of the $i^{th}$ logit.

\begin{equation}
\label{actual_logit_derivative}
\frac{\partial L}{\partial z_i} = p_i -q_i
\end{equation}
In general, the range of the $|\frac{\partial L}{\partial z_i}|$ is [$\delta$,
$\epsilon$] where, $\delta\rightarrow 0$ and $\epsilon\rightarrow 1$ in traditional
training (Eq. \ref{actual_logit_derivative} procedures of neural networks. Here, we
propose a new method of neural network training which alters the gradients at the time
of backward pass by modifying the predicted probability vector using PGT (Eq.
\ref{PGT_logit_derivative}) such that the directional derivative for each logit does not
become too small or too large,

With inclusion of PGT in the last layer, the range of the $\widehat{|\frac{\partial
L}{\partial z_i}|}$ becomes [$\delta^+$, $\epsilon^-$] where, $\delta^+>\delta$ and
$\epsilon^-<\epsilon$

Now, we demonstrate how PGT restricts the magnitude of the directional derivative for
each logit. For a training example, a neural network can assign the highest probability
to either (i) a wrong class or to the (ii) correct class.

\textbf{(i) Wrong Class Predicted:}
If the network assigns the highest probability for any class except the the true class,
then wrong class assignment happens. We analyze the effect on gradient due to PGT for
this case. There are three possibilities

(a) Let, $i^{th}$ class be the class for which the network assigns low probability, but
it is the true class, then

$p_i < p_{th} \Rightarrow p'_i > p_i \Rightarrow \widehat{\frac{\partial L}{\partial
z_i}} = p'_i -1 >  p_i -1 $ ; where $p_i\rightarrow 0$ \& $q_i = 1$

So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| <  |p_i -1| \Rightarrow
|\widehat{\frac{\partial L}{\partial z_i}}| <\epsilon $ where; $\epsilon\rightarrow 1$

(b) Let, $j^{th}$ class be the class (not true class) for which the network assigns the
highest probability, then

$p_j > p_{th} \Rightarrow p'_j < p_j \Rightarrow \widehat{\frac{\partial L}{\partial
z_j}} = p'_j <  p_j$ ; where $p_j\rightarrow 1$ \& $q_j = 0$

So, $|\widehat{\frac{\partial L}{\partial z_j}}| <\epsilon$ where; $\epsilon\rightarrow
1$

(c) Let $k$ denote the index for all those classes for which the predicted probability
is low and also the ground truth class label is $0$, then

$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$

So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow$ 0

\textbf{(ii) Correct Class Predicted:}
For a given training example, if the network assigns the highest probability to the true
class then it assigns correct class to the training example. Now we observe the effect
on gradient due to PGT for this case. There are two possibilities,

(a) Let, $i^{th}$ class be the class for which the network assigns the highest
probability and it is also the true class, then

$p_i > p_{th} \Rightarrow p'_i < p_i \Rightarrow \widehat{\frac{\partial L}{\partial
z_i}} = p'_i -1 <  p_i -1 $ ; where $p_i\rightarrow 1$ \& $q_j = 1$

So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| >  |p_i -1|  \Rightarrow
|\widehat{\frac{\partial L}{\partial z_i}}| >\delta $ where; $\delta\rightarrow 0$

(b) Let, $k^{th}$ index is used for all those classes for which the predicted
probability is low and also the actual label value $0$, then

$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$

So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow 0$

After analyzing all these cases ((i)a,b,c \& (ii)a,b), we are able to deduce that PGT
helps to restrict the directional derivative to be within limit such that for every
logit/direction the directional derivative is not very small or very large.

\subsection{Effect of PGT on weight gradients}
\label{sec:pgt_weight_gradients}

\begin{figure}[t]
\centering
\includegraphics[scale=1.0,width=0.5\columnwidth]{simplenn.jpg}
\caption{Fully connected layer of a convolutional neural network} 
\label{fig:simplenn}
\end{figure}

In this section, we discuss the impact of PGT on weight gradients. Let the input and
output of the last convolutional layer of a convolutional neural network be \textbf{x}
and \textbf{a} respectively (Fig. \ref{fig:simplenn}). The last layer which is a fully
connected layer, produces logits \textbf{z} by combining \textbf{a} with weights and
thereafter the network generates the predicted probability vector \textbf{p} from the
logits using softmax activation function. We note that a$_i \geq 0$ $ \forall i$  as,
\textbf{a} is the output of ReLU activation functions.

Here, $z_i = \sum_{j=1}^{C}w_{ij}a_j$, so $\frac{\partial L}{\partial w_{ij}} =
a_j\frac{\partial L}{\partial z_{i}} $. However, if we apply PGT while training the
neural network, then $\widehat{\frac{\partial L}{\partial w_{ij}}}
=a_j\widehat{\frac{\partial L}{\partial z_i}}$

If the activation of the $j^{th}$ neuron is zero, i.e. $a_j = 0$ then there is zero
gradient for all weights which acts on $a_j$ both for normal training and training with
PGT, as $a_j$ has no contribution in the computation of logits. However, the interesting
thing is to observe the effect of PGT when $a_j > 0$. As, the activation values $(a_j)$
are positive, so $ \widehat{\frac{\partial L}{\partial z_i}}  > \frac{\partial
L}{\partial z_{i}} \Rightarrow$ $\widehat{\frac{\partial L}{\partial w_{ij}}} >
\frac{\partial L}{\partial w_{ij}} \forall j \in (1,2,\ldots C)$ and similar
relationship for $a_j < 0$. With PGT, the weight updating rule for $(t+1)^{th}$
iteration then becomes,
\begin{equation}
\label{weight_update_PGT}
w_{ij}^{t+1} = w_{ij}^{t} - \eta \widehat{\frac{\partial L}{\partial w_{ij}^{t}}}
\end{equation}
The weight update equation using PGT, can also be viewed in terms of the following
equation,
\begin{equation}
\label{weight_update_final_PGT}
w_{ij}^{t+1} = w_{ij}^{t} - \frac{\eta}{C_{PGT}} \frac{\partial L}{\partial w_{ij}^{t}}
\end{equation}

For neural networks, the exponential nature of the softmax function produces sharp
distribution for predicted probability vector. For example, if \textbf{z} is $[20, 30,
10]$ then \textbf{p} is equal to $[4\times 10^{-5},  0.99995, 2\times10^{-9}]$. When the
network is at the initial phase of training, most of the training examples are
misclassified. Under this scenario, PGT helps to update the weights in a better manner
than traditional gradients. Suppose the label vector \textbf{q} is $[1, 0, 0]$ then it
indicates that $z_1$ is the logit corresponding to ground truth class and $z_2$ is the
logit where the network has assigned the highest probability. Similarly, $z_3$ is the
logit where the network has assigned a low probability as well as the label vector is
also $0$ for this logit. Here $\frac{\partial L}{\partial w_{1j}^{t}} = \epsilon a_j > 0
\Rightarrow$ $ \widehat{\frac{\partial L}{\partial w_{1j}^{t}}} < \frac{\partial
L}{\partial w_{1j}^{t}} \Rightarrow C_{PGT} > 1$ $\forall j \in (1, 2,\ldots C)$.
Advanced gradient updation algorithms like AdaGrad \cite{duchi2011adaptive} reduces the
gradient in the direction where the directional derivative for a weight is high. It
reduces the learning rate for the weight so that the gradient in that direction is not
too large in order to make the update process less volatile. The formulation of PGT also
implicitly reduces the partial derivative in the same manner for the high derivative
directions, the only difference with the AdaGrad is that $C_{PGT}$ depends only on the
current iteration's gradient and not on previous iterations like AdaGrad. Similarly we
can observe that for this example, $\frac{\partial L}{\partial w_{3j}^{t}} = \delta a_j
\approx 0 \Rightarrow$ $  \widehat{\frac{\partial L}{\partial w_{3j}^{t}}} >
\frac{\partial L}{\partial w_{3j}^{t}} \Rightarrow C_{PGT} < 1$ $\forall j \in (1,
2,\ldots C)$. As directional derivative for the weights associated with third logit is
extremely low, PGT helps to increase the derivative in these directions so that the
training process does not become too slow.

\subsection{Effect of PGT on class separation}
\label{sec:pgt_class_separation}

During final iterations of training, we analyze the iteration-wise gradient behaviour
where the network assigns the highest probability to the correct class for a given
training example; we deduce the effects of PGT in the next immediate training iteration.
For example, if \textbf{z} is $[30, 20, 10]$ then \textbf{p} is $[0.99995, 4\times
10^{-5}, 2\times10^{-9}]$ and \textbf{q} is $[1, 0, 0]$. There is no change in the
weight update due to PGT when $a_j = 0$. However, when $i \neq 1$ and $a_j > 0$,
$|\widehat{\frac{\partial L}{\partial w_{1j}^{t}}}| > |\frac{\partial L}{\partial
w_{1j}^{t}}|$ and $\widehat{\frac{\partial L}{\partial w_{ij}^{t}}} > \frac{\partial
L}{\partial w_{ij}^{t}}$. Here in this example $\frac{\partial L}{\partial z_{1}^{t}} <
0 \Rightarrow \frac{\partial L}{\partial w_{1j}^{t}} < 0 \Rightarrow$
$\widehat{w_{1j}^{t+1}} > w_{1j}^{t+1} > w_{1j}^{t} \forall j$. We also know that
$z_i^{t+1} = \sum_{j=1}^{C}w_{ij}^{t+1}a_j$ and $\widehat{z_i^{t+1}} =
\sum_{j=1}^{C}\widehat{w_{ij}^{t+1}}a_j$. It indicates that $\widehat{z_{1}^{t+1}} >
z_{1}^{t+1} > z_{1}^{t}$ as all the weights associated with first logit is larger while
using PGT in $(t+1)^{th}$ iteration. Similarly, for the non-true classes
$\widehat{w_{ij}^{t+1}} < w_{ij}^{t+1} < w_{ij}^{t}$ when $i \neq 1$ $\forall j$ as
$\frac{\partial L}{\partial z_{i}^{t}} > 0$. So in $(t+1)^{th}$ iteration,
$\widehat{z_{i}^{t+1}} < z_{i}^{t+1} < z_{i}^{t}$ when $i \neq 1$. We observe that
$dist(\widehat{z_{1}^{t+1}}, \widehat{z_{i}^{t+1}}) > $  $dist(z_{1}^{t+1},
z_{i}^{t+1})$ where the first logit corresponds to the true class and all other logits
indexed by $i$ are non-true classes. Therefore the distance between correct and
incorrect class logits increases due to PGT and this leads to better class separation.

\section{Experiments}
\label{sec:Expe}

\subsection{Experiments on Normalized Networks}
\label{sec:bn}

\begin{table}[t]
\centering
\caption{ Results and comparisons for networks trained on ImageNet-1K. All experiments
are performed with a 100 epoch budget. }
\label{tab:imagenet_table}
\vspace{5pt}
\begin{tabular}{cccccc}
\textbf{Model} & \textbf{Scheduler} & \textbf{Method} & \textbf{Train Acc} &
\textbf{Test Acc} & \textbf{Diff} \\
\midrule
ResNet-18 & Step & - & 69.95 & 69.704 & - \\
& Step & PGT & 70.3 & \textbf{69.844} & 0.14 \\
\midrule
ResNet-18 & Cosine & - & 70.38 & 70.208 & - \\
& Cosine & PGT & 70.53 & \textbf{70.298} & 0.09 \\
\midrule
ResNet-50 & Step & - & 78.99 & 75.97 & - \\
& Step & PGT & 79.56 & \textbf{76.494} & 0.524 \\
\midrule
ResNet-50 & Cosine & - & 79.18 & 76.56 & - \\
& Cosine & PGT & 79.68 & \textbf{77.216} & 0.656 \\
\midrule
ResNet-101 & Cosine & - & 82.29 & 77.896 & - \\
& Cosine & PGT & 83.1 & \textbf{78.258} & 0.362 \\
\midrule
SEResNet-18 & Cosine & - & 71.42 & 71.09 & - \\
& Cosine & PGT & 71.6 & \textbf{71.436} & 0.346 \\
\midrule
SEResNet-50 & Cosine & - & 81.5 & 77.218 & - \\
& Cosine & PGT & 82.47 & \textbf{77.952} & 0.734 \\
\end{tabular}
\end{table}

We perform experiments on various BN variants ResNets using the ImageNet-1K dataset
\cite{deng2009imagenet}. We observe a constant $0.09\%$ to $0.14\%$ rise in ResNet-18's
test accuracy. What's more interesting is that we are seeing comparable improvements in
training accuracy. PGT increases the generalization gap, as evident from this
observation. However, it allows the network to better fit the dataset. In section
\ref{sec:Abla}, we show how combining PowerGrad Transform with regularization techniques
such as label smoothing may improve network performance and simultaneously reduce the
generalization gap.

We find significantly larger improvements in experiments using ResNet-50, a $0.52\%$
increase over the step scheduling baseline and a $0.65\%$ increase over the cosine
scheduling baseline. In studies with Squeeze-and-Excitation networks, we use the cosine
scheduler. In these experiments, we find consistent improvements across the board, with
SEResNet-18 \cite{hu2018squeeze} gaining $0.34\%$ and SEResNet-50 \cite{hu2018squeeze}
gaining $0.73\%$. As we saw with ResNet-18, we find continuous improvements in training
performance because the networks are just training better and arriving to better optimas
at convergence. ResNet-101 training provides a modest advantage of $0.36\%$.

All models are trained on four V100 GPUs with a batch size of $1024$. We utilize the
same set of hyperparameters in each experiment, which are as follows: 100 epoch budget,
5 epochs linear warmup phase beginning with a learning rate of $4\times 10^{-4}$ and
ending with a peak learning rate of $0.4$. In our studies, we employ either a step
scheduler (dividing the learning rate by $10$ at the $30^{th}$, $60^{th}$, and $90^{th}$
epochs) or a cosine decay scheduler\footnote{Reproducible code, training recipes for the
experiments, pretrained checkpoints and training logs are provided at: \\
\url{https://github.com/bishshoy/power-grad-transform}. Code is adapted from
\cite{rw2019timm}.}. We observe a consistent improvement in test metrics when utilizing
cosine scheduling across all of our experiments. Finally, towards the conclusion of
training, we impose a 10 epoch cooling phase in which the network is trained with a
constant learning rate schedule set at $4\times 10^{-4}$. Other hyperparameters include
a weight decay value of $5\times 10^{-4}$ and a momentum of $0.9$ for the SGD Nesterov
optimizer. All of our models are trained using mixed floating point precision. The
findings are shown in Table \ref{tab:imagenet_table}. With these findings, we observe
that our proposed altering gradient mechanism PGT improves the model performance further
even for batch normalized networks.

\subsection{Empirical Studies on Residual Networks without Batch Normalization}
\label{sec:Empi}

This section empirically examines some of the issues that occur when networks are
trained without normalization layers. ResNet-18 \cite{he2016deep} is used as the
foundation model for all empirical trials owing to its popularity among deep learning
practitioners and the relative simplicity with which it can be trained on a big dataset
such as ImageNet \cite{russakovsky2015imagenet}. ResNet-18 is composed of $20$
convolutional layers and one fully connected (FC) layer. At the input, it has a single
convolutional layer with filters of size $7\times 7$. Three convolutional layers are
placed in the downsampling skip connections. The remaining convolutional layers have
several $3\times 3$ filters that are grouped together to create a block. Each basic
block is connected by a skip connection. At the end of all convolutional layers, a
global average pooling layer downsamples the features to a fixed-size feature vector and
transfers it to the FC layer, which fits the features using regression and outputs
logits. The logits $z_i$ produced by the FC layer are fed into a softmax layer, which
transforms them to predicted probabilities $p_i$ using the equation $p_i=\frac
{e^{z_i}}{\sum _i e^{z_i}}$.

We concentrate on the following metric: the per-filter L2-norm of each layer. Throughout
the training process, we monitor variations in the per-filter weight norm. Because a
full plot of each layer's data would use an inordinate amount of space, we focus on a
few critical layers' plots. The supplementary section contains plots for each layer.
ResNet-18 without any normalization layers is used for all experiments in this section.
The norm of the weights of the $11\textit{th}$ convolutional layer is shown in Fig.
\ref{fig:norm_plots}a. Layer $11$ includes $256$ filters. Each colour on the figure
represents the evolution of a particular filter throughout the course of training
iterations. The critical point to note is that some filters achieve a norm of zero
during training. This event is referred to as `Zeroing Out', and occurs when a channel
of a weight tensor gets fully filled with zeros. If all the coefficients of the filter
is 0 then only L2 norm of a filter can become zero. If we observe zero filter norm for a
filter, then that filter does not contribute at all to determine the input-output
relationship of a dataset. According to our findings, when a filter tensor becomes
zeroed out, it does not recover with further training.

\begin{figure}[t]
\centering
\captionsetup{font=scriptsize}

\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{baseline-w-layer-4-2}
\caption{Norm of filters of convolutional layer 11 (without PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{baseline-w-layer-7-2}
\caption{Norm of filters of the final convolutional layer (without PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{baseline-f-layer-22-1}
\caption{Norm of final layer's output features (without PGT)}
\end{subfigure}

\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{pgt-w-layer-4-2}
\caption{Norm of filters of convolutional layer 11 (with PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{pgt-w-layer-7-2}
\caption{Norm of filters of the final convolutional layer (with PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{pgt-f-layer-22-1}
\caption{Norm of final layer's output features (with PGT)}
\end{subfigure}
\captionsetup{font=normalsize}
\caption{ Experiment demonstrating the Zeroing Out phenomena, as observed in different
filters and feature vectors. PGT greatly reduces the zeroing out of filters and promote
better learning. }
\label{fig:norm_plots}
\end{figure}

Similar to the plot \ref{fig:norm_plots}a, we see the plot of the final convolutional
layer in Fig. \ref{fig:norm_plots}b, where we observe a number of filters / units
completely dropping out towards the end of training leading to underutilization of the
network's resources. We also plot the final layer's feature norm computed in a per
channel fashion. In Fig. \ref{fig:norm_plots}c, we see the norm of the feature vector at
the output side of the global average pooling layer of ResNet-18. The feature vector
here directly interfaces with the fully connected layer immediately after. Any zeroed
out regions in this tensor directly leads to permanent information loss, as it does not
contribute to the learning of decision boundaries in the upcoming fully connected layer.

We observe both effects \textbf{1)} zeroed-out filter \textbf{2)} zeroed-out feature.
The impact of having a zeroed-out filter or a feature might have a variety of
consequences for training and optimization. If a zeroed out filter is present in a
convolutional layer, there are two possible outcomes. For starters, if the layer is
immediately followed by a fully connected regression layer, there is an abrupt loss of
information. For the reason that a zeroed out filter always generates zeroed out
features regardless of the input image, when this tensor is fed to a fully connected
layer, it does not communicate any information that is useful for classification. In the
event that the zeroed out filter has a skip connection linked across it, this may be
circumvented as long as there are no overlapping zeroed out channels in the two feature
maps.

On the other hand, if a feature tensor has a zeroed out channel at the input of a layer,
the layer creates a zeroed out feature channel at the output, regardless of the weights.
However, if there is a zeroed out channel in the feature map just before the fully
connected layer, then this results in a permanent loss of information as it does not
contribute to the learning of decision boundaries. ResNet-18 suffers from the same kind
of information loss, as the final feature tensor (which is generated after the global
average pooling layer) has multiple channels that are zeroed out [Fig.
\ref{fig:norm_plots}c]. This information collapse phenomenon occurs when a channel that
has been zeroed out stays zeroed out for all images in the dataset.

\begin{figure}[t]
\centering
\includegraphics[width=0.45\columnwidth]{baseline_high_bs-w-layer2-2}
\caption{ Per filter norm vs. iteration plot of layer 5 of an unnormalized ResNet-18,
trained with a very high batch size of 1024. The weight matrix of the entire layer
zeroes out during training. }
\label{fig:high_bs}
\end{figure}

The impact of a feature tensor or weight tensor being zeroed out is seen to be
particularly common at larger batch size training procedures in networks without
normalization layers. As a result, the network's capacity is underutilized since some of
the filters are not involved in extracting any useful information that is important for
performing classification. When working with large batch sizes, it is possible that an
entire layer zeroes out [Fig. \ref{fig:high_bs}]. In a network without skip connections,
this can cause a full collapse in training. For example if a weight tensor completely
zeroes out, and it always produces a zero feature tensor, then the input feature tensor
to the next convolutional layer is a zero tensor and thus the output of the next layer
is always a zero tensor irrespective of its weights, and this continues until the end of
the neural network pipeline leading to zero logits. Also, zeroed out weights tensors
lead to zeroed out gradients hence stopping training for all subsequent iterations
leading to a collapse in training. Although residual networks bypass this difficulty via
the use of the skip connection, it does not solve the layer's non-participation in the
overall objective of classification. It is possible to solve this issue by using methods
such as gradient clipping (GC), adaptive gradient clipping (AGC), and PowerGrad
Transform (PGT), all of which regulate the amount of gradient that flows at various
junctions of the network. GC and AGC modify the gradients at each and every filter of
every layer. PGT performs this operation at only one junction, which is at the softmax
layer.

\begin{figure}[t]
\centering
\captionsetup{font=scriptsize}

\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{baseline-w-layer-5-3}
\caption{Norm of filters of layer 15 (without AGC/PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{pgt-w-layer-5-3}
\caption{Norm of filters of layer 15 (with PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{agc_pgt-w-layer-5-3}
\caption{Norm of filters of layer 15 (with AGC+PGT)}
\end{subfigure}

\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{baseline-w-layer-1-2}
\caption{Norm of filters of layer 2 (without AGC/PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{pgt-w-layer-1-2}
\caption{Norm of filters of layer 2 (with PGT)}
\end{subfigure}
\begin{subfigure}[t]{0.32\textwidth}
\includegraphics[width=\linewidth]{agc_pgt-w-layer-1-2}
\caption{Norm of filters of layer 2 (with AGC+PGT)}
\end{subfigure}
\captionsetup{font=normalsize}
\caption{ Experiment demonstrating the efficacy of PGT and AGC+PGT, as observed in
different layers. }
\label{fig:agc_pgt}
\end{figure}

In a similar fashion to the study in section \ref{sec:Empi}, we examine the norm
profiles of features and weights in the PGT and AGC. The supplementary section includes
all the detailed graphs of each layer. In Fig. \ref{fig:norm_plots}d, we show the
per-filter norm plot of the $11\textit{th}$ convolutional layer as it progresses through
a PGT enabled training session. We observe that the number of zeroed out filters has
considerably reduced. We observe similar benefits in plot of Fig. \ref{fig:norm_plots}e,
which is the plot of the per-filter norm of the final convolutional layer. Coming to
benefits in feature norms, we observe in plot of Fig. \ref{fig:norm_plots}f, the final
feature tensor obtained with PGT enabled training do not contain any zeroed out regions,
leading us to assert that information loss is heavily mitigated as the features pass on
from the feature extracting convolutional layers to the fully connected stage for
regression.

AGC is excellent at eliminating the issue of filter and feature zeroing out. However,
AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
it also mitigates the zeroing out phenomenon in most layers. In Fig.\ref{fig:agc_pgt}a,
we see a plot of the $15\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
As we can see that there has been quite a few filter dropouts in the baseline. PGT
recovers most of the filters [Fig. \ref{fig:agc_pgt}b]. In some layers however, the
performance of PGT worsens. The $2\textit{nd}$ convolutional layer has more zeroed-out
filters in PGT [Fig. \ref{fig:agc_pgt}e] than in the baseline [Fig. \ref{fig:agc_pgt}d].
Despite that, PGT achieves a greater overall accuracy. This necessitates the
simultaneous use of AGC and PGT.

The advantage of combining GC with PGT is not significant. When AGC and PGT are
combined, we see a tremendous increase in accuracy of over $2.06\%$ over the baseline.
As seen in Fig. \ref{fig:agc_pgt}c and \ref{fig:agc_pgt}f, activating AGC and PGT
concurrently results in no filter dropouts in all layers while also improving the
accuracy by a significant margin. Additionally, we give comprehensive graphs of the norm
profiles of the AGC + PGT method for all layers in the supplementary section.

\subsection{Towards removing Normalization}
\label{sec:Towa}

\begin{table}[t]
\centering
\caption{ Results for non-normalized ResNets on ImageNet-1K. Each experiment is
conducted with a 100 epoch budget. }
\label{tab:wobn_table}
\vspace{5pt}
\begin{tabular}{cccccc}
\textbf{Model} & \textbf{Batch Size} & \textbf{Method} & \textbf{Train Acc} &
\textbf{Test Acc} & \textbf{Diff} \\
\midrule
ResNet-18 (non-BN) & 1024 & - & 67.19 & 64.93 & - \\
\midrule
ResNet-18 (non-BN) & 512 & - & 68.02 & 66.552 & - \\
& 512 & PGT & 69.5 & \textbf{67.236} & 0.684 \\
\midrule
ResNet-18 (non-BN) & 256 & - & 68.86 & 66.796 & - \\
& 256 & GC & 69.04 & 67.064 & 0.268 \\
& 256 & AGC & 69.06 & 67.298 & 0.502 \\
& 256 & PGT & 69.97 & \textbf{67.814} & 1.018 \\
& 256 & GC+PGT & 68.67 & 67.088 & 0.292 \\
& 256 & AGC+PGT & 70.92 & \textbf{68.856} & 2.06 \\
\end{tabular}
\end{table}

Experiments are carried out on non-Batch Normalized networks. We specifically utilize
the ResNet-18 model and train it on ImageNet-1K which has 1.28 million images divided
into 1000 classes. ImageNet-1K's test set consists of 50 images per class. Table
\ref{tab:wobn_table} shows the findings.

To begin, we use a batch size of 1024. This is only to illustrate that in unnormalized
networks, accuracy suffers significantly when batch sizes are large. Additionally, the
training process is often unstable, resulting in frequent layer dropouts (as seen in
\ref{sec:Empi}) and sometimes infinite loss values. Thus, for all subsequent tests, we
mostly use batch sizes of 512 and 256. We find $\alpha=0.25$ and $\alpha=0.05$ to be
mostly optimal for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$
also have good performance as well. We explore different values of $\alpha$ in section
\ref{sec:Abla}. For ResNet-18, invoking PGT with $\alpha=0.25$ improves the baseline by
$0.68\%$ when the batch size is 512, and by over $1\%$ when the batch size is 256. In
comparison, the improvement obtained by GC and AGC is much less, at $0.27\%$ and
$0.5\%$, respectively.

On the metric of training accuracy, we get a $1.48\%$ increase in training accuracy at a
batch size of 512 and an improvement of $1.11\%$ at a batch size of 256, leading us to
infer that when PowerGrad Transform is used, the network's convergence optima is
significantly superior. We also notice that the training versus testing gap increases
with PGT as compared to the gap obtained at baseline. This allows us to infer that PGT
increases the generalization gap between the training and the test set. The improvement
in training and test accuracies are obtained not through regularization like it does in
methods such as label smoothing. The improvement in both training and test accuracies
are obtained as PGT enables the network to utilize its learning capacity better and
arrive at a better optima at convergence.

\subsection{Effect on Loss, Logits and other Metrics}
\label{sec:Effe}

\begin{figure}[t]
\centering
\begin{subfigure}{.33\columnwidth}
\centering
\includegraphics[width=0.96\columnwidth]{logit_vs_alpha}
\caption{\\Logit vs. $\alpha$}
\end{subfigure}%
\begin{subfigure}{.33\columnwidth}
\centering
\includegraphics[width=0.96\columnwidth]{loss_vs_alpha}
\caption{\\Cross-entropy loss vs. $\alpha$}
\end{subfigure}%
\begin{subfigure}{.33\columnwidth}
\centering
\includegraphics[width=0.96\columnwidth]{logit_vs_loss}
\caption{\\Logit-Loss correlation}
\end{subfigure}
\caption{ Plots of various statistical measures: \textbf{(a)} Variations in the logit
norm vs. $\alpha$. The logit norm is calculated per image over the test set of
ImageNet-1K (at the linear layer of the ResNet-18 architecture) and then averaged.
\textbf{(b)} Variations in the final loss values obtained for different $\alpha$
settings. \textbf{(c)} Correlation between loss and logits. Regression line equation
:$(0.01074 x + 0.25816)$. }
\label{fig:stats}
\end{figure}

As seen in Fig. \ref{fig:stats}a, the logit norm increases as $\alpha$ falls from $1$ to
$0$. In the initial training phase, PGT causes the gradient in both the valid and wrong
classes to rise as the network misclassifies majority of the examples. Due to the
exponential nature of the softmax function, the predicted probability distribution may
frequently be close to a delta distribution, even in instances of misclassifications.
PGT reshapes the predicted probability distribution. It redistributes probability from
the confident classes to the less confident ones. Gradient rises in all classes. We also
see that the final value of the loss rises as $\alpha$ drops, and that the logit norm
and the final value of the loss are linearly correlated Fig. \ref{fig:stats}c. What is
surprising is that greater accuracies may be acquired at larger loss values, at
convergence. This is markedly different from the coupled gradient descent based training
procedures. From these plots of logits and losses we find that with the inclusion of PGT
in the training process, the model can have higher loss values at convergence compared
to the loss when not using PGT. However, this decoupled gradient based training provides
greater accuracies than the standard way of training even though the loss value is
higher.

\begin{figure}[!t]
\centering
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r18}
\caption{Plot of training and test accuracies (ResNet-18)}
\end{subfigure}%
\begin{subfigure}{.5\columnwidth}
\centering
\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r50}
\caption{Plot of training and test accuracies (ResNet-50)}
\end{subfigure}
\caption{ Log-linear plots of training and test accuracies and comparison with baseline
of batch-normalized variants: \textbf{(a)} ResNet-18 ($\alpha=0.25$), \textbf{(b)}
ResNet-50 ($\alpha=0.3$). PGT (PowerGrad Transform) improves upon baseline in all cases.
}
\label{fig:metrics}
\end{figure}

ResNet-18 and ResNet-50 per epoch training and test accuracies with and without PGT are
shown in Fig. \ref{fig:metrics}a,b. In all instances, we observe a gain. This means that
PowerGrad Transform forces the network to learn better representations.

\section{Ablation Study}
\label{sec:Abla}

\begin{table}[t]
\centering
\caption{ Ablation study for ResNet-50 on ImageNet-1K. LS column denotes whether label
smoothing has been applied. A smoothing hyperparameter value of $0.1$ has been used. The
hyperparameter column that controls the amount of PGT gradient modification injected at
softmax, eq. (\ref{transformed_probabilities}). }
\label{tab:ablation_table}
\vspace{5pt}
\begin{tabular}{ccccccc}
\textbf{Model} & \textbf{Scheduler} & \textbf{LS} & \textbf{Hparams} & \textbf{Train
Acc} & \textbf{Test Acc} & \textbf{Gap} \\
\midrule
ResNet-50 & Step &  & - & 78.99 & 75.97 & 3.02 \\
& Step &  & 0.3 & 79.56 & 76.494 & 3.066 \\
& Cosine &  & - & 79.18 & 76.56 & 2.62 \\
& Cosine & $\checkmark$ & - & 78.81 & 76.698 & 2.112 \\
& Cosine &  & 0.3 & 79.43 & 76.886 & 2.544 \\
& Cosine & $\checkmark$ & 0.3 & 78.47 & 76.968 & 1.502 \\
& Cosine &  & 0.05 & 79.68 & \textbf{77.216} & 2.464 \\
& Cosine & $\checkmark$ & 0.05 & 77.69 & 76.39 & 1.3 \\
\end{tabular}
\end{table}

We conduct an ablation study to investigate the consequences of PowerGrad Transform. We
have previously demonstrated from the experiments depicted in section \ref{sec:Expe}
that PowerGrad Transform causes the network to arrive at a better optima and fit a
particular training dataset better than the baseline. We combine our approach with other
regularization techniques such as label smoothing. The findings are shown in the Table
\ref{tab:ablation_table}. We see that adding PowerGrad Transform increases the gap
between the training and test accuracies, while introducing a label smoothing
regularizer decreases the gap. We also investigate the two optimas obtained from
ResNet-50's grid search. PGT with label smoothing improves the accuracy at $\alpha=0.3$
but it hurts the accuracy at $\alpha=0.05$.

\section{Discussions}
\label{sec:Disc}

We introduced PowerGrad Transform, which decouples the backward and forward passes of
neural network training and enables a significantly better fit to the dataset, as
measured by training and test accuracy metrics. We show the application of PowerGrad
Transform, a simple yet powerful technique for modifying the gradient flow behaviour. We
investigate experimentally the degenerate behavior of non-BN networks, which is
frequently observed during the standard training approach, especially with higher batch
sizes. With PGT, gradient behavior is enhanced and the likelihood of weights attaining
degenerate states is diminished. We provide a theoretical analysis of the PowerGrad
transformation. With the use of different network topologies and datasets, we are able
to show the potential of PGT and explore its impacts from a theoretical standpoint. We
provide comprehensive results on a number of models (non-BN and BN ResNets, SE-ResNets)
using the ImageNet dataset. PGT helps the network to improve its learning capabilities
by locating a more optimum convergence point. In addition, we undertake an ablation
research and compare its effects to those of regularization methods such label
smoothing.

\bibliographystyle{model2-names}
\bibliography{references}

\newpage

\section{Supplementary section}
\label{sec:Supp}

\subsection{Sample Code}

In PyTorch, the available method to modify gradients in the backward pass is using the
\codeword{register_hook} function which is called in the \codeword{forward} function.

\begin{figure}[h] \hrule \lstinputlisting[language=Python]{code.py} \hrule
\vspace{0.1cm} \caption{Python code of PowerGrad Transform based on PyTorch. }
\label{fig:code}
\end{figure}

\newpage \subsubsection{Norm plots of weights (baseline)} \begin{figure}[h] \centering
\includegraphics[width=0.55\textwidth]{baseline-w_norm} \caption{ Method: baseline. This
is a plot depicting the evolution of the norm of each filter in each layer. Each subplot
corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
indices are shown. Each filter in each subplot is shown with a separate and randomly
determined colour. } \end{figure}

\newpage \subsubsection{Norm plots of weights (baseline (high batch size))}
\begin{figure}[h] \centering
\includegraphics[width=0.55\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
baseline (high batch size). This is a plot depicting the evolution of the norm of each
filter in each layer. Each subplot corresponds to a layer of the unnormalized ResNet-18.
Above each subplot, the layer indices are shown. Each filter in each subplot is shown
with a separate and randomly determined colour. } \end{figure}

\newpage \subsubsection{Norm plots of weights (PGT)} \begin{figure}[h] \centering
\includegraphics[width=0.55\textwidth]{pgt-w_norm} \caption{ Method: PGT. This is a plot
depicting the evolution of the norm of each filter in each layer. Each subplot
corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
indices are shown. Each filter in each subplot is shown with a separate and randomly
determined colour. } \end{figure}

\newpage \subsubsection{Norm plots of weights (AGC + PGT)} \begin{figure}[h] \centering
\includegraphics[width=0.55\textwidth]{agc_pgt-w_norm} \caption{ Method: AGC + PGT. This
is a plot depicting the evolution of the norm of each filter in each layer. Each subplot
corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
indices are shown. Each filter in each subplot is shown with a separate and randomly
determined colour. } \end{figure}

\newpage \subsection{ResNet-18 Per Layer Index Diagram}

For the layer indices used in figures \ref{fig:norm_plots}, \ref{fig:high_bs},
\ref{fig:agc_pgt}, we provide a block diagram depicting the different convolutional
layers of ResNet-18 with corresponding layer indices in Fig. \ref{fig:resnet18}. Each
layer represents a convolutional layer, with the layer index and the number of filters
denoted alongside it. Downsampling convolutional layers in skip connections are drawn
and their layer indices are shown on the skip connection paths.

\ \\ \ \\ \ \\ \ \\ \ \\

\begin{figure}[h] \centering \includegraphics[width=0.96\columnwidth]{drawing} \caption{
ResNet-18 architecture with layer indices of convolutional layers and the number of
filters in each layer. } \label{fig:resnet18} \end{figure}

\end{document}

### 2022-09-15 code.py

@@ -1,0 +1,24 @@
+# Modification of gradients take place
+# inside the forward function
+def forward(self, x, labels):
+    # x: Input tensor with shape [N, C, H, W]
+    # labels: One hot ground truth with shape [N, C]
+    logits = self.layers(x)
+
+    def powerGradTransform(grad):
+        alpha = 0.3
+        # Recover the predicted probabilites
+        # generated in the forward pass
+        grad_temp = grad / grad.shape[0]
+        pred = grad_temp + labels
+        # Transform the predicted probabilities
+        pred = pred**alpha
+        pred /= pred.sum(-1, keepdim=True)
+        # Modify the gradient
+        modified_grad = pred - labels
+        modified_grad = modified_grad * grad.shape[0]
+        return modified_grad
+
+    if self.training:
+        logits.register_hook(powerGradTransform)
+    return logits

### 2022-09-15 paper.tex
Modifications 1
Modifications 2

@@
 training data.
+
+Self-knowledge distillation is used in many papers in different ways, which mainly
+focuses on improving the performance of the model by either creating multiple ensemble
+student models \cite{zhang2018deep}, using different distorted versions of same data
+with two different versions of a single network which guides one another
+\cite{xu2019data} or injecting multiple sub-modules in the network to improve its
+capacity \cite{zhang2019your}. However none of these methods use normal traditional
+training without introducing sub-modules in the architecture. PGT differs from
+self-knowledge distillation as it neither introduces any additional sub-modules nor
+creates different ensembles to improve the performance of the model. PGT follows the
+standard neural network training mechanism with a single change which is altering the
+last layer's gradients in the backward pass.
 
@@
 
-\newpage \subsection{ResNet-18 Per Layer Index Diagram}
-
-For the layer indices used in figures \ref{fig:norm_plots}, \ref{fig:high_bs},
+\newpage \subsection{ResNet-18 Layer Index Diagram}
+
+For the layer indices used in figures Fig. \ref{fig:norm_plots}, \ref{fig:high_bs},
 \ref{fig:agc_pgt}, we provide a block diagram depicting the different convolutional

### 2022-09-16 paper.tex
Modifications 3

@@
 point very early in the backward graph and allow this changed gradient to also affect
-the rest of the network's gradients as the it backpropagates through the backward
+the rest of the network's gradients as it backpropagates through the backward
 computation graph.
@@
 chain rule, network parameters whose gradients are connected to the point of alteration
-the computation graph also gets subsequently altered. It can be described as:
+in the computation graph also gets subsequently altered. It can be described as:
 
@@
 W_D^{t+1} &= W_D^t - \lambda\ f(\nabla_{W_D}(L)) & \\
-W_i^{t+1} &= W_i^t - \lambda\ \nabla_{W_i}(L) &\hspace{-3cm} i=D-1,\ldots,1
+W_i^{t+1} &= W_i^t - \lambda\ \nabla_{W_i}(L)^{*} &\hspace{-3cm} i=D-1,\ldots,1
 \end{align}
@@
 function `$f$' and then this transformed gradient is propagated through the rest of the
-backward graph and all other gradient vectors are computed as is. Because of the early
-injection of the transformed gradients, all other gradient vectors that are connected to
-the transformed gradient through the chain rule, gets subsequently altered.
+backward graph and all other gradient vectors $\nabla_{W_i}(L)^{*}$ are computed as is.
+Because of the early injection of the transformed gradients, all other gradient vectors
+that are connected to the transformed gradient through the chain rule, gets subsequently
+altered.
 
@@
 PowerGrad Transform modifies the gradients at the softmax layer, which is the earliest
-part of most convolutional neural network's computation graph. By changing the gradient
-at this stage, all other gradients for all other parameters in the network such as the
-linear layer, batchnorm statistics and convolutional layer's filter parameters are
-affected. We analyze and explore the behaviour of such alteration both mathematically
-and through empirical experiments conducted on networks without batch-normalization.
-Experimentally, we see that PGT improved training and test accuracies for both
-normalized and non-normalized networks. PGT improves a network's learning capacity and
-results in a better fit of the training data.
+part of most convolutional neural network's backward computation graph. By changing the
+gradient at this stage, all other gradients for all other parameters in the network such
+as the linear layer, batchnorm statistics and convolutional layer's filter parameters
+are affected. We analyze and explore the behaviour of such alteration both
+mathematically and through empirical experiments conducted on networks without
+batch-normalization. Experimentally, we see that PGT improved training and test
+accuracies for both normalized and non-normalized networks. PGT improves a network's
+learning capacity and results in a better fit of the training data.
 
@@
 passes of neural network training and enables a considerably better fit to the dataset,
-as assessed by both training and test accuracy measures. PGT is a gradient enhancement
-method that alters the gradients in the backward pass before the update step leading to
-significant boost in network's predictive performance.
+as assessed by both training and test accuracy measures. PGT is a performance
+enhancement method that alters the gradients in the backward pass before the update step
+leading to significant boost in network's predictive performance.
 
@@
 pass, then the following equation is used to modify the gradient of the $i^\textit{th}$
-filter before its update. In most cases, AGC is used for training non-BN networks. Our
-experiments demonstrate that PGT outperforms AGC, when used independently. When used
-together, it can further improve a network's performance.
+filter before its update:
+
+\begin{equation}
+G_{i}^{l}\rightarrow
+\begin{cases}
+\lambda \frac{\norm{W_{i}^{l} }_{F}^{*}}{\norm{G_{i}^{l} }_{F}} G_{i}^{l} & if\ \frac{\norm{G_{i}^{l} }_{F}}{\norm{W_{i}^{l} }_{F}^{*}} \  >\ \lambda \\
+G_{i}^{l} & otherwise
+\end{cases}
+\end{equation}
+
+The hyperparameter $\lambda$ of AGC is less sensitive to changes in batch size and depth
+as observed in \cite{brock2021high}. However, when applied to vanilla residual networks
+(ResNet-18), the performance is still afar from normalized variants as demonstrated by
+our experiments. In most cases, AGC is used for training networks without
+batch-normalization. Our experiments demonstrate that PGT outperforms AGC, when used
+independently. When used together, it can further improve a network's performance
+(section \ref{sec:Empi}).
 
@@
 effect of label smoothing. From our studies, we show that PGT in fact leads to larger
-generalization gap between the training and test dataset. Coupled with PGT, label
-smoothing can narrow the generalization gap.
+generalization gap between the training and test dataset, although it does lead to an
+increase in both training and test accuracies at convergence. Coupled with PGT, label
+smoothing can narrow the generalization gap, while retaining the benefits of PGT.
 
@@
 and experiments to support the proposed method.
-
-Two other methods that also changes the gradients in the backward pass are \textbf{i)}
-Gradient Clipping \cite{pascanu2013difficulty} \textbf{ii)} Adaptive Gradient Clipping
-\cite{brock2021high}. We introduce these two methods before delving into the inner
-workings of PowerGrad Transform.
-
-\begin{equation}
-G_{i}^{l}\rightarrow
-\begin{cases}
-\lambda \frac{\norm{W_{i}^{l} }_{F}^{*}}{\norm{G_{i}^{l} }_{F}} G_{i}^{l} & if\ \frac{\norm{G_{i}^{l} }_{F}}{\norm{W_{i}^{l} }_{F}^{*}} \  >\ \lambda \\
-G_{i}^{l} & otherwise
-\end{cases}
-\end{equation}
-
-The hyperparameter $\lambda$ of AGC is less sensitive to changes in batch size and depth
-as observed in \cite{brock2021high}. However, when applied to vanilla residual networks
-(ResNet-18), the performance is still afar from normalized variants as demonstrated by
-our experiments.
 
@@
 transformation. With the use of different network topologies and datasets, we are able
-to show the potential of PGT and explore its impacts from a theoretical standpoint. We
+to show the potential of PGT and explore its impacts from an empirical standpoint. We
 provide comprehensive results on a number of models (non-BN and BN ResNets, SE-ResNets)

### 2022-09-16 paper.tex
Modifications 4

@@
   \vspace*{-1pc}
-\fi
-
-\begin{table*}[!th]
-\ifpreprint\else\vspace*{-5pc}\fi
-
-\section*{Graphical Abstract (Optional)}
-
-\vskip1pc \fbox{
-\begin{tabular}{p{.4\textwidth}p{.5\textwidth}} \bf Altering Backward Pass Gradients
-improves Convergence  \\
-Bishshoy Das, Milton Mondal\\ Brejesh Lall, Shiv Dutt Joshi, Sumantra Dutta Roy \\[1pc]
-& In standard neural network training, the gradients in the backward pass are determined
-by the forward pass. As a result, the two stages are coupled. This is how most neural
-networks are trained currently. However, gradient modification in the backward pass has
-seldom been studied in the literature. In this paper we explore decoupled training,
-where we alter the gradients in the backward pass. We propose a simple yet powerful
-method called PowerGrad Transform, that alters the gradients before the weight update in
-the backward pass and significantly enhances the predictive performance of the neural
-network. PowerGrad Transform trains the network to arrive at a better optima at
-convergence. It is computationally extremely efficient, virtually adding no additional
-cost to either memory or compute, but results in improved final accuracies on both the
-training and test sets. PowerGrad Transform is easy to integrate into existing training
-routines, requiring just a few lines of code. PowerGrad Transform makes it possible for
-the network to better fit the training data. With decoupled training, PowerGrad
-Transform improves baseline accuracies for ResNet-50 by 0.73\%, for SE-ResNet-50 by
-0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network on the ImageNet
-classification task.
-\end{tabular}
-}
-
-\end{table*}
-
-\clearpage
-\thispagestyle{empty}
-
-\ifpreprint
-  \vspace*{-1pc}
-\else
-%  \vspace*{-6pc}
 \fi
@@
 function `$f$' and then this transformed gradient is propagated through the rest of the
-backward graph and all other gradient vectors $\nabla_{W_i}(L)^{*}$ are computed as is.
-Because of the early injection of the transformed gradients, all other gradient vectors
-that are connected to the transformed gradient through the chain rule, gets subsequently
-altered.
+backward graph. All other gradient vectors $\nabla_{W_i}(L)^{*}$ are computed as is, but
+because of the early injection of the transformed gradient $\nabla_{W_D}(L)$, all other
+gradient vectors that are connected to the transformed gradient through the chain rule
+($\nabla_{W_i}(L)^{*}$,\ \ i=D-1,\ldots,1), gets subsequently altered.
 

### 2022-09-16 paper.tex
Modifications 5.

@@
 transforms them to predicted probabilities $p_i$ using the equation $p_i=\frac
-{e^{z_i}}{\sum _i e^{z_i}}$.
+{e^{z_i}}{\sum _j e^{z_j}}$.
 
@@
 ResNet-18 without any normalization layers is used for all experiments in this section.
-The norm of the weights of the $11\textit{th}$ convolutional layer is shown in Fig.
+The norm of the weights of the $11^\textit{th}$ convolutional layer is shown in Fig.
 \ref{fig:norm_plots}a. Layer $11$ includes $256$ filters. Each colour on the figure
@@
 all the detailed graphs of each layer. In Fig. \ref{fig:norm_plots}d, we show the
-per-filter norm plot of the $11\textit{th}$ convolutional layer as it progresses through
+per-filter norm plot of the $11^\textit{th}$ convolutional layer as it progresses through
 a PGT enabled training session. We observe that the number of zeroed out filters has
@@
 it also mitigates the zeroing out phenomenon in most layers. In Fig.\ref{fig:agc_pgt}a,
-we see a plot of the $15\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
+we see a plot of the $15^\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
 As we can see that there has been quite a few filter dropouts in the baseline. PGT

### 2022-09-18 paper.tex
Mod6

@@
 \usepackage{amsthm}
+\usepackage{pifont}
 \usepackage{graphicx}
+
+% \usepackage{xcolor}
+% \usepackage{pagecolor}
+% \pagecolor{black}
+% \color{white}
 
@@
 \NewDocumentCommand{\codeword}{v}{\texttt{\textcolor{orange}{#1}}}
+\newcommand{\cmark}{\ding{51}} \newcommand{\xmark}{\ding{55}}
 
@@
 
-\item If \underline {presented} at or \underline {submitted} to or \underline
-{published }at a conference(s), the conference(s) is (are) identified and  substantial
-\underline {justification for re-publication} is presented  below. A \underline {copy of
-conference paper(s) }is(are) uploaded with the  manuscript.
+\item If \underline {presented} at or \underline {submitted} to or \underline {published
+}at a conference(s), the conference(s) is (are) identified and  substantial \underline
+{justification for re-publication} is presented  below. A \underline {copy of conference
+paper(s) }is(are) uploaded with the  manuscript.
 
@@
 standard calculus rules and use the chain rule to calculate the gradients of the rest of
-the network's parameters, also known as backpropagation (Fig.
-\ref{fig:gradient_altering}a). The network is then updated with the gradient descent
+the network's parameters, also known as backpropagation as portrayed in Fig.
+\ref{fig:gradient_altering}(a). The network is then updated with the gradient descent
 equation:
@@
 the gradients are first computed using standard procedure and then individually altered
-(Fig. \ref{fig:gradient_altering}b). Gradient clipping \cite{pascanu2013difficulty} and
-Adaptive gradient clipping \cite{brock2021high} are examples of such modifications. In
-both of these methods, the gradients are first computed using standard rules and then
-they are modified using some function. It can be described as:
+as shown in Fig. \ref{fig:gradient_altering}(b). Gradient clipping
+\cite{pascanu2013difficulty} and Adaptive gradient clipping \cite{brock2021high} are
+examples of such modifications. In both of these methods, the gradients are first
+computed using standard rules and then they are modified using some function. It can be
+described as:
 
@@
 computation graph and then all subsequent gradients are generated using the values
-obtained with the modified gradients (Fig. \ref{fig:gradient_altering}c). Because of the
-chain rule, network parameters whose gradients are connected to the point of alteration
-in the computation graph also gets subsequently altered. It can be described as:
+obtained with the modified gradients. We illustrate this type of modification in Fig.
+\ref{fig:gradient_altering}(c). Because of the chain rule, network parameters whose
+gradients are connected to the point of alteration in the computation graph also gets
+subsequently altered. It can be described as:
 
@@
 
-The formulation is as follows: If $G$ is a gradient vector which is $\partial
-L / \partial \theta$ (the gradient of the loss $L$ with respect to
-the parameters $\theta$). In clip by value, the gradients are modified as:
+The formulation is as follows: If $G$ is a gradient vector which is $\partial L /
+\partial \theta$ (the gradient of the loss $L$ with respect to the parameters $\theta$).
+In clip by value, the gradients are modified as:
 
@@
 distribution) is calculated. If the loss function is cross-entropy loss, then the value
-of the loss is given as $L=-\sum _{i=1}^C q_i \log \left(p_i\right)$ where
-$q_i$ is the ground truth label of the $i^{th}$ class for a particular training example.
-By standard gradient update rule, we can calculate the gradient of the loss with respect
-to the logits which takes the form $\frac{\partial L}{\partial z_i}=p_i-q_i$.
+of the loss is given as $L=-\sum _{i=1}^C q_i \log \left(p_i\right)$ where $q_i$ is the
+ground truth label of the $i^{th}$ class for a particular training example. By standard
+gradient update rule, we can calculate the gradient of the loss with respect to the
+logits which takes the form $\frac{\partial L}{\partial z_i}=p_i-q_i$.
 
@@
 
-\subsection{Experiments on Normalized Networks}
+\subsection{Experiments on convolutional neural networks (with batch-normalization)}
 \label{sec:bn}
 
-\begin{table}[t]
-\centering
-\caption{ Results and comparisons for networks trained on ImageNet-1K. All experiments
-are performed with a 100 epoch budget. }
+\begin{figure}[!t]
+\centering
+\begin{subfigure}{.5\columnwidth}
+\centering
+\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r18}
+\caption{Plot of training and test accuracies (ResNet-18)}
+\end{subfigure}%
+\begin{subfigure}{.5\columnwidth}
+\centering
+\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r50}
+\caption{Plot of training and test accuracies (ResNet-50)}
+\end{subfigure}
+\caption{ Log-linear plots of training and test accuracies and comparison with baseline
+of batch-normalized variants: \textbf{(a)} ResNet-18 ($\alpha=0.25$), \textbf{(b)}
+ResNet-50 ($\alpha=0.3$). For better visualization of accuracies obtained in the final
+iterations, we plot the y-axis in log scale. PowerGrad Transform improves upon the
+baseline in all cases. The corresponding accuracies are mentioned in table
+\ref{tab:imagenet_table}. }
+\label{fig:metrics}
+\end{figure}
+
+\begin{table}[!t]
+\centering
+\caption{ Results and comparison table for CNN networks trained on ImageNet-1K (sorted
+in descending order w.r.t. the last column denoting test accuracy improvements over
+baseline). Best training and test accuracies are highlighted in red and blue
+respectively. Accuracy differences are highlighted in yellow. In all experiments, we
+observe that PGT positively improves both training and test accuracies. We also observe
+that deeper networks such as SE-ResNet-50 and ResNet-50 accrue higher gains over the
+baselines as compared to shallow networks like ResNet-18. }
 \label{tab:imagenet_table}
-\vspace{5pt}
-\begin{tabular}{cccccc}
-\textbf{Model} & \textbf{Scheduler} & \textbf{Method} & \textbf{Train Acc} &
-\textbf{Test Acc} & \textbf{Diff} \\
+\begin{tabular}{cccccccc}
+\multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Scheduler}} &
+\multirow{2}{*}{\textbf{Method}} & \textbf{PowerGrad} & \textbf{\hspace{-0.65cm} Train}
+& \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
+& & & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} & \textbf{Diff (\%)} &
+\textbf{Accuracy (\%)} & \textbf{Diff (\%)} \\
 \midrule
-ResNet-18 & Step & - & 69.95 & 69.704 & - \\
-& Step & PGT & 70.3 & \textbf{69.844} & 0.14 \\
+\multirow{2}{*}{SE-ResNet-50} & \multirow{2}{*}{Cosine} & Baseline & - & 81.5 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.97}}} & 77.218 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.734}}} \\
+& & PGT & 0.3 & \textcolor{red}{\textbf{82.47}} & & \textcolor{blue}{\textbf{77.952}} & \\
 \midrule
-ResNet-18 & Cosine & - & 70.38 & 70.208 & - \\
-& Cosine & PGT & 70.53 & \textbf{70.298} & 0.09 \\
+\multirow{2}{*}{ResNet-50} & \multirow{2}{*}{Cosine} & Baseline & - & 79.18 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.5}}} & 76.56 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.656}}} \\
+& & PGT & 0.05 & \textcolor{red}{\textbf{79.68}} & & \textcolor{blue}{\textbf{77.216}} & \\
 \midrule
-ResNet-50 & Step & - & 78.99 & 75.97 & - \\
-& Step & PGT & 79.56 & \textbf{76.494} & 0.524 \\
+\multirow{2}{*}{ResNet-50} & \multirow{2}{*}{Step} & Baseline & - & 78.99 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.57}}} & 75.97 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.524}}} \\
+& & PGT & 0.05 & \textcolor{red}{\textbf{79.56}} & & \textcolor{blue}{\textbf{76.494}} & \\
 \midrule
-ResNet-50 & Cosine & - & 79.18 & 76.56 & - \\
-& Cosine & PGT & 79.68 & \textbf{77.216} & 0.656 \\
+\multirow{2}{*}{ResNet-101} & \multirow{2}{*}{Cosine} & Baseline & - & 82.29 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.81}}} & 77.896 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.362}}} \\
+& & PGT & 0.3 & \textcolor{red}{\textbf{83.1}} & & \textcolor{blue}{\textbf{78.258}} & \\
 \midrule
-ResNet-101 & Cosine & - & 82.29 & 77.896 & - \\
-& Cosine & PGT & 83.1 & \textbf{78.258} & 0.362 \\
+\multirow{2}{*}{SE-ResNet-18} & \multirow{2}{*}{Cosine} & Baseline & - & 71.42 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.18}}} & 71.09 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.346}}} \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{71.6}} & & \textcolor{blue}{\textbf{71.436}} & \\
 \midrule
-SEResNet-18 & Cosine & - & 71.42 & 71.09 & - \\
-& Cosine & PGT & 71.6 & \textbf{71.436} & 0.346 \\
+\multirow{2}{*}{ResNet-18} & \multirow{2}{*}{Step} & Baseline & - & 69.95 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.35}}} & 69.704 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.14}}} \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{70.3}} & & \textcolor{blue}{\textbf{69.844}} & \\
 \midrule
-SEResNet-50 & Cosine & - & 81.5 & 77.218 & - \\
-& Cosine & PGT & 82.47 & \textbf{77.952} & 0.734 \\
+\multirow{2}{*}{ResNet-18} & \multirow{2}{*}{Cosine} & Baseline & - & 70.38 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.15}}} & 70.208 &
+\textcolor{olive}{\multirow{2}{*}{\textbf{+0.09}}} \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{70.53}} & & \textcolor{blue}{\textbf{70.298}} & \\
 \end{tabular}
@@
 
-We perform experiments on various BN variants ResNets using the ImageNet-1K dataset
-\cite{deng2009imagenet}. We observe a constant $0.09\%$ to $0.14\%$ rise in ResNet-18's
-test accuracy. What's more interesting is that we are seeing comparable improvements in
-training accuracy. PGT increases the generalization gap, as evident from this
-observation. However, it allows the network to better fit the dataset. In section
-\ref{sec:Abla}, we show how combining PowerGrad Transform with regularization techniques
-such as label smoothing may improve network performance and simultaneously reduce the
-generalization gap.
-
-We find significantly larger improvements in experiments using ResNet-50, a $0.52\%$
-increase over the step scheduling baseline and a $0.65\%$ increase over the cosine
-scheduling baseline. In studies with Squeeze-and-Excitation networks, we use the cosine
-scheduler. In these experiments, we find consistent improvements across the board, with
-SEResNet-18 \cite{hu2018squeeze} gaining $0.34\%$ and SEResNet-50 \cite{hu2018squeeze}
-gaining $0.73\%$. As we saw with ResNet-18, we find continuous improvements in training
-performance because the networks are just training better and arriving to better optimas
-at convergence. ResNet-101 training provides a modest advantage of $0.36\%$.
-
-All models are trained on four V100 GPUs with a batch size of $1024$. We utilize the
-same set of hyperparameters in each experiment, which are as follows: 100 epoch budget,
-5 epochs linear warmup phase beginning with a learning rate of $4\times 10^{-4}$ and
-ending with a peak learning rate of $0.4$. In our studies, we employ either a step
-scheduler (dividing the learning rate by $10$ at the $30^{th}$, $60^{th}$, and $90^{th}$
-epochs) or a cosine decay scheduler\footnote{Reproducible code, training recipes for the
-experiments, pretrained checkpoints and training logs are provided at: \\
-\url{https://github.com/bishshoy/power-grad-transform}. Code is adapted from
-\cite{rw2019timm}.}. We observe a consistent improvement in test metrics when utilizing
-cosine scheduling across all of our experiments. Finally, towards the conclusion of
-training, we impose a 10 epoch cooling phase in which the network is trained with a
-constant learning rate schedule set at $4\times 10^{-4}$. Other hyperparameters include
-a weight decay value of $5\times 10^{-4}$ and a momentum of $0.9$ for the SGD Nesterov
-optimizer. All of our models are trained using mixed floating point precision. The
-findings are shown in Table \ref{tab:imagenet_table}. With these findings, we observe
-that our proposed altering gradient mechanism PGT improves the model performance further
-even for batch normalized networks.
-
-\subsection{Empirical Studies on Residual Networks without Batch Normalization}
+We perform experiments on different variants ResNets using the ImageNet-1K dataset
+\cite{deng2009imagenet}. All models are trained on four V100 GPUs with a batch size of
+$1024$. We utilize a common set of hyperparameters for all experiments, which are as
+follows: 100 epoch budget, 5 epochs linear warmup phase beginning with a learning rate
+of $4\times 10^{-4}$ and ending with a peak learning rate of $0.4$, a momentum of $0.9$
+and weight decay of $5\times 10^{-4}$, the SGD Nesterov optimizer and mixed precision.
+In our studies, we employ either a step scheduler (dividing the learning rate by $10$ at
+the $30^{th}$, $60^{th}$, and $90^{th}$ epochs) or a cosine decay scheduler
+\cite{loshchilov2016sgdr}. We find $\alpha=0.25$ and $\alpha=0.05$ to be good choices
+for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$ also have good
+performance as well. The experimental results are shown in table
+\ref{tab:imagenet_table}, and we mention the value of the PGT hyperparmeter ($\alpha$)
+in each experiment. We explore different values of $\alpha$ in section \ref{sec:Abla}
+and provide detailed grid plots in the supplementary section.
+
+In our experiments with Squeeze-and-Excitation variant of ResNet-50 i.e.
+SE-ResNet-50\cite{hu2018squeeze}, we observe significant improvements; a $0.97\%$ boost
+in training accuracy and a $0.734\%$ increase in test accuracy. These improvements are
+obtained with the number of epochs being kept the same as that of the baseline model.
+Alternatively, practioners can choose to employ PGT and have the networks converge to
+accuracies similar to that of the baseline with a much lower epoch budget. We explore
+this in detail in section \ref{sec:speed}. Therefore, PGT also speeds up training. In
+our experiments with ResNet-50, we find an increase of $0.5\%$ and $0.656\%$ performance
+enhancement over the baseline while using the cosine scheduler. The corresponding
+improvement over the step scheduler baseline for ResNet-50 is $0.57\%$ and $0.524\%$
+respectively for training and testing. ResNet-101 sees a higher improvement in training
+fit, $0.81\%$ to be exact, while the improvement over the test set is $0.362\%$. Smaller
+networks such as SE-ResNet-18 and ResNet-18 sees accuracy boosts which are smaller but
+nevertheless positive. Therefore we conclude that with PGT, the consistent improvements
+in training accuracies across all cases is because the networks train better and arrive
+at better optimas during convergence. Per epoch training and test accuracy plots of
+ResNet-18 and ResNet-50 (both with and without PGT) are shown in Fig.
+\ref{fig:metrics}(a, b). For better visualization of the top end of the plots, we scale
+the y-axis in log scale. Our experiments with batch-normalized networks indicate that
+PowerGrad Transform speeds up training and allows the network to learn better
+representations\footnote{Reproducible code, training recipes for the experiments,
+pretrained checkpoints and training logs are provided at: \\
+\url{ https://github.com/bishshoy/power-grad-transform }. Code is adapted from
+\cite{rw2019timm}.}.
+
+\subsection{Empirical studies on networks without Batch Normalization}
 \label{sec:Empi}
@@
 This section empirically examines some of the issues that occur when networks are
-trained without normalization layers. ResNet-18 \cite{he2016deep} is used as the
+trained without normalization layers and we provide insight into the effects of
+PowerGrad transform on such networks. We use ResNet-18 \cite{he2016deep} as the
 foundation model for all empirical trials owing to its popularity among deep learning
 practitioners and the relative simplicity with which it can be trained on a big dataset
-such as ImageNet \cite{russakovsky2015imagenet}. ResNet-18 is composed of $20$
-convolutional layers and one fully connected (FC) layer. At the input, it has a single
-convolutional layer with filters of size $7\times 7$. Three convolutional layers are
-placed in the downsampling skip connections. The remaining convolutional layers have
-several $3\times 3$ filters that are grouped together to create a block. Each basic
-block is connected by a skip connection. At the end of all convolutional layers, a
-global average pooling layer downsamples the features to a fixed-size feature vector and
-transfers it to the FC layer, which fits the features using regression and outputs
-logits. The logits $z_i$ produced by the FC layer are fed into a softmax layer, which
-transforms them to predicted probabilities $p_i$ using the equation $p_i=\frac
-{e^{z_i}}{\sum _j e^{z_j}}$.
-
-We concentrate on the following metric: the per-filter L2-norm of each layer. Throughout
-the training process, we monitor variations in the per-filter weight norm. Because a
-full plot of each layer's data would use an inordinate amount of space, we focus on a
-few critical layers' plots. The supplementary section contains plots for each layer.
-ResNet-18 without any normalization layers is used for all experiments in this section.
-The norm of the weights of the $11^\textit{th}$ convolutional layer is shown in Fig.
-\ref{fig:norm_plots}a. Layer $11$ includes $256$ filters. Each colour on the figure
-represents the evolution of a particular filter throughout the course of training
-iterations. The critical point to note is that some filters achieve a norm of zero
-during training. This event is referred to as `Zeroing Out', and occurs when a channel
-of a weight tensor gets fully filled with zeros. If all the coefficients of the filter
-is 0 then only L2 norm of a filter can become zero. If we observe zero filter norm for a
-filter, then that filter does not contribute at all to determine the input-output
-relationship of a dataset. According to our findings, when a filter tensor becomes
-zeroed out, it does not recover with further training.
+such as ImageNet-1K \cite{russakovsky2015imagenet}. We found deeper networks (such as
+ResNet-34 and ResNet-50 models) are impossible to train without BatchNormalization
+because of increased depth. So in order to conduct our empirical experiments, we solely
+focus on the ResNet-18 architecture. For our experiments, we designate different layers
+with their corresponding layer indices. We provide a detailed layer-wise index list of
+ResNet-18 in the supplementary section.
+
+We concentrate on the following metric: the per-filter L2-norm of each layer's weight
+tensor. Throughout the training process, we monitor variations in the per-filter weight
+norm. Because a full plot of each layer's data would use an inordinate amount of space,
+we focus on a few critical layers' plots. The supplementary section contains full plots
+for all layer for various settings. The norm of the weights of the $11^\textit{th}$
+convolutional layer is shown in Fig. \ref{fig:norm_plots}(a). Layer $11$ includes $256$
+filters. Each colour on the figure represents the iteration-wise evolution of a
+particular filter throughout the course of training. The critical point to note is that
+some filters achieve a norm of zero during training. We refer to this event as `Zeroing
+Out', and occurs when a channel of a weight tensor gets fully filled with zeros. If all
+the coefficients of the filter becomes zero, only then the L2 norm of that filter
+becomes zero. If we observe zero filter norm for a filter, then that filter does not
+contribute at all to determine the input-output relationship of a dataset, as the
+feature tensor it produces is also an all-zero tensor. According to our findings, when a
+filter tensor becomes zeroed out, it does not recover with further training. We observe
+both effects \textbf{1)} zeroed out filter \textbf{2)} zeroed out feature.
 
@@
 
-Similar to the plot \ref{fig:norm_plots}a, we see the plot of the final convolutional
-layer in Fig. \ref{fig:norm_plots}b, where we observe a number of filters / units
+Similar to the plot \ref{fig:norm_plots}(a), we see the plot of the final convolutional
+layer in Fig. \ref{fig:norm_plots}(b), where we observe a number of filters / units
 completely dropping out towards the end of training leading to underutilization of the
-network's resources. We also plot the final layer's feature norm computed in a per
-channel fashion. In Fig. \ref{fig:norm_plots}c, we see the norm of the feature vector at
-the output side of the global average pooling layer of ResNet-18. The feature vector
-here directly interfaces with the fully connected layer immediately after. Any zeroed
-out regions in this tensor directly leads to permanent information loss, as it does not
-contribute to the learning of decision boundaries in the upcoming fully connected layer.
-
-We observe both effects \textbf{1)} zeroed-out filter \textbf{2)} zeroed-out feature.
-The impact of having a zeroed-out filter or a feature might have a variety of
-consequences for training and optimization. If a zeroed out filter is present in a
-convolutional layer, there are two possible outcomes. For starters, if the layer is
-immediately followed by a fully connected regression layer, there is an abrupt loss of
-information. For the reason that a zeroed out filter always generates zeroed out
+network's parameters. In Fig. \ref{fig:norm_plots}(c), we see the norm of the feature
+vector at the output side of the global average pooling layer of ResNet-18. The feature
+vector here directly interfaces with the fully connected layer immediately after. Any
+zeroed out regions in this tensor directly leads to permanent information loss, as it
+does not contribute to the learning of decision boundaries in the upcoming fully
+connected layer, and indeed there exists a few zeroed out features in the baseline
+model, which we discuss next. The impact of having a zeroed out filter or a feature
+might have a variety of consequences for training and optimization. If a zeroed out
+filter is present in a convolutional layer, there are two possible outcomes. If the
+layer is immediately followed by a fully connected regression layer, there is an abrupt
+loss of information. For the reason that a zeroed out filter always generates zeroed out
 features regardless of the input image, when this tensor is fed to a fully connected
@@
 circumvented as long as there are no overlapping zeroed out channels in the two feature
-maps.
-
-On the other hand, if a feature tensor has a zeroed out channel at the input of a layer,
-the layer creates a zeroed out feature channel at the output, regardless of the weights.
-However, if there is a zeroed out channel in the feature map just before the fully
-connected layer, then this results in a permanent loss of information as it does not
-contribute to the learning of decision boundaries. ResNet-18 suffers from the same kind
-of information loss, as the final feature tensor (which is generated after the global
-average pooling layer) has multiple channels that are zeroed out [Fig.
-\ref{fig:norm_plots}c]. This information collapse phenomenon occurs when a channel that
-has been zeroed out stays zeroed out for all images in the dataset.
+maps. On the other hand, if a feature tensor has a zeroed out channel at the input of a
+layer, the layer creates a zeroed out feature channel at the output, regardless of the
+weights. However, if there is a zeroed out channel in the feature map just before the
+fully connected layer, then this results in a permanent loss of information as it does
+not contribute to the learning of decision boundaries. ResNet-18 suffers from the same
+kind of information loss, as the final feature tensor (which is generated after the
+global average pooling layer) has multiple channels that are zeroed out [Fig.
+\ref{fig:norm_plots}(c)]. This information collapse phenomenon occurs when a channel
+that has been zeroed out stays zeroed out for all images in the dataset.
 
@@
 The impact of a feature tensor or weight tensor being zeroed out is seen to be
-particularly common at larger batch size training procedures in networks without
-normalization layers. As a result, the network's capacity is underutilized since some of
-the filters are not involved in extracting any useful information that is important for
-performing classification. When working with large batch sizes, it is possible that an
-entire layer zeroes out [Fig. \ref{fig:high_bs}]. In a network without skip connections,
-this can cause a full collapse in training. For example if a weight tensor completely
-zeroes out, and it always produces a zero feature tensor, then the input feature tensor
-to the next convolutional layer is a zero tensor and thus the output of the next layer
-is always a zero tensor irrespective of its weights, and this continues until the end of
-the neural network pipeline leading to zero logits. Also, zeroed out weights tensors
-lead to zeroed out gradients hence stopping training for all subsequent iterations
-leading to a collapse in training. Although residual networks bypass this difficulty via
-the use of the skip connection, it does not solve the layer's non-participation in the
-overall objective of classification. It is possible to solve this issue by using methods
-such as gradient clipping (GC), adaptive gradient clipping (AGC), and PowerGrad
-Transform (PGT), all of which regulate the amount of gradient that flows at various
-junctions of the network. GC and AGC modify the gradients at each and every filter of
-every layer. PGT performs this operation at only one junction, which is at the softmax
-layer.
+particularly common in training procedures involving larger batch sizes in networks
+without normalization layers. As a result, the network's parameters are underutilized
+since some of the filters are not involved in extracting any useful information that is
+important for performing classification. When working with large batch sizes, it is
+possible that an entire layer zeroes out [Fig. \ref{fig:high_bs}]. In a network without
+skip connections, this can cause a full collapse in training. For example, if a weight
+tensor completely zeroes out, and it always produces a zero feature tensor, then the
+input feature tensor to the next convolutional layer is a zero tensor and thus the
+output of the next layer is always a zero tensor irrespective of its weights, and this
+continues until the end of the neural network pipeline leading to zero logits. Also,
+zeroed out weights tensors lead to zeroed out gradients hence stopping training for all
+subsequent iterations leading to a collapse in training. Although residual networks
+bypass this difficulty via the use of the skip connection, it does not solve the layer's
+non-participation in the overall objective of classification. It is possible to solve
+this issue by using methods such as gradient clipping (GC), adaptive gradient clipping
+(AGC), and PowerGrad Transform (PGT), all of which regulate the amount of gradient that
+flows at various junctions of the network. GC and AGC modify the gradients at each and
+every filter of every layer. PGT performs this operation at only one junction, which is
+at the softmax layer.
 
@@
 profiles of features and weights in the PGT and AGC. The supplementary section includes
-all the detailed graphs of each layer. In Fig. \ref{fig:norm_plots}d, we show the
-per-filter norm plot of the $11^\textit{th}$ convolutional layer as it progresses through
-a PGT enabled training session. We observe that the number of zeroed out filters has
-considerably reduced. We observe similar benefits in plot of Fig. \ref{fig:norm_plots}e,
-which is the plot of the per-filter norm of the final convolutional layer. Coming to
-benefits in feature norms, we observe in plot of Fig. \ref{fig:norm_plots}f, the final
-feature tensor obtained with PGT enabled training do not contain any zeroed out regions,
-leading us to assert that information loss is heavily mitigated as the features pass on
-from the feature extracting convolutional layers to the fully connected stage for
-regression.
+all the detailed graphs of all layer. In Fig. \ref{fig:norm_plots}(d), we show the
+per-filter norm plot of the $11^\textit{th}$ convolutional layer as it progresses
+through a PGT enabled training session. We observe that the number of zeroed out filters
+has considerably reduced. We observe similar benefits in plot of Fig.
+\ref{fig:norm_plots}(e), which is the plot of the per-filter norm of the final
+convolutional layer. Coming to benefits in feature norms, we observe in plot of Fig.
+\ref{fig:norm_plots}(f), the final feature tensor obtained with PGT enabled training do
+not contain any zeroed out regions, leading us to assert that information loss is
+heavily mitigated as the features pass on from the feature extracting convolutional
+layers to the fully connected stage for regression.
 
@@
 AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
-it also mitigates the zeroing out phenomenon in most layers. In Fig.\ref{fig:agc_pgt}a,
-we see a plot of the $15^\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
-As we can see that there has been quite a few filter dropouts in the baseline. PGT
-recovers most of the filters [Fig. \ref{fig:agc_pgt}b]. In some layers however, the
-performance of PGT worsens. The $2\textit{nd}$ convolutional layer has more zeroed-out
-filters in PGT [Fig. \ref{fig:agc_pgt}e] than in the baseline [Fig. \ref{fig:agc_pgt}d].
-Despite that, PGT achieves a greater overall accuracy. This necessitates the
-simultaneous use of AGC and PGT.
-
-The advantage of combining GC with PGT is not significant. When AGC and PGT are
-combined, we see a tremendous increase in accuracy of over $2.06\%$ over the baseline.
-As seen in Fig. \ref{fig:agc_pgt}c and \ref{fig:agc_pgt}f, activating AGC and PGT
-concurrently results in no filter dropouts in all layers while also improving the
-accuracy by a significant margin. Additionally, we give comprehensive graphs of the norm
-profiles of the AGC + PGT method for all layers in the supplementary section.
-
-\subsection{Towards removing Normalization}
-\label{sec:Towa}
-
-\begin{table}[t]
+it also mitigates the zeroing out phenomenon in most layers. In
+Fig.\ref{fig:agc_pgt}(a), we see a plot of the $15^\textit{th}$ layer of the non-BN
+ResNet-18 trained without PGT. As we can see that there has been quite a few filter
+dropouts in the baseline. PGT recovers most of the filters [Fig. \ref{fig:agc_pgt}(b)].
+In some layers however, the performance of PGT is slightly weaker than the baseline. The
+$2^\textit{nd}$ convolutional layer has more zeroed out filters in PGT [Fig.
+\ref{fig:agc_pgt}(e)] than in the baseline [Fig. \ref{fig:agc_pgt}(d)]. Despite that,
+PGT achieves a higher overall accuracy (table \ref{tab:wobn_table}). This necessitates
+the simultaneous use of AGC and PGT, for the non-BN ResNet-18.
+
+\begin{table}[!t]
 \centering
 \caption{ Results for non-normalized ResNets on ImageNet-1K. Each experiment is
-conducted with a 100 epoch budget. }
+conducted with a 100 epoch budget using the cosine decay scheduler. Best training and
+test accuracies are highlighted in red and blue respectively. Top differences in
+training and test accuracies are marked in yellow. The baselines in each segment is
+taken as the reference point for all `Diff' accuracies in that segment. The first three
+segments highlight comparison for experiments where the methods are applied
+individually, while the last segment denotes performance metrics for combinations of
+different methods (GC + PGT, AGC + PGT). We see that at any batch size, with PGT, the
+performance of non-normalized version of ResNet-18 improves significantly over the
+baseline owing to the improvements in training as observed in Fig. \ref{fig:agc_pgt}. }
 \label{tab:wobn_table}
-\vspace{5pt}
-\begin{tabular}{cccccc}
-\textbf{Model} & \textbf{Batch Size} & \textbf{Method} & \textbf{Train Acc} &
-\textbf{Test Acc} & \textbf{Diff} \\
+\begin{tabular}{cccccccc}
+\multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Batch Size}} &
+\multirow{2}{*}{\textbf{Method}} & \textbf{PowerGrad} & \textbf{\hspace{-0.65cm} Train}
+& \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
+& & & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} & \textbf{Diff (\%)} &
+\textbf{Accuracy (\%)} & \textbf{Diff (\%)} \\
 \midrule
-ResNet-18 (non-BN) & 1024 & - & 67.19 & 64.93 & - \\
-\midrule
-ResNet-18 (non-BN) & 512 & - & 68.02 & 66.552 & - \\
-& 512 & PGT & 69.5 & \textbf{67.236} & 0.684 \\
-\midrule
-ResNet-18 (non-BN) & 256 & - & 68.86 & 66.796 & - \\
-& 256 & GC & 69.04 & 67.064 & 0.268 \\
-& 256 & AGC & 69.06 & 67.298 & 0.502 \\
-& 256 & PGT & 69.97 & \textbf{67.814} & 1.018 \\
-& 256 & GC+PGT & 68.67 & 67.088 & 0.292 \\
-& 256 & AGC+PGT & 70.92 & \textbf{68.856} & 2.06 \\
+ResNet-18 (non-BN) & 1024 & Baseline & - & 67.19 & - & 64.93 & - \\
+\cmidrule{2-8}
+& 512 & Baseline & - & 68.02 & - & 66.552 & - \\
+& 512 & PGT & 0.25 & \textcolor{red}{\textbf{69.5}} & \textcolor{olive}{\textbf{+1.48}} &
+\textcolor{blue}{\textbf{67.236}} & \textcolor{olive}{\textbf{+0.684}} \\
+\cmidrule{2-8}
+& 256 & Baseline & - & 68.86 & - & 66.796 & - \\
+& 256 & GC & - & 69.04 & +0.18 & 67.064 & +0.268 \\
+& 256 & AGC & - & 69.06 & +0.2 & 67.298 & +0.502 \\
+& 256 & PGT & 0.25 & \textcolor{red}{\textbf{69.97}} & \textcolor{olive}{\textbf{+1.11}} &
+\textcolor{blue}{\textbf{67.814}} & \textcolor{olive}{\textbf{+1.018}} \\
+\cmidrule{2-8}
+& 256 & Baseline & - & 68.86 & - & 66.796 & - \\
+& 256 & GC+PGT & 0.25 & 68.67 & -0.19 & 67.088 & +0.292 \\
+& 256 & AGC+PGT & 0.25 & \textcolor{red}{\textbf{70.92}} & \textcolor{olive}{\textbf{+2.06}} &
+\textcolor{blue}{\textbf{68.856}} & \textcolor{olive}{\textbf{+2.06}} \\
 \end{tabular}
@@
 
-Experiments are carried out on non-Batch Normalized networks. We specifically utilize
-the ResNet-18 model and train it on ImageNet-1K which has 1.28 million images divided
-into 1000 classes. ImageNet-1K's test set consists of 50 images per class. Table
-\ref{tab:wobn_table} shows the findings.
-
-To begin, we use a batch size of 1024. This is only to illustrate that in unnormalized
-networks, accuracy suffers significantly when batch sizes are large. Additionally, the
-training process is often unstable, resulting in frequent layer dropouts (as seen in
-\ref{sec:Empi}) and sometimes infinite loss values. Thus, for all subsequent tests, we
-mostly use batch sizes of 512 and 256. We find $\alpha=0.25$ and $\alpha=0.05$ to be
-mostly optimal for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$
-also have good performance as well. We explore different values of $\alpha$ in section
-\ref{sec:Abla}. For ResNet-18, invoking PGT with $\alpha=0.25$ improves the baseline by
-$0.68\%$ when the batch size is 512, and by over $1\%$ when the batch size is 256. In
-comparison, the improvement obtained by GC and AGC is much less, at $0.27\%$ and
-$0.5\%$, respectively.
-
-On the metric of training accuracy, we get a $1.48\%$ increase in training accuracy at a
-batch size of 512 and an improvement of $1.11\%$ at a batch size of 256, leading us to
-infer that when PowerGrad Transform is used, the network's convergence optima is
-significantly superior. We also notice that the training versus testing gap increases
-with PGT as compared to the gap obtained at baseline. This allows us to infer that PGT
-increases the generalization gap between the training and the test set. The improvement
-in training and test accuracies are obtained not through regularization like it does in
-methods such as label smoothing. The improvement in both training and test accuracies
-are obtained as PGT enables the network to utilize its learning capacity better and
-arrive at a better optima at convergence.
-
-\subsection{Effect on Loss, Logits and other Metrics}
+We provide accuracy results for different batch sizes for the non-BN ResNet-18 trained
+on ImageNet-1K and also experiment with different methods (GC, AGC, PGT along with their
+combinations). Table \ref{tab:wobn_table} shows the findings. To begin, we use a batch
+size of 1024. This is only to illustrate that in unnormalized networks, accuracy suffers
+significantly when batch sizes are large. Additionally, the training process is often
+unstable at this batch size, resulting in frequent layer dropouts (as seen in
+\ref{sec:Empi}) and infinite loss values. Thus for all subsequent tests, we use batch
+sizes of 512 and 256. At a batch size of 512, invoking PGT with $\alpha=0.25$ improves
+the training accuracy baseline by $1.48\%$ and the test accuracy baseline by $0.684\%$,
+while at a batch size of 256, the improvement in training and test accuracies are
+$1.11\%$ and $1.018\%$ respectively. In comparison, the test accuracy improvements
+obtained by GC and AGC at batch size of 256 is much less, at $0.27\%$ and $0.5\%$,
+respectively.
+
+On the training accuracy front, since we get a significant boost ($1.48\%$ at batch size
+of 512 and $1.11\%$ at batch size of 256), it leads us to infer that when PowerGrad
+Transform is used, the network fits the training dataset more tightly and the
+convergence optima is significantly superior. We also notice that the training versus
+testing gap increases with PGT as compared to the gap at baseline even though both
+training and test performance is improved. This allows us to infer that PGT increases
+the generalization gap between the training and the test set, despite improvements in
+both training and test performance. Therefore it allows us to conclude that the
+improvements in training and test accuracies are obtained not through regularization
+like it does in methods such as label smoothing; rather the improvements are obtained as
+PGT enables the network to utilize its learning capacity better and arrive at a higher
+performing optima at convergence, something which is inaccessible through traditional
+gradient based training.
+
+We also experiment with combinations of GC/AGC and PGT. The advantage of combining GC
+with PGT is not very significant. When AGC and PGT are combined, we see a tremendous
+increase in test accuracy of over $2.06\%$ over the baseline. As seen in Fig.
+\ref{fig:agc_pgt}(c) and \ref{fig:agc_pgt}(f), activating AGC and PGT concurrently
+results in no filter dropouts in all layers while also improving the accuracy by a
+significant margin. Additionally, we provide comprehensive graphs of the norm profiles
+of the AGC + PGT method for all layers in the supplementary section.
+
+\subsection{Effect on Loss, Logits and other metrics}
 \label{sec:Effe}
@@
 
-As seen in Fig. \ref{fig:stats}a, the logit norm increases as $\alpha$ falls from $1$ to
-$0$. In the initial training phase, PGT causes the gradient in both the valid and wrong
-classes to rise as the network misclassifies majority of the examples. Due to the
+As seen in Fig. \ref{fig:stats}(a), the logit norm increases as $\alpha$ falls from $1$
+to $0$. In the initial training phase, PGT causes the gradient in both the valid and
+wrong classes to rise as the network misclassifies majority of the examples. Due to the
 exponential nature of the softmax function, the predicted probability distribution may
@@
 see that the final value of the loss rises as $\alpha$ drops, and that the logit norm
-and the final value of the loss are linearly correlated Fig. \ref{fig:stats}c. What is
-surprising is that greater accuracies may be acquired at larger loss values, at
-convergence. This is markedly different from the coupled gradient descent based training
-procedures. From these plots of logits and losses we find that with the inclusion of PGT
-in the training process, the model can have higher loss values at convergence compared
-to the loss when not using PGT. However, this decoupled gradient based training provides
-greater accuracies than the standard way of training even though the loss value is
-higher.
-
-\begin{figure}[!t]
-\centering
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r18}
-\caption{Plot of training and test accuracies (ResNet-18)}
-\end{subfigure}%
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r50}
-\caption{Plot of training and test accuracies (ResNet-50)}
-\end{subfigure}
-\caption{ Log-linear plots of training and test accuracies and comparison with baseline
-of batch-normalized variants: \textbf{(a)} ResNet-18 ($\alpha=0.25$), \textbf{(b)}
-ResNet-50 ($\alpha=0.3$). PGT (PowerGrad Transform) improves upon baseline in all cases.
-}
-\label{fig:metrics}
-\end{figure}
-
-ResNet-18 and ResNet-50 per epoch training and test accuracies with and without PGT are
-shown in Fig. \ref{fig:metrics}a,b. In all instances, we observe a gain. This means that
-PowerGrad Transform forces the network to learn better representations.
+and the final value of the loss are linearly correlated Fig. \ref{fig:stats}(c). What is
+surprising is that: \textit{greater accuracies may be acquired at larger loss values at
+convergence}. This is markedly different from the coupled gradient descent based
+training procedures. PGT enables the network to arrive at such optima where the loss
+values are high, but both training and test perfomance is better. From these plots of
+logits and losses we find that with the inclusion of PGT in the training process, the
+model can have access to such regions of the loss landscape that is otherwise
+inaccessible to traditional gradient based training procedures.
 
@@
 
-\begin{table}[t]
+\begin{table}[!t]
 \centering
@@
 hyperparameter column that controls the amount of PGT gradient modification injected at
-softmax, eq. (\ref{transformed_probabilities}). }
+softmax, Eq. \ref{transformed_probabilities}. }
 \label{tab:ablation_table}
-\vspace{5pt}
 \begin{tabular}{ccccccc}
-\textbf{Model} & \textbf{Scheduler} & \textbf{LS} & \textbf{Hparams} & \textbf{Train
-Acc} & \textbf{Test Acc} & \textbf{Gap} \\
+\multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} &
+\textbf{PowerGrad} & \textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap}} \\
+& & \textbf{Smoothing} & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} &
+\textbf{Accuracy (\%)} & \\
 \midrule
-ResNet-50 & Step &  & - & 78.99 & 75.97 & 3.02 \\
-& Step &  & 0.3 & 79.56 & 76.494 & 3.066 \\
-& Cosine &  & - & 79.18 & 76.56 & 2.62 \\
-& Cosine & $\checkmark$ & - & 78.81 & 76.698 & 2.112 \\
-& Cosine &  & 0.3 & 79.43 & 76.886 & 2.544 \\
-& Cosine & $\checkmark$ & 0.3 & 78.47 & 76.968 & 1.502 \\
-& Cosine &  & 0.05 & 79.68 & \textbf{77.216} & 2.464 \\
-& Cosine & $\checkmark$ & 0.05 & 77.69 & 76.39 & 1.3 \\
+ResNet-50 & Step & \xmark & \xmark & 78.99 & 75.97 & 3.02 \\
+& Step & \xmark & 0.3 & 79.56 & 76.494 & 3.066 \\
+\cmidrule{2-7}
+& Cosine & \xmark & \xmark & 79.18 & 76.56 & 2.62 \\
+& Cosine & 0.1 & \xmark & 78.81 & 76.698 & 2.112 \\
+\cmidrule{2-7}
+& Cosine & \xmark & 0.3 & 79.43 & 76.886 & 2.544 \\
+& Cosine & 0.1 & 0.3 & 78.47 & 76.968 & 1.502 \\
+\cmidrule{2-7}
+& Cosine & \xmark & 0.05 & 79.68 & 77.216 & 2.464 \\
+& Cosine & 0.1 & 0.05 & 77.69 & 76.39 & 1.3 \\
 \end{tabular}
@@
 
-We conduct an ablation study to investigate the consequences of PowerGrad Transform. We
-have previously demonstrated from the experiments depicted in section \ref{sec:Expe}
-that PowerGrad Transform causes the network to arrive at a better optima and fit a
-particular training dataset better than the baseline. We combine our approach with other
-regularization techniques such as label smoothing. The findings are shown in the Table
-\ref{tab:ablation_table}. We see that adding PowerGrad Transform increases the gap
-between the training and test accuracies, while introducing a label smoothing
-regularizer decreases the gap. We also investigate the two optimas obtained from
-ResNet-50's grid search. PGT with label smoothing improves the accuracy at $\alpha=0.3$
-but it hurts the accuracy at $\alpha=0.05$.
+We conduct an ablation study to investigate the effects of PowerGrad Transform for
+different values of the PGT hyperparameter($\alpha$). Grid plots are made available in
+the supplementary section. We have previously demonstrated from the experiments depicted
+in section \ref{sec:Expe}, that PowerGrad Transform causes the network to arrive at a
+better optima and fit the training dataset better than the baseline. In this ablation
+study, we use the ResNet-50 architecture and we combine our proposed method with other
+regularization techniques such as label smoothing and report our findings in table
+\ref{tab:ablation_table}. We also investigate the two optimal peaks of the
+hyperparameter ($\alpha$) obtained from ResNet-50's grid search ($\alpha=0.05$ and
+$\alpha=0.3$). First, we examine the effect of PGT on the step scheduler baseline in
+order to later compare it to the cosine scheduler baseline. PGT improves upon the step
+scheduler baseline (test set) by a substantial margin with $0.524\%$ ($76.494\%$ as
+opposed to $75.97\%$). Introducing the cosine scheduler yields a $0.59\%$ improvement
+($76.56\%$ vs. $75.97\%$) over the step scheduler. After introducing label smoothing,
+the test accuracy relative to the cosine scheduler baseline increases by only $0.138\%$
+(from $76.56\%$ to $76.698\%$). However, introducing PGT with $\alpha=0.3$ alone
+(without label smoothing) improves the cosine scheduler baseline by $0.326\%$
+($76.886\%$ vs. $76.56\%$). Combining PGT ($\alpha=0.3$) with label smoothing improves
+the performance on the test set further by $0.408\%$ (from $76.56\%$ to $76.968\%$) and
+reduces the generalization gap (from $2.54\%$ to $1.5\%$). However, the impact of
+combining PGT with label smoothing can vary depending on the value of the hyperparameter
+$\alpha$. With a PGT hyperparameter value of $\alpha=0.05$, we notice the greatest
+performance improvement, $1.246\%$ over the step scheduler test baseline and $0.656\%$
+over the cosine scheduler test baseline. Adding label smoothing to PGT ($\alpha=0.05$)
+hurts performance even though it reduces the generalization gap.
 
@@
 
-\newpage \subsubsection{Norm plots of weights (baseline)} \begin{figure}[h] \centering
-\includegraphics[width=0.55\textwidth]{baseline-w_norm} \caption{ Method: baseline. This
-is a plot depicting the evolution of the norm of each filter in each layer. Each subplot
+\newpage \subsection{Grid search for different values of $\alpha$}
+
+\newpage \subsection{Norm plots of weights (baseline (low batch size)) of non-BN
+ResNet-18}
+\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{baseline-w_norm}
+\caption{ Method: baseline, batch size = 256. This is a plot depicting the
+iteration-wise evolution of the norm of each filter in each layer. Each subplot
 corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
@@
 
-\newpage \subsubsection{Norm plots of weights (baseline (high batch size))}
+\newpage \subsection{Norm plots of weights (baseline (high batch size)) of non-BN
+ResNet-18}
 \begin{figure}[h] \centering
 \includegraphics[width=0.55\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
-baseline (high batch size). This is a plot depicting the evolution of the norm of each
-filter in each layer. Each subplot corresponds to a layer of the unnormalized ResNet-18.
-Above each subplot, the layer indices are shown. Each filter in each subplot is shown
-with a separate and randomly determined colour. } \end{figure}
-
-\newpage \subsubsection{Norm plots of weights (PGT)} \begin{figure}[h] \centering
-\includegraphics[width=0.55\textwidth]{pgt-w_norm} \caption{ Method: PGT. This is a plot
-depicting the evolution of the norm of each filter in each layer. Each subplot
+baseline (high batch size), batch size = 1024. This is a plot depicting the
+iteration-wise evolution of the norm of each filter in each layer. Each subplot
 corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
@@
 
-\newpage \subsubsection{Norm plots of weights (AGC + PGT)} \begin{figure}[h] \centering
-\includegraphics[width=0.55\textwidth]{agc_pgt-w_norm} \caption{ Method: AGC + PGT. This
-is a plot depicting the evolution of the norm of each filter in each layer. Each subplot
+\newpage \subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
+\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{pgt-w_norm}
+\caption{ Method: PGT, batch size = 256. This is a plot depicting the iteration-wise
+evolution of the norm of each filter in each layer. Each subplot corresponds to a layer
+of the unnormalized ResNet-18. Above each subplot, the layer indices are shown. Each
+filter in each subplot is shown with a separate and randomly determined colour. }
+\end{figure}
+
+\newpage \subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
+\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{agc_pgt-w_norm}
+\caption{ Method: AGC + PGT, batch size = 256. This is a plot depicting the
+iteration-wise evolution of the norm of each filter in each layer. Each subplot
 corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
@@
 
-\ \\ \ \\ \ \\ \ \\ \ \\
+ResNet-18 is composed of $20$ convolutional layers and one fully connected (FC) layer.
+At the input, it has a single convolutional layer with filters of size $7\times 7$.
+Three convolutional layers are placed in the downsampling skip connections. The
+remaining convolutional layers have several $3\times 3$ filters that are grouped
+together to create a block. Each basic block is connected by a skip connection. At the
+end of all convolutional layers, a global average pooling layer downsamples the features
+to a fixed-size feature vector and transfers it to the FC layer, which fits the features
+using regression and outputs logits. The logits $z_i$ produced by the FC layer are fed
+into a softmax layer, which transforms them to predicted probabilities $p_i$ using the
+equation $p_i=\frac {e^{z_i}}{\sum _j e^{z_j}}$.
 

### 2022-09-19 paper.tex
Mod7

@@
 ResNet-50 ($\alpha=0.3$). For better visualization of accuracies obtained in the final
-iterations, we plot the y-axis in log scale. PowerGrad Transform improves upon the
+iterations, we plot both axes in log scale. PowerGrad Transform improves upon the
 baseline in all cases. The corresponding accuracies are mentioned in table
@@
 \ref{fig:metrics}(a, b). For better visualization of the top end of the plots, we scale
-the y-axis in log scale. Our experiments with batch-normalized networks indicate that
+both axes in log scale. Our experiments with batch-normalized networks indicate that
 PowerGrad Transform speeds up training and allows the network to learn better
@@
 \midrule
-ResNet-18 (non-BN) & 1024 & Baseline & - & 67.19 & - & 64.93 & - \\
+ResNet-18 (non-BN) & 1024 & Baseline & - & 66.27 & - & 64.816 & - \\
+& 1024 & PGT & 0.92 & \textcolor{red}{\textbf{66.62}} &
+\textcolor{olive}{\textbf{+0.35}} & \textcolor{blue}{\textbf{65.498}} &
+\textcolor{olive}{\textbf{+0.682}} \\
 \cmidrule{2-8}
@@
 
-We provide accuracy results for different batch sizes for the non-BN ResNet-18 trained
-on ImageNet-1K and also experiment with different methods (GC, AGC, PGT along with their
-combinations). Table \ref{tab:wobn_table} shows the findings. To begin, we use a batch
-size of 1024. This is only to illustrate that in unnormalized networks, accuracy suffers
-significantly when batch sizes are large. Additionally, the training process is often
-unstable at this batch size, resulting in frequent layer dropouts (as seen in
-\ref{sec:Empi}) and infinite loss values. Thus for all subsequent tests, we use batch
-sizes of 512 and 256. At a batch size of 512, invoking PGT with $\alpha=0.25$ improves
-the training accuracy baseline by $1.48\%$ and the test accuracy baseline by $0.684\%$,
-while at a batch size of 256, the improvement in training and test accuracies are
-$1.11\%$ and $1.018\%$ respectively. In comparison, the test accuracy improvements
-obtained by GC and AGC at batch size of 256 is much less, at $0.27\%$ and $0.5\%$,
-respectively.
+We provide performance metrics on the test set of ImageNet-1K for different batch sizes
+for the non-BN ResNet-18 and also experiment with different methods (GC, AGC, PGT along
+with their combinations). Table \ref{tab:wobn_table} shows the findings. To begin, we
+use a batch size of 1024. The baseline performance for this batch size is drastically
+inferior to baselines for other batch sizes. PGT helps regain some of the lost
+performance by $0.682\%$ ($65.498\%$ vs. $64.816\%$). This illustrates that in
+unnormalized networks, accuracy suffers significantly when batch sizes are large.
+Additionally, the training process is often unstable at this batch size, resulting in
+filters and sometimes entire layers frequently zeroing out (as seen in \ref{sec:Empi}).
+At a batch size of 512, invoking PGT improves the training accuracy baseline by $1.48\%$
+and the test accuracy baseline by $0.684\%$, while at a batch size of 256, the
+improvement in training and test accuracies are $1.11\%$ and $1.018\%$ respectively. In
+comparison, the test accuracy improvements obtained by GC and AGC at batch size of 256
+is much less, at $0.27\%$ and $0.5\%$, respectively.
 
@@
 We conduct an ablation study to investigate the effects of PowerGrad Transform for
-different values of the PGT hyperparameter($\alpha$). Grid plots are made available in
-the supplementary section. We have previously demonstrated from the experiments depicted
-in section \ref{sec:Expe}, that PowerGrad Transform causes the network to arrive at a
+different values of the hyperparameter ($\alpha$). Grid plots are made available in the
+supplementary section. We have previously demonstrated from the experiments depicted in
+section \ref{sec:Expe}, that PowerGrad Transform causes the network to arrive at a
 better optima and fit the training dataset better than the baseline. In this ablation
@@
 
-\newpage
+\clearpage
 
@@
 
-\newpage \subsection{Grid search for different values of $\alpha$}
-
-\newpage \subsection{Norm plots of weights (baseline (low batch size)) of non-BN
+\clearpage
+
+\subsection{Grid search for different values of $\alpha$}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{r18_grid_search}
+\caption{ Grid search for various values of $\alpha$ for ResNet-18 architecture with the
+ImageNet dataset. Each datapoint is collected over a 50 epoch budget, with other
+hyperparameters kept the same as mentioned in section \ref{sec:bn}. }
+\label{fig:grid_searcha}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{r50_grid_search}
+\caption{ Grid search for various values of $\alpha$ for ResNet-50 architecture with the
+ImageNet dataset. Each datapoint is collected over a 50 epoch budget, with other
+hyperparameters kept the same as mentioned in section \ref{sec:bn}. }
+\label{fig:grid_searchb}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\columnwidth]{cifar10_grid_search}
+\caption{\\ResNet-20 CIFAR-10}
+\caption{ Grid search for various values of $\alpha$ for ResNet-20 architecture with the
+CIFAR-10 dataset. Each datapoint is collected over a 360 epoch budget. }
+\label{fig:grid_searchc}
+\end{figure}
+
+\clearpage
+
+\subsection{Norm plots of weights (baseline (low batch size)) of non-BN
 ResNet-18}
@@
 
-\newpage \subsection{Norm plots of weights (baseline (high batch size)) of non-BN
+\clearpage
+
+\subsection{Norm plots of weights (baseline (high batch size)) of non-BN
 ResNet-18}
@@
 
-\newpage \subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
+\clearpage
+
+\subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
 \begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{pgt-w_norm}
@@
 
-\newpage \subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
+\clearpage
+
+\subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
 \begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{agc_pgt-w_norm}
@@
 
-\newpage \subsection{ResNet-18 Layer Index Diagram}
+\clearpage
+
+\subsection{ResNet-18 Layer Index Diagram}
 

### 2022-09-19 paper.tex
Mod8

@@
 \centering
-\includegraphics[width=1.0\columnwidth]{PGT_Types}
+\includegraphics[width=\textwidth]{PGT_Types}
 \caption{ Different ways of altering gradients in the backward pass. Blue blocks denote
@@
 enhancement method that alters the gradients in the backward pass before the update step
-leading to significant boost in network's predictive performance.
+leading to accelerated training and a significant boost in the network's predictive
+performance.
 
@@
 \centering
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.96\columnwidth]{prob_transform_unordered}
+\begin{subfigure}{.5\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{prob_transform_unordered}
 \caption{Unordered distribution}
 \end{subfigure}%
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.96\columnwidth]{prob_transform_ordered}
+\begin{subfigure}{.5\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{prob_transform_ordered}
 \caption{Rank ordered distribution}
@@
 \centering
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r18}
+\begin{subfigure}{.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{acc_vs_epoch_r18}
 \caption{Plot of training and test accuracies (ResNet-18)}
 \end{subfigure}%
-\begin{subfigure}{.5\columnwidth}
-\centering
-\includegraphics[width=0.9\columnwidth]{acc_vs_epoch_r50}
+\begin{subfigure}{.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{acc_vs_epoch_r50}
 \caption{Plot of training and test accuracies (ResNet-50)}
+\end{subfigure}%
+\begin{subfigure}{.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{acc_vs_epoch_r50_60epochs}
+\caption{Plot of test accuracies (ResNet-50)}
 \end{subfigure}
@@
 baseline in all cases. The corresponding accuracies are mentioned in table
-\ref{tab:imagenet_table}. }
+\ref{tab:imagenet_table}. \textbf{(c)} Training speed comparison between PGT ($60$
+epochs) and baseline ($100$ epochs). They both converge to the same test accuracy
+($76.5\%$) on ImageNet-1K. PGT's accelerated training saves $40\%$ of the epoch budget.
+}
 \label{fig:metrics}
@@
 SE-ResNet-50\cite{hu2018squeeze}, we observe significant improvements; a $0.97\%$ boost
-in training accuracy and a $0.734\%$ increase in test accuracy. These improvements are
-obtained with the number of epochs being kept the same as that of the baseline model.
-Alternatively, practioners can choose to employ PGT and have the networks converge to
-accuracies similar to that of the baseline with a much lower epoch budget. We explore
-this in detail in section \ref{sec:speed}. Therefore, PGT also speeds up training. In
-our experiments with ResNet-50, we find an increase of $0.5\%$ and $0.656\%$ performance
-enhancement over the baseline while using the cosine scheduler. The corresponding
-improvement over the step scheduler baseline for ResNet-50 is $0.57\%$ and $0.524\%$
-respectively for training and testing. ResNet-101 sees a higher improvement in training
-fit, $0.81\%$ to be exact, while the improvement over the test set is $0.362\%$. Smaller
-networks such as SE-ResNet-18 and ResNet-18 sees accuracy boosts which are smaller but
-nevertheless positive. Therefore we conclude that with PGT, the consistent improvements
-in training accuracies across all cases is because the networks train better and arrive
-at better optimas during convergence. Per epoch training and test accuracy plots of
-ResNet-18 and ResNet-50 (both with and without PGT) are shown in Fig.
-\ref{fig:metrics}(a, b). For better visualization of the top end of the plots, we scale
-both axes in log scale. Our experiments with batch-normalized networks indicate that
-PowerGrad Transform speeds up training and allows the network to learn better
-representations\footnote{Reproducible code, training recipes for the experiments,
-pretrained checkpoints and training logs are provided at: \\
+in training accuracy and a $0.734\%$ increase in test accuracy. Therefore, PGT also
+speeds up training. In our experiments with ResNet-50, we find an increase of $0.5\%$
+and $0.656\%$ performance enhancement over the baseline while using the cosine
+scheduler. The corresponding improvement over the step scheduler baseline for ResNet-50
+is $0.57\%$ and $0.524\%$ respectively for training and testing. ResNet-101 sees a
+higher improvement in training fit, $0.81\%$ to be exact, while the improvement over the
+test set is $0.362\%$. Smaller networks such as SE-ResNet-18 and ResNet-18 sees accuracy
+boosts which are smaller but nevertheless positive. Therefore we conclude that with PGT,
+the consistent improvements in training accuracies across all cases is because the
+networks train better and arrive at better optimas during convergence. Per epoch
+training and test accuracy plots of ResNet-18 and ResNet-50 (both with and without PGT)
+are shown in Fig. \ref{fig:metrics}(a, b). For better visualization of the top end of
+the plots, we scale both axes in log scale. Our experiments with batch-normalized
+networks indicate that PowerGrad Transform speeds up training and allows the network to
+learn better representations\footnote{Reproducible code, training recipes for the
+experiments, pretrained checkpoints and training logs are provided at: \\
 \url{ https://github.com/bishshoy/power-grad-transform }. Code is adapted from
 \cite{rw2019timm}.}.
+
+These improvements are obtained with the number of epochs being kept the same as that of
+the baseline model. Alternatively, practitioners can apply PGT and have the networks
+converge to accuracy levels comparable to the baseline with a significantly reduced
+epoch budget. In Fig. \ref{fig:metrics}(c), we illustrate one such instance in which we
+train a ResNet-50 to its cosine scheduler baseline in just $60$ epochs with PGT, and
+match its baseline performance that is acquired in $100$ epochs. Because of the
+acceleration in training that is made possible by PGT, the total amount of training time
+can be decreased by $40\%$.
 
@@
 
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{baseline-w-layer-4-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{baseline-w-layer-4-2}
 \caption{Norm of filters of convolutional layer 11 (without PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{baseline-w-layer-7-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{baseline-w-layer-7-2}
 \caption{Norm of filters of the final convolutional layer (without PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{baseline-f-layer-22-1}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{baseline-f-layer-22-1}
 \caption{Norm of final layer's output features (without PGT)}
@@
 
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{pgt-w-layer-4-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{pgt-w-layer-4-2}
 \caption{Norm of filters of convolutional layer 11 (with PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{pgt-w-layer-7-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{pgt-w-layer-7-2}
 \caption{Norm of filters of the final convolutional layer (with PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{pgt-f-layer-22-1}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{pgt-f-layer-22-1}
 \caption{Norm of final layer's output features (with PGT)}
@@
 
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{baseline-w-layer-5-3}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{baseline-w-layer-5-3}
 \caption{Norm of filters of layer 15 (without AGC/PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{pgt-w-layer-5-3}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{pgt-w-layer-5-3}
 \caption{Norm of filters of layer 15 (with PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{agc_pgt-w-layer-5-3}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{agc_pgt-w-layer-5-3}
 \caption{Norm of filters of layer 15 (with AGC+PGT)}
@@
 
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{baseline-w-layer-1-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{baseline-w-layer-1-2}
 \caption{Norm of filters of layer 2 (without AGC/PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{pgt-w-layer-1-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{pgt-w-layer-1-2}
 \caption{Norm of filters of layer 2 (with PGT)}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\linewidth]{agc_pgt-w-layer-1-2}
+\begin{subfigure}[t]{0.33\textwidth}
+\includegraphics[width=\textwidth]{agc_pgt-w-layer-1-2}
 \caption{Norm of filters of layer 2 (with AGC+PGT)}
@@
 \centering
-\begin{subfigure}{.33\columnwidth}
-\centering
-\includegraphics[width=0.96\columnwidth]{logit_vs_alpha}
+\begin{subfigure}{0.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{logit_vs_alpha}
 \caption{\\Logit vs. $\alpha$}
 \end{subfigure}%
-\begin{subfigure}{.33\columnwidth}
-\centering
-\includegraphics[width=0.96\columnwidth]{loss_vs_alpha}
+\begin{subfigure}{0.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{loss_vs_alpha}
 \caption{\\Cross-entropy loss vs. $\alpha$}
 \end{subfigure}%
-\begin{subfigure}{.33\columnwidth}
-\centering
-\includegraphics[width=0.96\columnwidth]{logit_vs_loss}
+\begin{subfigure}{0.33\textwidth}
+\centering
+\includegraphics[width=0.98\textwidth]{logit_vs_loss}
 \caption{\\Logit-Loss correlation}
@@
 using the ImageNet dataset. PGT helps the network to improve its learning capabilities
-by locating a more optimum convergence point. In addition, we undertake an ablation
-research and compare its effects to those of regularization methods such label
-smoothing.
+by locating a more optimum convergence point and simultaneously speeds up training. In
+addition, we undertake an ablation research and compare its effects to those of
+regularization methods such label smoothing.
 
@@
 
-\begin{figure}[h] \hrule \lstinputlisting[language=Python]{code.py} \hrule
+\begin{figure}[ht] \hrule \lstinputlisting[language=Python]{code.py} \hrule
 \vspace{0.1cm} \caption{Python code of PowerGrad Transform based on PyTorch. }
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 ResNet-18}
-\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{baseline-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{baseline-w_norm}
 \caption{ Method: baseline, batch size = 256. This is a plot depicting the
@@
 ResNet-18}
-\begin{figure}[h] \centering
+\begin{figure}[ht] \centering
 \includegraphics[width=0.55\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
@@
 \subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
-\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{pgt-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{pgt-w_norm}
 \caption{ Method: PGT, batch size = 256. This is a plot depicting the iteration-wise
@@
 \subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
-\begin{figure}[h] \centering \includegraphics[width=0.55\textwidth]{agc_pgt-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{agc_pgt-w_norm}
 \caption{ Method: AGC + PGT, batch size = 256. This is a plot depicting the
@@
 
-\begin{figure}[h] \centering \includegraphics[width=0.96\columnwidth]{drawing} \caption{
+\begin{figure}[ht] \centering \includegraphics[width=0.96\columnwidth]{drawing} \caption{
 ResNet-18 architecture with layer indices of convolutional layers and the number of

### 2022-09-19 paper.tex
Mod9

@@
 
-% \usepackage{xcolor}
-% \usepackage{pagecolor}
-% \pagecolor{black}
-% \color{white}
+% \usepackage{xcolor} \usepackage{pagecolor} \pagecolor{black} \color{white}
 
@@
 iterations, we plot both axes in log scale. PowerGrad Transform improves upon the
-baseline in all cases. The corresponding accuracies are mentioned in table
+baseline in all cases. The corresponding accuracies are mentioned in Table
 \ref{tab:imagenet_table}. \textbf{(c)} Training speed comparison between PGT ($60$
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.734}}} \\
-& & PGT & 0.3 & \textcolor{red}{\textbf{82.47}} & & \textcolor{blue}{\textbf{77.952}} & \\
+& & PGT & 0.3 & \textcolor{red}{\textbf{82.47}} & & \textcolor{blue}{\textbf{77.952}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.656}}} \\
-& & PGT & 0.05 & \textcolor{red}{\textbf{79.68}} & & \textcolor{blue}{\textbf{77.216}} & \\
+& & PGT & 0.05 & \textcolor{red}{\textbf{79.68}} & & \textcolor{blue}{\textbf{77.216}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.524}}} \\
-& & PGT & 0.05 & \textcolor{red}{\textbf{79.56}} & & \textcolor{blue}{\textbf{76.494}} & \\
+& & PGT & 0.05 & \textcolor{red}{\textbf{79.56}} & & \textcolor{blue}{\textbf{76.494}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.362}}} \\
-& & PGT & 0.3 & \textcolor{red}{\textbf{83.1}} & & \textcolor{blue}{\textbf{78.258}} & \\
+& & PGT & 0.3 & \textcolor{red}{\textbf{83.1}} & & \textcolor{blue}{\textbf{78.258}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.346}}} \\
-& & PGT & 0.25 & \textcolor{red}{\textbf{71.6}} & & \textcolor{blue}{\textbf{71.436}} & \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{71.6}} & & \textcolor{blue}{\textbf{71.436}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.14}}} \\
-& & PGT & 0.25 & \textcolor{red}{\textbf{70.3}} & & \textcolor{blue}{\textbf{69.844}} & \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{70.3}} & & \textcolor{blue}{\textbf{69.844}} &
+\\
 \midrule
@@
 \textcolor{olive}{\multirow{2}{*}{\textbf{+0.09}}} \\
-& & PGT & 0.25 & \textcolor{red}{\textbf{70.53}} & & \textcolor{blue}{\textbf{70.298}} & \\
+& & PGT & 0.25 & \textcolor{red}{\textbf{70.53}} & & \textcolor{blue}{\textbf{70.298}} &
+\\
 \end{tabular}
@@
 for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$ also have good
-performance as well. The experimental results are shown in table
+performance as well. The experimental results are shown in Table
 \ref{tab:imagenet_table}, and we mention the value of the PGT hyperparmeter ($\alpha$)
@@
 SE-ResNet-50\cite{hu2018squeeze}, we observe significant improvements; a $0.97\%$ boost
-in training accuracy and a $0.734\%$ increase in test accuracy. Therefore, PGT also
-speeds up training. In our experiments with ResNet-50, we find an increase of $0.5\%$
-and $0.656\%$ performance enhancement over the baseline while using the cosine
-scheduler. The corresponding improvement over the step scheduler baseline for ResNet-50
-is $0.57\%$ and $0.524\%$ respectively for training and testing. ResNet-101 sees a
-higher improvement in training fit, $0.81\%$ to be exact, while the improvement over the
-test set is $0.362\%$. Smaller networks such as SE-ResNet-18 and ResNet-18 sees accuracy
-boosts which are smaller but nevertheless positive. Therefore we conclude that with PGT,
-the consistent improvements in training accuracies across all cases is because the
-networks train better and arrive at better optimas during convergence. Per epoch
-training and test accuracy plots of ResNet-18 and ResNet-50 (both with and without PGT)
-are shown in Fig. \ref{fig:metrics}(a, b). For better visualization of the top end of
-the plots, we scale both axes in log scale. Our experiments with batch-normalized
-networks indicate that PowerGrad Transform speeds up training and allows the network to
-learn better representations\footnote{Reproducible code, training recipes for the
-experiments, pretrained checkpoints and training logs are provided at: \\
+in training accuracy and a $0.734\%$ increase in test accuracy. In our experiments with
+ResNet-50, we find an increase of $0.5\%$ and $0.656\%$ performance enhancement over the
+cosine scheduler baseline for training and testing respectively. The corresponding
+improvement over the step scheduler baseline for ResNet-50 is $0.57\%$ (training) and
+$0.524\%$ (testing). ResNet-101 sees a higher improvement in training fit, $0.81\%$ to
+be exact, while the improvement over the test set is $0.362\%$. Smaller networks such as
+SE-ResNet-18 and ResNet-18 sees accuracy boosts which are smaller but nevertheless
+positive. Therefore we conclude that with PGT, the consistent improvements in training
+accuracies across all cases is because the networks train better and arrive at better
+optimas during convergence. Per epoch training and test accuracy plots of ResNet-18 and
+ResNet-50 (both with and without PGT) are shown in Fig. \ref{fig:metrics}(a, b). For
+better visualization of the top end of the plots, we scale both axes in log scale.
+Because both training and test accuracies improve, it indicates that PowerGrad Transform
+allows the network to learn better representations\footnote{Reproducible code, training
+recipes for the experiments, pretrained checkpoints and training logs are provided at:
+\\
 \url{ https://github.com/bishshoy/power-grad-transform }. Code is adapted from
@@
 train a ResNet-50 to its cosine scheduler baseline in just $60$ epochs with PGT, and
-match its baseline performance that is acquired in $100$ epochs. Because of the
-acceleration in training that is made possible by PGT, the total amount of training time
-can be decreased by $40\%$.
+match its baseline performance that is acquired in $100$ epochs. This hints to the
+observation that PGT not only makes the network perform better but also speeds up
+training. Because of the acceleration in training that is made possible by PGT, the
+total amount of training time can be decreased by $40\%$.
 
@@
 AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
-it also mitigates the zeroing out phenomenon in most layers. In
-Fig.\ref{fig:agc_pgt}(a), we see a plot of the $15^\textit{th}$ layer of the non-BN
+it also mitigates the zeroing out phenomenon in most layers. In Fig.
+\ref{fig:agc_pgt}(a), we see a plot of the $15^\textit{th}$ layer of the non-BN
 ResNet-18 trained without PGT. As we can see that there has been quite a few filter
@@
 \ref{fig:agc_pgt}(e)] than in the baseline [Fig. \ref{fig:agc_pgt}(d)]. Despite that,
-PGT achieves a higher overall accuracy (table \ref{tab:wobn_table}). This necessitates
+PGT achieves a higher overall accuracy (Table \ref{tab:wobn_table}). This necessitates
 the simultaneous use of AGC and PGT, for the non-BN ResNet-18.
@@
 & 512 & Baseline & - & 68.02 & - & 66.552 & - \\
-& 512 & PGT & 0.25 & \textcolor{red}{\textbf{69.5}} & \textcolor{olive}{\textbf{+1.48}} &
-\textcolor{blue}{\textbf{67.236}} & \textcolor{olive}{\textbf{+0.684}} \\
+& 512 & PGT & 0.25 & \textcolor{red}{\textbf{69.5}} & \textcolor{olive}{\textbf{+1.48}}
+& \textcolor{blue}{\textbf{67.236}} & \textcolor{olive}{\textbf{+0.684}} \\
 \cmidrule{2-8}
@@
 & 256 & AGC & - & 69.06 & +0.2 & 67.298 & +0.502 \\
-& 256 & PGT & 0.25 & \textcolor{red}{\textbf{69.97}} & \textcolor{olive}{\textbf{+1.11}} &
-\textcolor{blue}{\textbf{67.814}} & \textcolor{olive}{\textbf{+1.018}} \\
+& 256 & PGT & 0.25 & \textcolor{red}{\textbf{69.97}} & \textcolor{olive}{\textbf{+1.11}}
+& \textcolor{blue}{\textbf{67.814}} & \textcolor{olive}{\textbf{+1.018}} \\
 \cmidrule{2-8}
@@
 & 256 & GC+PGT & 0.25 & 68.67 & -0.19 & 67.088 & +0.292 \\
-& 256 & AGC+PGT & 0.25 & \textcolor{red}{\textbf{70.92}} & \textcolor{olive}{\textbf{+2.06}} &
-\textcolor{blue}{\textbf{68.856}} & \textcolor{olive}{\textbf{+2.06}} \\
+& 256 & AGC+PGT & 0.25 & \textcolor{red}{\textbf{70.92}} &
+\textcolor{olive}{\textbf{+2.06}} & \textcolor{blue}{\textbf{68.856}} &
+\textcolor{olive}{\textbf{+2.06}} \\
 \end{tabular}
@@
 Additionally, the training process is often unstable at this batch size, resulting in
-filters and sometimes entire layers frequently zeroing out (as seen in \ref{sec:Empi}).
-At a batch size of 512, invoking PGT improves the training accuracy baseline by $1.48\%$
-and the test accuracy baseline by $0.684\%$, while at a batch size of 256, the
-improvement in training and test accuracies are $1.11\%$ and $1.018\%$ respectively. In
-comparison, the test accuracy improvements obtained by GC and AGC at batch size of 256
-is much less, at $0.27\%$ and $0.5\%$, respectively.
+filters and sometimes entire layers frequently zeroing out (as seen in section
+\ref{sec:Empi}). At a batch size of 512, invoking PGT improves the training accuracy
+baseline by $1.48\%$ and the test accuracy baseline by $0.684\%$, while at a batch size
+of 256, the improvement in training and test accuracies are $1.11\%$ and $1.018\%$
+respectively. In comparison, the test accuracy improvements obtained by GC and AGC at
+batch size of 256 is much less, at $0.27\%$ and $0.5\%$, respectively.
 
@@
 testing gap increases with PGT as compared to the gap at baseline even though both
-training and test performance is improved. This allows us to infer that PGT increases
-the generalization gap between the training and the test set, despite improvements in
-both training and test performance. Therefore it allows us to conclude that the
+training and test accuracies improve. Therefore it allows us to conclude that the
 improvements in training and test accuracies are obtained not through regularization
@@
 and the final value of the loss are linearly correlated Fig. \ref{fig:stats}(c). What is
-surprising is that: \textit{greater accuracies may be acquired at larger loss values at
-convergence}. This is markedly different from the coupled gradient descent based
-training procedures. PGT enables the network to arrive at such optima where the loss
-values are high, but both training and test perfomance is better. From these plots of
-logits and losses we find that with the inclusion of PGT in the training process, the
-model can have access to such regions of the loss landscape that is otherwise
-inaccessible to traditional gradient based training procedures.
+surprising is that: \textit{it is possible to achieve higher accuracies even though the
+loss values are larger during convergence}. In our studies with ResNet-50, for instance,
+if PGT is not invoked, the corresponding test accuracy is $76.56\%$ (Table
+\ref{tab:imagenet_table}) with a cross-entropy loss value of $0.852$. However, if PGT is
+activated with $\alpha=0.05$, the corresponding test accuracy is $77.216\%$ (Table
+\ref{tab:imagenet_table}) with a loss value of $2.360$. This is markedly different from
+the coupled gradient descent based training procedures. PGT enables the network to
+arrive at such optima where the loss values are high, but both training and test
+perfomance is better. From these plots of logits and losses we find that with the
+inclusion of PGT in the training process, the model can have access to such regions of
+the loss landscape that is otherwise inaccessible to traditional gradient based training
+procedures.
 
@@
 \label{tab:ablation_table}
-\begin{tabular}{ccccccc}
-\multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} &
-\textbf{PowerGrad} & \textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap}} \\
-& & \textbf{Smoothing} & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} &
+\begin{tabular}{cccccccc}
+\multirow{2}{*}{\textbf{\#Row}} & \multirow{2}{*}{\textbf{Model}} &
+\multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} & \textbf{PowerGrad} &
+\textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap (\%)}} \\
+& & & \textbf{Smoothing} & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} &
 \textbf{Accuracy (\%)} & \\
 \midrule
-ResNet-50 & Step & \xmark & \xmark & 78.99 & 75.97 & 3.02 \\
-& Step & \xmark & 0.3 & 79.56 & 76.494 & 3.066 \\
-\cmidrule{2-7}
-& Cosine & \xmark & \xmark & 79.18 & 76.56 & 2.62 \\
-& Cosine & 0.1 & \xmark & 78.81 & 76.698 & 2.112 \\
-\cmidrule{2-7}
-& Cosine & \xmark & 0.3 & 79.43 & 76.886 & 2.544 \\
-& Cosine & 0.1 & 0.3 & 78.47 & 76.968 & 1.502 \\
-\cmidrule{2-7}
-& Cosine & \xmark & 0.05 & 79.68 & 77.216 & 2.464 \\
-& Cosine & 0.1 & 0.05 & 77.69 & 76.39 & 1.3 \\
+1. & ResNet-50 & Step & \xmark & \xmark & 78.99 & 75.97 & 3.02 \\
+2. & ResNet-50 & Step & \xmark & 0.3 & 79.56 & 76.494 & 3.066 \\
+\cmidrule{2-8}
+3. & ResNet-50 & Cosine & \xmark & \xmark & 79.18 & 76.56 & 2.62 \\
+4. & ResNet-50 & Cosine & 0.1 & \xmark & 78.81 & 76.698 & 2.112 \\
+\cmidrule{2-8}
+5. & ResNet-50 & Cosine & \xmark & 0.3 & 79.43 & 76.886 & 2.544 \\
+6. & ResNet-50 & Cosine & 0.1 & 0.3 & 78.47 & 76.968 & 1.502 \\
+\cmidrule{2-8}
+7. & ResNet-50 & Cosine & \xmark & 0.05 & 79.68 & 77.216 & 2.464 \\
+8. & ResNet-50 & Cosine & 0.1 & 0.05 & 77.69 & 76.39 & 1.3 \\
 \end{tabular}
@@
 study, we use the ResNet-50 architecture and we combine our proposed method with other
-regularization techniques such as label smoothing and report our findings in table
+regularization techniques such as label smoothing and report our findings in Table
 \ref{tab:ablation_table}. We also investigate the two optimal peaks of the
@@
 $\alpha=0.3$). First, we examine the effect of PGT on the step scheduler baseline in
-order to later compare it to the cosine scheduler baseline. PGT improves upon the step
+order to later compare it to the cosine scheduler baseline. \textbf{Row-1)} We begin
+with the step scheduler baseline ($75.97\%$). \textbf{Row-2)} PGT improves upon the step
 scheduler baseline (test set) by a substantial margin with $0.524\%$ ($76.494\%$ as
-opposed to $75.97\%$). Introducing the cosine scheduler yields a $0.59\%$ improvement
-($76.56\%$ vs. $75.97\%$) over the step scheduler. After introducing label smoothing,
-the test accuracy relative to the cosine scheduler baseline increases by only $0.138\%$
-(from $76.56\%$ to $76.698\%$). However, introducing PGT with $\alpha=0.3$ alone
-(without label smoothing) improves the cosine scheduler baseline by $0.326\%$
-($76.886\%$ vs. $76.56\%$). Combining PGT ($\alpha=0.3$) with label smoothing improves
-the performance on the test set further by $0.408\%$ (from $76.56\%$ to $76.968\%$) and
-reduces the generalization gap (from $2.54\%$ to $1.5\%$). However, the impact of
-combining PGT with label smoothing can vary depending on the value of the hyperparameter
-$\alpha$. With a PGT hyperparameter value of $\alpha=0.05$, we notice the greatest
+opposed to $75.97\%$). \textbf{Row-3)} Introducing the cosine scheduler yields a
+$0.59\%$ improvement ($76.56\%$ vs. $75.97\%$) over the step scheduler. \textbf{Row-4)}
+After introducing label smoothing, the test accuracy relative to the cosine scheduler
+baseline increases by only $0.138\%$ (from $76.56\%$ to $76.698\%$). \textbf{Row-5)}
+However, introducing PGT with $\alpha=0.3$ alone (without label smoothing) improves the
+cosine scheduler baseline by $0.326\%$ ($76.886\%$ vs. $76.56\%$). \textbf{Row-6)}
+Combining PGT ($\alpha=0.3$) with label smoothing improves the performance on the test
+set further by $0.408\%$ (from $76.56\%$ to $76.968\%$) and reduces the generalization
+gap (from $2.54\%$ to $1.5\%$). However, the impact of combining PGT with label
+smoothing can vary depending on the value of the hyperparameter ($\alpha$).
+\textbf{Row-7)} With a PGT hyperparameter value of $\alpha=0.05$, we notice the greatest
 performance improvement, $1.246\%$ over the step scheduler test baseline and $0.656\%$
-over the cosine scheduler test baseline. Adding label smoothing to PGT ($\alpha=0.05$)
-hurts performance even though it reduces the generalization gap.
+over the cosine scheduler test baseline. \textbf{Row-8)} Adding label smoothing to PGT
+($\alpha=0.05$) hurts performance even though it reduces the generalization gap.
 
@@
 
+There is no effect of PGT on small datasets such as CIFAR-10. In these scenarios,
+networks almost always achieve $100\%$ training accuracy, leaving no room for PGT to
+assist.
+
 \clearpage
 
-\subsection{Norm plots of weights (baseline (low batch size)) of non-BN
-ResNet-18}
+\subsection{Norm plots of weights (baseline (low batch size)) of non-BN ResNet-18}
 \begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{baseline-w_norm}
@@
 
-\subsection{Norm plots of weights (baseline (high batch size)) of non-BN
-ResNet-18}
+\subsection{Norm plots of weights (baseline (high batch size)) of non-BN ResNet-18}
 \begin{figure}[ht] \centering
@@
 
-\begin{figure}[ht] \centering \includegraphics[width=0.96\columnwidth]{drawing} \caption{
-ResNet-18 architecture with layer indices of convolutional layers and the number of
-filters in each layer. } \label{fig:resnet18} \end{figure}
+\begin{figure}[ht] \centering \includegraphics[width=0.96\columnwidth]{drawing}
+\caption{ ResNet-18 architecture with layer indices of convolutional layers and the
+number of filters in each layer. } \label{fig:resnet18} \end{figure}
 

### 2022-11-14 paper.tex
Mod10
Mod11
Title change

@@
 
-\title{Altering Backward Pass Gradients improves Convergence}
+\title{Altering Backward Pass Gradients to improve Convergence}
 
@@
 training and test sets. PowerGrad Transform is easy to integrate into existing training
-routines, requiring just a few lines of code. PowerGrad Transform makes it possible for
-the network to better fit the training data. With decoupled training, PowerGrad
-Transform improves baseline accuracies for ResNet-50 by 0.73\%, for SE-ResNet-50 by
-0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network on the ImageNet
-classification task.
+routines, requiring just a few lines of code. PowerGrad Transform accelerates training
+and makes it possible for the network to better fit the training data. With decoupled
+training, PowerGrad Transform improves baseline accuracies for ResNet-50 by 0.73\%, for
+SE-ResNet-50 by 0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network
+on the ImageNet classification task.
 \end{abstract}
@@
 
+\clearpage
+
 \begin{figure}[ht]

### 2022-11-14 0100.tex
Added PAKDD version.

--- 2022-11-14 paper.tex
@@
-\documentclass[times,sort&compress]{elsarticle}
-\usepackage{ycviu}
+\documentclass[runningheads]{llncs}
+\usepackage[T1]{fontenc}
 \usepackage{framed}
@@
 \usepackage{xcolor}
-\usepackage{hyperref}
 \usepackage{subcaption}
@@
 \usepackage{booktabs}
-\usepackage{amsthm}
+% \usepackage{amsthm}
 \usepackage{pifont}
 \usepackage{graphicx}
-
-% \usepackage{xcolor} \usepackage{pagecolor} \pagecolor{black} \color{white}
-
+\usepackage{hyperref}
+\usepackage{color}
 \graphicspath{{images/}}
-
+\renewcommand\UrlFont{\color{blue}\rmfamily}
 \newcommand{\norm}[1]{\left\lVert#1\right\rVert}
@@
 \newcommand{\cmark}{\ding{51}} \newcommand{\xmark}{\ding{55}}
-
 \definecolor{newcolor}{rgb}{.8,.349,.1}
@@
 
-\journal{Computer Vision and Image Understanding}
-
 \begin{document}
-
-\thispagestyle{empty}
-
-\begin{table*}[!th]
-
-\begin{minipage}{.9\textwidth}
-\baselineskip12pt
-\ifpreprint
-  \vspace*{1pc}
-\else
-  \vspace*{-6pc}
-\fi
-
-\noindent {\LARGE\itshape Computer Vision and Image Understanding}
-\vskip6pt
-
-\noindent {\Large\bfseries Authorship Confirmation}
-
-\vskip1pc
-
-{\bf Please save a copy of this file, complete and upload as the ``Confirmation of
-Authorship'' file.}
-
-\vskip1pc
-
-As corresponding author I, \underline{\hphantom{\hspace*{7cm}}}, hereby confirm on
-behalf of all authors that:
-
-\vskip1pc
-
-\begin{enumerate}
-\itemsep=3pt
-\item This manuscript, or a large part of it, \underline {has not been published,  was
-not, and is not being submitted to} any other journal.
-
-\item If \underline {presented} at or \underline {submitted} to or \underline {published
-}at a conference(s), the conference(s) is (are) identified and  substantial \underline
-{justification for re-publication} is presented  below. A \underline {copy of conference
-paper(s) }is(are) uploaded with the  manuscript.
-
-\item If the manuscript appears as a preprint anywhere on the web, e.g. arXiv,  etc., it
-is identified below. The \underline {preprint should include a  statement that the paper
-is under consideration at Computer Vision and Image Understanding}.
-
-\item All text and graphics, except for those marked with sources, are \underline
-{original works} of the authors, and all necessary permissions for  publication were
-secured prior to submission of the manuscript.
-
-\item All authors each made a significant contribution to the research reported  and
-have \underline {read} and \underline {approved} the submitted  manuscript. 
-\end{enumerate}
-
-Signature\underline{\hphantom{\hspace*{7cm}}} Date\underline{\hphantom{\hspace*{4cm}}}
-\vskip1pc
-
-\rule{\textwidth}{2pt}
-\vskip1pc
-
-{\bf List any pre-prints:} \vskip5pc
-
-\rule{\textwidth}{2pt}
-\vskip1pc
-
-{\bf Relevant Conference publication(s) (submitted, accepted, or published):} \vskip5pc
-
-{\bf Justification for re-publication:}
-
-\end{minipage}
-\end{table*}
-
-\clearpage
-\thispagestyle{empty}
-\ifpreprint
-  \vspace*{-1pc}
-\fi
-
-\begin{table*}[!t]
-\ifpreprint\else\vspace*{-15pc}\fi
-
-\section*{Research Highlights (Required)}
-
-To create your highlights, please type the highlights against each \verb+\item+ command.
-
-\vskip1pc
-
-\fboxsep=6pt \fbox{
-\begin{minipage}{.95\textwidth}
-It should be short collection of bullet points that convey the core findings of the
-article. It should  include 3 to 5 bullet points (maximum 85 characters, including
-spaces, per bullet point.)  
-\vskip1pc
-\begin{itemize}
-
-\item Altering gradients in the backward pass can significantly alter a CNN's
-performance.
-
-\item PowerGrad Transform is proposed that enhances the backward pass gradients.
-
-\item Theoretical analysis of the impact of PowerGrad transformation is conducted.
-
-\item Efficacy of PowerGrad Transform is demonstrated through experiments.
-
-\end{itemize}
-\vskip1pc
-\end{minipage}
-}
-
-\end{table*}
-
-\clearpage
-
-\ifpreprint
-  \setcounter{page}{1}
-\else
-  \setcounter{page}{1}
-\fi
-
-\begin{frontmatter}
-
 \title{Altering Backward Pass Gradients to improve Convergence}
-
-\author[1]{Bishshoy Das\corref{cor1}}
-\ead{bishshoy.das@ee.iitd.ac.in}
-
-\author[1]{Milton Mondal\corref{cor2}}
-\ead{milton.mondal@ee.iitd.ac.in}
-
-\author[1]{\\ Brejesh Lall}
-\ead{brejesh@ee.iitd.ac.in}
-
-\author[1]{Shiv Dutt Joshi}
-\ead{sdjoshi@ee.iitd.ac.in}
-
-\author[1]{Sumantra Dutta Roy}
-\ead{sumantra@ee.iitd.ac.in}
-
-\cortext[cor1]{Corresponding author}
-\cortext[cor2]{Equal contribution}
-
-\address[1]{Electrical Engineering Department, Indian Institute of Technology Delhi, Hauz Khas, New Delhi - 110016, India}
-
-\received{1 May 2013}
-\finalform{10 May 2013}
-\accepted{13 May 2013}
-\availableonline{15 May 2013}
-\communicated{S. Sarkar}
+\author{Bishshoy Das\inst{1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{1}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Brejesh Lall\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Shiv Dutt Joshi\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Sumantra Dutta Roy\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}}
+\authorrunning{B. Das et al.}
+\institute{Indian Institute of Technology Delhi, New Delhi - 110016, India \\ \email{\{bishshoy.das, milton.mondal, brejesh, sdjoshi, sumantra\}@ee.iitd.ac.in}}
+\maketitle
 
@@
 where we alter the gradients in the backward pass. We propose a simple yet powerful
-method called PowerGrad Transform, that alters the gradients before the weight update in
-the backward pass and significantly enhances the predictive performance of the neural
-network. PowerGrad Transform trains the network to arrive at a better optima at
-convergence. It is computationally extremely efficient, virtually adding no additional
-cost to either memory or compute, but results in improved final accuracies on both the
-training and test sets. PowerGrad Transform is easy to integrate into existing training
-routines, requiring just a few lines of code. PowerGrad Transform accelerates training
-and makes it possible for the network to better fit the training data. With decoupled
-training, PowerGrad Transform improves baseline accuracies for ResNet-50 by 0.73\%, for
-SE-ResNet-50 by 0.66\% and by more than 1.0\% for the non-normalized ResNet-18 network
-on the ImageNet classification task.
+method called PowerGrad Transform (PGT), that alters the gradients before the weight
+update in the backward pass and significantly enhances the predictive performance of the
+neural network. PGT trains the network to arrive at a better optima at convergence. It
+is computationally extremely efficient, virtually adding no additional cost to either
+memory or compute, but results in improved final accuracies on both the training and
+test sets. PGT is easy to integrate into existing training routines, requiring just a
+few lines of code. PGT accelerates training and makes it possible for the network to
+better fit the training data. With decoupled training, PGT improves baseline accuracies
+for ResNet-50 by 0.73\%, for SE-ResNet-50 by 0.66\% and by more than 1.0\% for the
+non-normalized ResNet-18 network on the ImageNet classification task.
+\keywords{backpropagation \and gradients \and neural networks \and softmax \and clipping.}
 \end{abstract}
-
-\begin{keyword}
-\KWD backpropagation \sep gradients \sep neural networks \sep softmax \sep clipping
-\end{keyword}
-
-\end{frontmatter}
 
@@
 \begin{equation}
-\hspace{4cm}
+% \hspace{4cm}
 W_i^{t+1} = W_i^t - \lambda\ \nabla_{W_i}(L)
@@
 \begin{equation}
-\hspace{4cm}
+% \hspace{4cm}
 W_i^{t+1} = W_i^t - \lambda\ f(\nabla_{W_i}(L))
@@
 \begin{align}
-\hspace{4cm}
+% \hspace{4cm}
 W_D^{t+1} &= W_D^t - \lambda\ f(\nabla_{W_D}(L)) & \\
-W_i^{t+1} &= W_i^t - \lambda\ \nabla_{W_i}(L)^{*} &\hspace{-3cm} i=D-1,\ldots,1
+W_i^{t+1} &= W_i^t - \lambda\ \nabla_{W_i}(L)^{*} &\hspace{-1cm} i=D-1,\ldots,1
 \end{align}
@@
 during backward propagation through the network and updating the weights using the
-clipped gradients \cite{zhang2019gradient, smith2020generalization}. By rescaling the
+clipped gradients \cite{zhang2019gradient} \cite{smith2020generalization}. By rescaling the
 gradients, the weight updates are likewise rescaled, significantly reducing the risk of
@@
 labels generated by the teacher network. Variants of knowledge distillation include
-self-distillation \cite{zhang2019your, wang2021memory}, identical student network
+self-distillation \cite{zhang2019your} \cite{wang2021memory}, identical student network
 distillation \cite{furlanello2018born}, channel distillation \cite{ge2019distilling},
 regularizing wrong predictions \cite{yun2020regularizing} and representation or
-embedding based knowledge distillation \cite{aguilar2020knowledge, yao2018deep,
-passalis2018unsupervised}. Distillation is applied to various tasks such as model
+embedding based knowledge distillation \cite{aguilar2020knowledge} \cite{yao2018deep}
+\cite{passalis2018unsupervised}. Distillation is applied to various tasks such as model
 compression \cite{wang2020real}, face recognition \cite{ge2018low}, enhancing ensemble
 performance \cite{zhang2018deep}, interpreting neural networks \cite{liu2018improving}.
-Efficacy of distillation is extensivly studied in \cite{cho2019efficacy,
-yuan2020revisiting}. Applications of logit suppression include: deep metric learning
-\cite{zhai2018classification}, non-parametric neighborhood component analysis
+Efficacy of distillation is extensivly studied in \cite{cho2019efficacy}
+\cite{yuan2020revisiting}. Applications of logit suppression include: deep metric
+learning \cite{zhai2018classification}, non-parametric neighborhood component analysis
 \cite{wu2018improving}, sequence-to-sequence modelling \cite{chorowski2016towards},
 reinforcement learning \cite{he2018determining} and face recognition
-\cite{liu2017sphereface, wang2018cosface}.
+\cite{liu2017sphereface} \cite{wang2018cosface}.
 
@@
 
-\begin{proof}[\unskip\nopunct] \begin{alignat}{2} &\ \ \frac{p_i^{\alpha }}{\sum
+\begin{proof} \begin{alignat}{2} &\ \ \frac{p_i^{\alpha }}{\sum
 _{j=1}^C p_j^{\alpha}} < p_i \\ \Rightarrow &\ p_i^{\alpha -1}<\sum _{j=1}^C p_j^{\alpha
@@
 
-\begin{proof}[\unskip\nopunct] \begin{equation} \frac{\partial}{\partial \alpha
+\begin{proof} \begin{equation} \frac{\partial}{\partial \alpha
 }\left(\sum _{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq \ \ 0
@@
 \multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Scheduler}} &
-\multirow{2}{*}{\textbf{Method}} & \textbf{PowerGrad} & \textbf{\hspace{-0.65cm} Train}
+\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{\hspace{-0.65cm} Train}
 & \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
-& & & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} & \textbf{Diff (\%)} &
-\textbf{Accuracy (\%)} & \textbf{Diff (\%)} \\
+& & & \textbf{($\alpha$)} & \textbf{Acc. (\%)} & \textbf{Diff (\%)} &
+\textbf{Acc. (\%)} & \textbf{Diff (\%)} \\
 \midrule
@@
 recipes for the experiments, pretrained checkpoints and training logs are provided at:
-\\
 \url{ https://github.com/bishshoy/power-grad-transform }. Code is adapted from
@@
 
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{baseline-w-layer-4-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{baseline-w-layer-7-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{baseline-f-layer-22-1}
@@
 
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{pgt-w-layer-4-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{pgt-w-layer-7-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{pgt-f-layer-22-1}
@@
 
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{baseline-w-layer-5-3}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{pgt-w-layer-5-3}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{agc_pgt-w-layer-5-3}
@@
 
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{baseline-w-layer-1-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{pgt-w-layer-1-2}
@@
 \end{subfigure}
-\begin{subfigure}[t]{0.33\textwidth}
+\begin{subfigure}[t]{0.32\textwidth}
 \includegraphics[width=\textwidth]{agc_pgt-w-layer-1-2}
@@
 \multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Batch Size}} &
-\multirow{2}{*}{\textbf{Method}} & \textbf{PowerGrad} & \textbf{\hspace{-0.65cm} Train}
+\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{\hspace{-0.65cm} Train}
 & \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
-& & & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} & \textbf{Diff (\%)} &
-\textbf{Accuracy (\%)} & \textbf{Diff (\%)} \\
+& & & \textbf{($\alpha$)} & \textbf{Acc. (\%)} & \textbf{Diff (\%)} &
+\textbf{Acc. (\%)} & \textbf{Diff (\%)} \\
 \midrule
-ResNet-18 (non-BN) & 1024 & Baseline & - & 66.27 & - & 64.816 & - \\
-& 1024 & PGT & 0.92 & \textcolor{red}{\textbf{66.62}} &
+ResNet-18 & 1024 & Baseline & - & 66.27 & - & 64.816 & - \\
+(non-BN) & 1024 & PGT & 0.92 & \textcolor{red}{\textbf{66.62}} &
 \textcolor{olive}{\textbf{+0.35}} & \textcolor{blue}{\textbf{65.498}} &
@@
 \multirow{2}{*}{\textbf{\#Row}} & \multirow{2}{*}{\textbf{Model}} &
-\multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} & \textbf{PowerGrad} &
+\multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} & \textbf{PGT} &
 \textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap (\%)}} \\
-& & & \textbf{Smoothing} & \textbf{Transform ($\alpha$)} & \textbf{Accuracy (\%)} &
-\textbf{Accuracy (\%)} & \\
+& & & \textbf{Smoothing} & \textbf{($\alpha$)} & \textbf{Acc. (\%)} &
+\textbf{Acc. (\%)} & \\
 \midrule
@@
 
-\bibliographystyle{model2-names}
+\bibliographystyle{splncs04}
 \bibliography{references}
@@
 \subsection{Norm plots of weights (baseline (low batch size)) of non-BN ResNet-18}
-\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{baseline-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{baseline-w_norm}
 \caption{ Method: baseline, batch size = 256. This is a plot depicting the
@@
 \begin{figure}[ht] \centering
-\includegraphics[width=0.55\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
+\includegraphics[width=0.57\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
 baseline (high batch size), batch size = 1024. This is a plot depicting the
@@
 \subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
-\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{pgt-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{pgt-w_norm}
 \caption{ Method: PGT, batch size = 256. This is a plot depicting the iteration-wise
@@
 \subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
-\begin{figure}[ht] \centering \includegraphics[width=0.55\textwidth]{agc_pgt-w_norm}
+\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{agc_pgt-w_norm}
 \caption{ Method: AGC + PGT, batch size = 256. This is a plot depicting the

### 2022-11-15 0100.tex
PAKDD Mod1

@@
 \usepackage{booktabs}
-% \usepackage{amsthm}
 \usepackage{pifont}
@@
 \title{Altering Backward Pass Gradients to improve Convergence}
-\author{Bishshoy Das\inst{1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{1}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Brejesh Lall\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Shiv Dutt Joshi\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Sumantra Dutta Roy\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}}
+\author{Bishshoy Das\inst{1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{1}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Brejesh Lall\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Shiv Dutt Joshi\inst{3}\orcidID{0000-0002-0326-0524}\and Sumantra Dutta Roy\inst{3}\orcidID{0000-0002-2141-5067}}
 \authorrunning{B. Das et al.}
@@
 
-\subsection{Properties of PGT}
+\subsection{Properties}
 \label{sec:pgt_prop}
@@
 
-\textit{Proof.} The proof follows directly from the definition of $p_i'$. We notice that
-if $p_i' < p_i$, we get:
-
-\begin{proof} \begin{alignat}{2} &\ \ \frac{p_i^{\alpha }}{\sum
-_{j=1}^C p_j^{\alpha}} < p_i \\ \Rightarrow &\ p_i^{\alpha -1}<\sum _{j=1}^C p_j^{\alpha
-} \\ \Rightarrow &\ p_i > \Big(\sum _{j=1}^C p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}}
-\end{alignat} \end{proof}
-
 We call this threshold, the \textit{stationary threshold}. The stationary threshold is
@@
 
-\textit{Proof.} From Eq. (\ref{eqn:threshold}), we see that the stationary threshold at
-$\alpha=0$ is $1/C$. Also, following from the definition of the transformed
-probabilities (Eq. \ref{transformed_probabilities}) we conclude that if $\alpha=0$, then
-all values of $p_i'$ are $1/C$. Therefore the transformed distribution at $\alpha=0$ is
-a uniform distribution.
-
-Since we have established that values of $p_i$ which are greater than the stationary
-threshold decreases and move down towards the stationary threshold, and values in $p_i$
-lower than the stationary threshold moves up towards the stationary threshold, this
+We have established that values of $p_i$ which are greater than the stationary threshold
+decreases and move down towards the stationary threshold, and values in $p_i$ lower than
+the stationary threshold moves up towards the stationary threshold. Therefore, this
 transformation makes the distribution more uniform (i.e. it smooths out the actual
 distribution) as $\alpha$ is decreased from $1$ and down towards $0$. This final
-observation we prove in the following theorem.
+observation in the following theorem.
 
@@
 \alpha \leq 1$ is a monotonically non-decreasing function with respect to $\alpha$.
-
-\textit{Proof.} To prove monotonicity, we first compute the gradient of the stationary
-threshold with respect to the variable in concern, $\alpha$.
-
-\begin{alignat}{2} & \frac{\partial }{\partial \alpha} \left(\sum _{j=1}^c
-p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ = \left(\sum _{j=1}^c
-p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left(\frac{\sum _{j=1}^c p_j^{\alpha } \log
-\left(p_j\right)}{(\alpha -1) \sum _{j=1}^c p_j^{\alpha}}-\frac{\log \left(\sum _{j=1}^c
-p_j^{\alpha}\right)}{(\alpha -1)^2}\right) \\
-& = \frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c p_j^{\alpha
-}\right){}^{\frac{1}{\alpha -1}} \left(\frac{(\alpha -1) \sum _{j=1}^C p_j^{\alpha }
-\log \left(p_j^{\alpha}\right)}{\sum _{j=1}^c p_j^{\alpha }}-\alpha \log \left(\sum
-_{j=1}^c p_j^{\alpha }\right)\right) \label{derivative} \end{alignat}
-
-If $a_1,\dots ,a_n$ and $b_1,\dots ,b_n$ are non-negative numbers, then using the log
-sum inequality, \\ we get $\sum _{j=1}^n a_j \log \left(\frac{a_j}{b_j}\right)\geq
-\left(\sum _{j=1}^n a_j\right) \log \left(\frac{\sum _{j=1}^n a_j}{\sum _{j=1}^n
-b_j}\right)$. Setting $a_j=p_j^\alpha$ and $b_j=1$, we get the following lower bound
-\begin{equation} \sum _{j=1}^C p_j^{\alpha } \log
-\left(p_j^{\alpha }\right)\geq \left(\sum _{j=1}^C
-p_j^{\alpha}\right) \log \left(\frac{1}{C} \sum _{j=1}^C
-p_j^{\alpha}\right) \label{lower_bound1} \end{equation}
-
-Substituting (\ref{lower_bound1}) in (\ref{derivative}), we get:
-
-\begin{equation} \frac{\partial }{\partial \alpha} \left(\sum
-_{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq
-\frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c
-p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left((1-\alpha )
-\log (C)-\log \left(\sum _{j=1}^C p_j^{\alpha }\right)\right)
-\label{lb_subs} \end{equation}
-
-$p^{\alpha}$ is concave, and so by Jensen's inequality we get the following upper bound
-for the second term:
-
-\begin{equation} \left(\frac{1}{C}{\sum _{j=1}^C
-p_j}{}\right){}^{\alpha }\ \ \geq\ \ \frac{1}{C}{\sum _{j=1}^C
-p_j^{\alpha }}{} \end{equation}
-
-\begin{equation}\Rightarrow\ \ \log \left(\sum _{j=1}^C p_j^{\alpha}\right)\leq
-(1-\alpha ) \log (C) \label{upper_bound} \end{equation}
-
-Substituting (\ref{upper_bound}) in (\ref{lb_subs}),
-
-\begin{proof} \begin{equation} \frac{\partial}{\partial \alpha
-}\left(\sum _{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq \ \ 0
-\end{equation} \end{proof}
 
@@
 
-\subsection{PGT restricts the partial derivative of the loss w.r.t. each logit from becoming too small or too high}
-\label{sec:pgt_logit_derivative}
-
-We know that neural networks uses the softmax function to generate prediction
-probabilities from logits in multi-class classification tasks. If we use cross-entropy
-loss to train the network then the partial derivative of the loss w.r.t. $i^{th}$ logit
-is dependent on the value of the predicted probability and the value of the class label
-(either 0 or 1) of the $i^{th}$ logit.
-
-\begin{equation}
-\label{actual_logit_derivative}
-\frac{\partial L}{\partial z_i} = p_i -q_i
-\end{equation}
-In general, the range of the $|\frac{\partial L}{\partial z_i}|$ is [$\delta$,
-$\epsilon$] where, $\delta\rightarrow 0$ and $\epsilon\rightarrow 1$ in traditional
-training (Eq. \ref{actual_logit_derivative} procedures of neural networks. Here, we
-propose a new method of neural network training which alters the gradients at the time
-of backward pass by modifying the predicted probability vector using PGT (Eq.
-\ref{PGT_logit_derivative}) such that the directional derivative for each logit does not
-become too small or too large,
-
-With inclusion of PGT in the last layer, the range of the $\widehat{|\frac{\partial
-L}{\partial z_i}|}$ becomes [$\delta^+$, $\epsilon^-$] where, $\delta^+>\delta$ and
-$\epsilon^-<\epsilon$
-
-Now, we demonstrate how PGT restricts the magnitude of the directional derivative for
-each logit. For a training example, a neural network can assign the highest probability
-to either (i) a wrong class or to the (ii) correct class.
-
-\textbf{(i) Wrong Class Predicted:}
-If the network assigns the highest probability for any class except the the true class,
-then wrong class assignment happens. We analyze the effect on gradient due to PGT for
-this case. There are three possibilities
-
-(a) Let, $i^{th}$ class be the class for which the network assigns low probability, but
-it is the true class, then
-
-$p_i < p_{th} \Rightarrow p'_i > p_i \Rightarrow \widehat{\frac{\partial L}{\partial
-z_i}} = p'_i -1 >  p_i -1 $ ; where $p_i\rightarrow 0$ \& $q_i = 1$
-
-So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| <  |p_i -1| \Rightarrow
-|\widehat{\frac{\partial L}{\partial z_i}}| <\epsilon $ where; $\epsilon\rightarrow 1$
-
-(b) Let, $j^{th}$ class be the class (not true class) for which the network assigns the
-highest probability, then
-
-$p_j > p_{th} \Rightarrow p'_j < p_j \Rightarrow \widehat{\frac{\partial L}{\partial
-z_j}} = p'_j <  p_j$ ; where $p_j\rightarrow 1$ \& $q_j = 0$
-
-So, $|\widehat{\frac{\partial L}{\partial z_j}}| <\epsilon$ where; $\epsilon\rightarrow
-1$
-
-(c) Let $k$ denote the index for all those classes for which the predicted probability
-is low and also the ground truth class label is $0$, then
-
-$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
-z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$
-
-So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow$ 0
-
-\textbf{(ii) Correct Class Predicted:}
-For a given training example, if the network assigns the highest probability to the true
-class then it assigns correct class to the training example. Now we observe the effect
-on gradient due to PGT for this case. There are two possibilities,
-
-(a) Let, $i^{th}$ class be the class for which the network assigns the highest
-probability and it is also the true class, then
-
-$p_i > p_{th} \Rightarrow p'_i < p_i \Rightarrow \widehat{\frac{\partial L}{\partial
-z_i}} = p'_i -1 <  p_i -1 $ ; where $p_i\rightarrow 1$ \& $q_j = 1$
-
-So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| >  |p_i -1|  \Rightarrow
-|\widehat{\frac{\partial L}{\partial z_i}}| >\delta $ where; $\delta\rightarrow 0$
-
-(b) Let, $k^{th}$ index is used for all those classes for which the predicted
-probability is low and also the actual label value $0$, then
-
-$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
-z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$
-
-So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow 0$
-
-After analyzing all these cases ((i)a,b,c \& (ii)a,b), we are able to deduce that PGT
-helps to restrict the directional derivative to be within limit such that for every
-logit/direction the directional derivative is not very small or very large.
-
-\subsection{Effect of PGT on weight gradients}
-\label{sec:pgt_weight_gradients}
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=1.0,width=0.5\columnwidth]{simplenn.jpg}
-\caption{Fully connected layer of a convolutional neural network} 
-\label{fig:simplenn}
-\end{figure}
-
-In this section, we discuss the impact of PGT on weight gradients. Let the input and
-output of the last convolutional layer of a convolutional neural network be \textbf{x}
-and \textbf{a} respectively (Fig. \ref{fig:simplenn}). The last layer which is a fully
-connected layer, produces logits \textbf{z} by combining \textbf{a} with weights and
-thereafter the network generates the predicted probability vector \textbf{p} from the
-logits using softmax activation function. We note that a$_i \geq 0$ $ \forall i$  as,
-\textbf{a} is the output of ReLU activation functions.
-
-Here, $z_i = \sum_{j=1}^{C}w_{ij}a_j$, so $\frac{\partial L}{\partial w_{ij}} =
-a_j\frac{\partial L}{\partial z_{i}} $. However, if we apply PGT while training the
-neural network, then $\widehat{\frac{\partial L}{\partial w_{ij}}}
-=a_j\widehat{\frac{\partial L}{\partial z_i}}$
-
-If the activation of the $j^{th}$ neuron is zero, i.e. $a_j = 0$ then there is zero
-gradient for all weights which acts on $a_j$ both for normal training and training with
-PGT, as $a_j$ has no contribution in the computation of logits. However, the interesting
-thing is to observe the effect of PGT when $a_j > 0$. As, the activation values $(a_j)$
-are positive, so $ \widehat{\frac{\partial L}{\partial z_i}}  > \frac{\partial
-L}{\partial z_{i}} \Rightarrow$ $\widehat{\frac{\partial L}{\partial w_{ij}}} >
-\frac{\partial L}{\partial w_{ij}} \forall j \in (1,2,\ldots C)$ and similar
-relationship for $a_j < 0$. With PGT, the weight updating rule for $(t+1)^{th}$
-iteration then becomes,
-\begin{equation}
-\label{weight_update_PGT}
-w_{ij}^{t+1} = w_{ij}^{t} - \eta \widehat{\frac{\partial L}{\partial w_{ij}^{t}}}
-\end{equation}
-The weight update equation using PGT, can also be viewed in terms of the following
-equation,
-\begin{equation}
-\label{weight_update_final_PGT}
-w_{ij}^{t+1} = w_{ij}^{t} - \frac{\eta}{C_{PGT}} \frac{\partial L}{\partial w_{ij}^{t}}
-\end{equation}
-
-For neural networks, the exponential nature of the softmax function produces sharp
-distribution for predicted probability vector. For example, if \textbf{z} is $[20, 30,
-10]$ then \textbf{p} is equal to $[4\times 10^{-5},  0.99995, 2\times10^{-9}]$. When the
-network is at the initial phase of training, most of the training examples are
-misclassified. Under this scenario, PGT helps to update the weights in a better manner
-than traditional gradients. Suppose the label vector \textbf{q} is $[1, 0, 0]$ then it
-indicates that $z_1$ is the logit corresponding to ground truth class and $z_2$ is the
-logit where the network has assigned the highest probability. Similarly, $z_3$ is the
-logit where the network has assigned a low probability as well as the label vector is
-also $0$ for this logit. Here $\frac{\partial L}{\partial w_{1j}^{t}} = \epsilon a_j > 0
-\Rightarrow$ $ \widehat{\frac{\partial L}{\partial w_{1j}^{t}}} < \frac{\partial
-L}{\partial w_{1j}^{t}} \Rightarrow C_{PGT} > 1$ $\forall j \in (1, 2,\ldots C)$.
-Advanced gradient updation algorithms like AdaGrad \cite{duchi2011adaptive} reduces the
-gradient in the direction where the directional derivative for a weight is high. It
-reduces the learning rate for the weight so that the gradient in that direction is not
-too large in order to make the update process less volatile. The formulation of PGT also
-implicitly reduces the partial derivative in the same manner for the high derivative
-directions, the only difference with the AdaGrad is that $C_{PGT}$ depends only on the
-current iteration's gradient and not on previous iterations like AdaGrad. Similarly we
-can observe that for this example, $\frac{\partial L}{\partial w_{3j}^{t}} = \delta a_j
-\approx 0 \Rightarrow$ $  \widehat{\frac{\partial L}{\partial w_{3j}^{t}}} >
-\frac{\partial L}{\partial w_{3j}^{t}} \Rightarrow C_{PGT} < 1$ $\forall j \in (1,
-2,\ldots C)$. As directional derivative for the weights associated with third logit is
-extremely low, PGT helps to increase the derivative in these directions so that the
-training process does not become too slow.
-
-\subsection{Effect of PGT on class separation}
-\label{sec:pgt_class_separation}
-
-During final iterations of training, we analyze the iteration-wise gradient behaviour
-where the network assigns the highest probability to the correct class for a given
-training example; we deduce the effects of PGT in the next immediate training iteration.
-For example, if \textbf{z} is $[30, 20, 10]$ then \textbf{p} is $[0.99995, 4\times
-10^{-5}, 2\times10^{-9}]$ and \textbf{q} is $[1, 0, 0]$. There is no change in the
-weight update due to PGT when $a_j = 0$. However, when $i \neq 1$ and $a_j > 0$,
-$|\widehat{\frac{\partial L}{\partial w_{1j}^{t}}}| > |\frac{\partial L}{\partial
-w_{1j}^{t}}|$ and $\widehat{\frac{\partial L}{\partial w_{ij}^{t}}} > \frac{\partial
-L}{\partial w_{ij}^{t}}$. Here in this example $\frac{\partial L}{\partial z_{1}^{t}} <
-0 \Rightarrow \frac{\partial L}{\partial w_{1j}^{t}} < 0 \Rightarrow$
-$\widehat{w_{1j}^{t+1}} > w_{1j}^{t+1} > w_{1j}^{t} \forall j$. We also know that
-$z_i^{t+1} = \sum_{j=1}^{C}w_{ij}^{t+1}a_j$ and $\widehat{z_i^{t+1}} =
-\sum_{j=1}^{C}\widehat{w_{ij}^{t+1}}a_j$. It indicates that $\widehat{z_{1}^{t+1}} >
-z_{1}^{t+1} > z_{1}^{t}$ as all the weights associated with first logit is larger while
-using PGT in $(t+1)^{th}$ iteration. Similarly, for the non-true classes
-$\widehat{w_{ij}^{t+1}} < w_{ij}^{t+1} < w_{ij}^{t}$ when $i \neq 1$ $\forall j$ as
-$\frac{\partial L}{\partial z_{i}^{t}} > 0$. So in $(t+1)^{th}$ iteration,
-$\widehat{z_{i}^{t+1}} < z_{i}^{t+1} < z_{i}^{t}$ when $i \neq 1$. We observe that
-$dist(\widehat{z_{1}^{t+1}}, \widehat{z_{i}^{t+1}}) > $  $dist(z_{1}^{t+1},
-z_{i}^{t+1})$ where the first logit corresponds to the true class and all other logits
-indexed by $i$ are non-true classes. Therefore the distance between correct and
-incorrect class logits increases due to PGT and this leads to better class separation.
+\subsection{Effects of PGT on logit, weight gradients and class separation}
+
+PGT, other than having the properties mentioned above, also:
+\begin{enumerate}
+\item restricts the magnitude of the directional derivative for each logit from becoming too small or too high
+\item enhances gradient behavior of weights in a manner that improves learning
+\item leads to better class separation.
+\end{enumerate}
+
+We demonstrate each of these effects in detail in supplementary sections
+\ref{sec:pgt_logit_derivative}, \ref{sec:pgt_weight_gradients} and
+\ref{sec:pgt_class_separation} respectively.
 
@@
 \multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Scheduler}} &
-\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{\hspace{-0.65cm} Train}
-& \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
-& & & \textbf{($\alpha$)} & \textbf{Acc. (\%)} & \textbf{Diff (\%)} &
-\textbf{Acc. (\%)} & \textbf{Diff (\%)} \\
+\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{Train}
+& \textbf{Train} & \textbf{Test} & \textbf{Test} \\
+& & & \textbf{($\alpha$)} & \textbf{Acc.(\%)} & \textbf{Diff(\%)} &
+\textbf{Acc.(\%)} & \textbf{Diff(\%)} \\
 \midrule
@@
 recipes for the experiments, pretrained checkpoints and training logs are provided at:
-\url{ https://github.com/bishshoy/power-grad-transform }. Code is adapted from
+\url{ https://github.com/skalien/power-grad-transform }. Code is adapted from
 \cite{rw2019timm}.}.
@@
 
-This section empirically examines some of the issues that occur when networks are
-trained without normalization layers and we provide insight into the effects of
-PowerGrad transform on such networks. We use ResNet-18 \cite{he2016deep} as the
-foundation model for all empirical trials owing to its popularity among deep learning
-practitioners and the relative simplicity with which it can be trained on a big dataset
-such as ImageNet-1K \cite{russakovsky2015imagenet}. We found deeper networks (such as
-ResNet-34 and ResNet-50 models) are impossible to train without BatchNormalization
-because of increased depth. So in order to conduct our empirical experiments, we solely
-focus on the ResNet-18 architecture. For our experiments, we designate different layers
-with their corresponding layer indices. We provide a detailed layer-wise index list of
-ResNet-18 in the supplementary section.
-
-We concentrate on the following metric: the per-filter L2-norm of each layer's weight
-tensor. Throughout the training process, we monitor variations in the per-filter weight
-norm. Because a full plot of each layer's data would use an inordinate amount of space,
-we focus on a few critical layers' plots. The supplementary section contains full plots
-for all layer for various settings. The norm of the weights of the $11^\textit{th}$
-convolutional layer is shown in Fig. \ref{fig:norm_plots}(a). Layer $11$ includes $256$
-filters. Each colour on the figure represents the iteration-wise evolution of a
-particular filter throughout the course of training. The critical point to note is that
-some filters achieve a norm of zero during training. We refer to this event as `Zeroing
-Out', and occurs when a channel of a weight tensor gets fully filled with zeros. If all
-the coefficients of the filter becomes zero, only then the L2 norm of that filter
-becomes zero. If we observe zero filter norm for a filter, then that filter does not
-contribute at all to determine the input-output relationship of a dataset, as the
-feature tensor it produces is also an all-zero tensor. According to our findings, when a
-filter tensor becomes zeroed out, it does not recover with further training. We observe
-both effects \textbf{1)} zeroed out filter \textbf{2)} zeroed out feature.
+We examine issues that occur in non-BatchNormalized networks and we provide insight into
+the effects of PGT on such networks. We use ResNet-18 \cite{he2016deep} as the
+foundation model trained on ImageNet-1K \cite{russakovsky2015imagenet}. We found deeper
+networks (such as ResNet-34 and ResNet-50 models) are impossible to train without Batch
+Normalization because of increased depth. So in order to conduct our empirical
+experiments, we solely focus on the ResNet-18 architecture. We designate different
+layers with their corresponding layer indices. We provide a detailed layer-wise index
+list of ResNet-18 in supplementary section \ref{sec:pgt_class_separation}.
+
+Throughout the training process, we monitor variations in the the per-filter L2-norm of
+each layer's weight tensor. Because a full plot of each layer's data would use an
+inordinate amount of space, we focus on a few critical layers' plots. The supplementary
+sections (\ref{sec:plots1} through \ref{sec:plots4}) contains full plots for all layer
+for various settings. The norm of the weights of the $11^\textit{th}$ convolutional
+layer is shown in Fig. \ref{fig:norm_plots}(a). Layer $11$ includes $256$ filters. Each
+colour on the figure represents the iteration-wise evolution of a particular filter
+throughout the course of training. The critical point to note is that some filters
+achieve a norm of zero during training. We refer to this event as `Zeroing Out', and
+occurs when a channel of a weight tensor gets fully filled with zeros. If all the
+coefficients of the filter becomes zero, only then the L2 norm of that filter becomes
+zero. If we observe zero filter norm for a filter, then that filter does not contribute
+at all to determine the input-output relationship of a dataset, as the feature tensor it
+produces is also an all-zero tensor. According to our findings, when a filter tensor
+becomes zeroed out, it does not recover with further training. We observe both effects
+\textbf{1)} zeroed out filter \textbf{2)} zeroed out feature.
 
@@
 \multirow{2}{*}{\textbf{Model}} & \multirow{2}{*}{\textbf{Batch Size}} &
-\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{\hspace{-0.65cm} Train}
-& \textbf{Train} & \textbf{\hspace{-0.65cm} Test} & \textbf{Test} \\
-& & & \textbf{($\alpha$)} & \textbf{Acc. (\%)} & \textbf{Diff (\%)} &
-\textbf{Acc. (\%)} & \textbf{Diff (\%)} \\
+\multirow{2}{*}{\textbf{Method}} & \textbf{PGT} & \textbf{Train}
+& \textbf{Train} & \textbf{Test} & \textbf{Test} \\
+& & & \textbf{($\alpha$)} & \textbf{Acc.(\%)} & \textbf{Diff(\%)} &
+\textbf{Acc.(\%)} & \textbf{Diff(\%)} \\
 \midrule
@@
 \textbf{(b)} Variations in the final loss values obtained for different $\alpha$
-settings. \textbf{(c)} Correlation between loss and logits. Regression line equation
-:$(0.01074 x + 0.25816)$. }
+settings. \textbf{(c)} Correlation between loss and logits. Regression line equation:
+$(0.01074 x + 0.25816)$. }
 \label{fig:stats}
@@
 \centering
-\caption{ Ablation study for ResNet-50 on ImageNet-1K. LS column denotes whether label
-smoothing has been applied. A smoothing hyperparameter value of $0.1$ has been used. The
-hyperparameter column that controls the amount of PGT gradient modification injected at
-softmax, Eq. \ref{transformed_probabilities}. }
+\caption{ Ablation study for ResNet-50 on ImageNet-1K. }
 \label{tab:ablation_table}
@@
 \multirow{2}{*}{\textbf{Scheduler}} & \textbf{Label} & \textbf{PGT} &
-\textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap (\%)}} \\
-& & & \textbf{Smoothing} & \textbf{($\alpha$)} & \textbf{Acc. (\%)} &
-\textbf{Acc. (\%)} & \\
+\textbf{Train} & \textbf{Test} & \multirow{2}{*}{\textbf{Gap(\%)}} \\
+& & & \textbf{Smoothing} & \textbf{($\alpha$)} & \textbf{Acc.(\%)} &
+\textbf{Acc.(\%)} & \\
 \midrule
@@
 
+% \end{document}
+
 \bibliographystyle{splncs04}
@@
 \label{sec:Supp}
+
+\subsection{Proofs}
+
+\textbf{Lemma 1.} For any arbitrary probability distribution $P$ with probability values
+given by $p_i$ for $i=1,\dots,C$, the corresponding transformed probability values
+$p_i'$ given by [Eq. \ref{transformed_probabilities}] has a threshold $\Big(\sum
+_{j=1}^C p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}}$ and
+
+\begin{equation} \begin{split} p_i' \geq p_i & \text{,\ \ if } p_i \leq
+\Big(\sum _{j=1}^C p_j^{\alpha }\Big)^{\frac{1}{\alpha-1}} \\ \hspace*{1cm}
+p_i' < p_i & \text{,\ \ if } p_i > \Big(\sum _{j=1}^C
+p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}} \end{split} \label{eqn:threshold}
+\end{equation}
+
+\textbf{Proposition 1.} At $\alpha=0$, the stationary threshold equals $1/C$ and all
+values of the transformed distribution $p_i'$ reduces reduces to the uniform
+distribution for $i=1,\dots,C$,.
+
+\textit{Proof.} From Eq. (\ref{eqn:threshold}), we see that the stationary threshold at
+$\alpha=0$ is $1/C$. Also, following from the definition of the transformed
+probabilities (Eq. \ref{transformed_probabilities}) we conclude that if $\alpha=0$, then
+all values of $p_i'$ are $1/C$. Therefore the transformed distribution at $\alpha=0$ is
+a uniform distribution.
+
+\textbf{Theorem 1.} For any arbitrary probability distribution $P$ with probability
+values $p_i$ for $i=1,\dots,C$, the stationary threshold of the transformed distribution
+$P'$ with probability values $p_i'=\frac{p_i^{\alpha}}{\sum _{j=1}^C p_j^\alpha}, 0 \leq
+\alpha \leq 1$ is a monotonically non-decreasing function with respect to $\alpha$.
+
+\textit{Proof.} To prove monotonicity, we first compute the gradient of the stationary
+threshold with respect to the variable in concern, $\alpha$.
+
+\begin{alignat}{2} & \frac{\partial }{\partial \alpha} \left(\sum _{j=1}^c
+p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ = \left(\sum _{j=1}^c
+p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left(\frac{\sum _{j=1}^c p_j^{\alpha } \log
+\left(p_j\right)}{(\alpha -1) \sum _{j=1}^c p_j^{\alpha}}-\frac{\log \left(\sum _{j=1}^c
+p_j^{\alpha}\right)}{(\alpha -1)^2}\right) \\
+& = \frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c p_j^{\alpha
+}\right){}^{\frac{1}{\alpha -1}} \left(\frac{(\alpha -1) \sum _{j=1}^C p_j^{\alpha }
+\log \left(p_j^{\alpha}\right)}{\sum _{j=1}^c p_j^{\alpha }}-\alpha \log \left(\sum
+_{j=1}^c p_j^{\alpha }\right)\right) \label{derivative} \end{alignat}
+
+If $a_1,\dots ,a_n$ and $b_1,\dots ,b_n$ are non-negative numbers, then using the log
+sum inequality, \\ we get $\sum _{j=1}^n a_j \log \left(\frac{a_j}{b_j}\right)\geq
+\left(\sum _{j=1}^n a_j\right) \log \left(\frac{\sum _{j=1}^n a_j}{\sum _{j=1}^n
+b_j}\right)$.
+
+Setting $a_j=p_j^\alpha$ and $b_j=1$, we get the following lower bound
+\begin{equation} \sum _{j=1}^C p_j^{\alpha } \log
+\left(p_j^{\alpha }\right)\geq \left(\sum _{j=1}^C
+p_j^{\alpha}\right) \log \left(\frac{1}{C} \sum _{j=1}^C
+p_j^{\alpha}\right) \label{lower_bound1} \end{equation}
+
+Substituting (\ref{lower_bound1}) in (\ref{derivative}), we get:
+
+\begin{equation} \frac{\partial }{\partial \alpha} \left(\sum
+_{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq
+\frac{1}{\alpha  (\alpha -1)^2} \left(\sum _{j=1}^c
+p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}} \left((1-\alpha )
+\log (C)-\log \left(\sum _{j=1}^C p_j^{\alpha }\right)\right)
+\label{lb_subs} \end{equation}
+
+$p^{\alpha}$ is concave, and so by Jensen's inequality we get the following upper bound
+for the second term:
+
+\begin{equation} \left(\frac{1}{C}{\sum _{j=1}^C
+p_j}{}\right){}^{\alpha }\ \ \geq\ \ \frac{1}{C}{\sum _{j=1}^C
+p_j^{\alpha }}{} \end{equation}
+
+\begin{equation}\Rightarrow\ \ \log \left(\sum _{j=1}^C p_j^{\alpha}\right)\leq
+(1-\alpha ) \log (C) \label{upper_bound} \end{equation}
+
+Substituting (\ref{upper_bound}) in (\ref{lb_subs}),
+
+\begin{equation} \frac{\partial}{\partial \alpha
+}\left(\sum _{j=1}^c p_j^{\alpha}\right){}^{\frac{1}{\alpha -1}}\ \ \geq \ \ 0
+\end{equation}
+
+\clearpage
+
+\subsection{PGT restricts the partial derivative of the loss w.r.t. each logit from becoming too small or too high}
+\label{sec:pgt_logit_derivative}
+
+We know that neural networks uses the softmax function to generate prediction
+probabilities from logits in multi-class classification tasks. If we use cross-entropy
+loss to train the network then the partial derivative of the loss w.r.t. $i^{th}$ logit
+is dependent on the value of the predicted probability and the value of the class label
+(either 0 or 1) of the $i^{th}$ logit.
+
+\begin{equation}
+\label{actual_logit_derivative}
+\frac{\partial L}{\partial z_i} = p_i -q_i
+\end{equation}
+In general, the range of the $|\frac{\partial L}{\partial z_i}|$ is [$\delta$,
+$\epsilon$] where, $\delta\rightarrow 0$ and $\epsilon\rightarrow 1$ in traditional
+training (Eq. \ref{actual_logit_derivative} procedures of neural networks. Here, we
+propose a new method of neural network training which alters the gradients at the time
+of backward pass by modifying the predicted probability vector using PGT (Eq.
+\ref{PGT_logit_derivative}) such that the directional derivative for each logit does not
+become too small or too large,
+
+With inclusion of PGT in the last layer, the range of the $\widehat{|\frac{\partial
+L}{\partial z_i}|}$ becomes [$\delta^+$, $\epsilon^-$] where, $\delta^+>\delta$ and
+$\epsilon^-<\epsilon$
+
+Now we demonstrate how PGT restricts the magnitude of the directional derivative for
+each logit. For a training example, a neural network can assign the highest probability
+to either (i) a wrong class or to the (ii) correct class.
+
+\textbf{(i) Wrong Class Predicted:}
+If the network assigns the highest probability for any class except the the true class,
+then wrong class assignment happens. We analyze the effect on gradient due to PGT for
+this case. There are three possibilities
+
+(a) Let, $i^{th}$ class be the class for which the network assigns low probability, but
+it is the true class, then
+
+$p_i < p_{th} \Rightarrow p'_i > p_i \Rightarrow \widehat{\frac{\partial L}{\partial
+z_i}} = p'_i -1 >  p_i -1 $ ; where $p_i\rightarrow 0$ \& $q_i = 1$
+
+So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| <  |p_i -1| \Rightarrow
+|\widehat{\frac{\partial L}{\partial z_i}}| <\epsilon $ where; $\epsilon\rightarrow 1$
+
+(b) Let, $j^{th}$ class be the class (not true class) for which the network assigns the
+highest probability, then
+
+$p_j > p_{th} \Rightarrow p'_j < p_j \Rightarrow \widehat{\frac{\partial L}{\partial
+z_j}} = p'_j <  p_j$ ; where $p_j\rightarrow 1$ \& $q_j = 0$
+
+So, $|\widehat{\frac{\partial L}{\partial z_j}}| <\epsilon$ where; $\epsilon\rightarrow
+1$
+
+(c) Let $k$ denote the index for all those classes for which the predicted probability
+is low and also the ground truth class label is $0$, then
+
+$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
+z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$
+
+So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow$ 0
+
+\textbf{(ii) Correct Class Predicted:}
+For a given training example, if the network assigns the highest probability to the true
+class then it assigns correct class to the training example. Now we observe the effect
+on gradient due to PGT for this case. There are two possibilities,
+
+(a) Let, $i^{th}$ class be the class for which the network assigns the highest
+probability and it is also the true class, then
+
+$p_i > p_{th} \Rightarrow p'_i < p_i \Rightarrow \widehat{\frac{\partial L}{\partial
+z_i}} = p'_i -1 <  p_i -1 $ ; where $p_i\rightarrow 1$ \& $q_j = 1$
+
+So; $|\widehat{\frac{\partial L}{\partial z_i}}| = |p'_i -1| >  |p_i -1|  \Rightarrow
+|\widehat{\frac{\partial L}{\partial z_i}}| >\delta $ where; $\delta\rightarrow 0$
+
+(b) Let, $k^{th}$ index is used for all those classes for which the predicted
+probability is low and also the actual label value $0$, then
+
+$p_k < p_{th} \Rightarrow p'_k > p_k \Rightarrow \widehat{\frac{\partial L}{\partial
+z_k}} = p'_k >  p_k$ ; where $p_k\rightarrow 0$ \& $q_k = 0$
+
+So, $|\widehat{\frac{\partial L}{\partial z_k}}| >\delta$ where; $\delta\rightarrow 0$
+
+After analyzing all these cases ((i)a,b,c \& (ii)a,b), we are able to deduce that PGT
+helps to restrict the directional derivative to be within limit such that for every
+logit/direction the directional derivative is not very small or very large.
+
+\subsection{Effect of PGT on weight gradients}
+\label{sec:pgt_weight_gradients}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=1.0,width=0.5\columnwidth]{simplenn.jpg}
+\caption{Fully connected layer of a convolutional neural network} 
+\label{fig:simplenn}
+\end{figure}
+
+In this section, we discuss the impact of PGT on weight gradients. Let the input and
+output of the last convolutional layer of a convolutional neural network be \textbf{x}
+and \textbf{a} respectively (Fig. \ref{fig:simplenn}). The last layer which is a fully
+connected layer, produces logits \textbf{z} by combining \textbf{a} with weights and
+thereafter the network generates the predicted probability vector \textbf{p} from the
+logits using softmax activation function. We note that a$_i \geq 0$ $ \forall i$  as,
+\textbf{a} is the output of ReLU activation functions.
+
+Here, $z_i = \sum_{j=1}^{C}w_{ij}a_j$, so $\frac{\partial L}{\partial w_{ij}} =
+a_j\frac{\partial L}{\partial z_{i}} $. However, if we apply PGT while training the
+neural network, then $\widehat{\frac{\partial L}{\partial w_{ij}}}
+=a_j\widehat{\frac{\partial L}{\partial z_i}}$
+
+If the activation of the $j^{th}$ neuron is zero, i.e. $a_j = 0$ then there is zero
+gradient for all weights which acts on $a_j$ both for normal training and training with
+PGT, as $a_j$ has no contribution in the computation of logits. However, the interesting
+thing is to observe the effect of PGT when $a_j > 0$. As, the activation values $(a_j)$
+are positive, so $ \widehat{\frac{\partial L}{\partial z_i}}  > \frac{\partial
+L}{\partial z_{i}} \Rightarrow$ $\widehat{\frac{\partial L}{\partial w_{ij}}} >
+\frac{\partial L}{\partial w_{ij}} \forall j \in (1,2,\ldots C)$ and similar
+relationship for $a_j < 0$. With PGT, the weight updating rule for $(t+1)^{th}$
+iteration then becomes,
+\begin{equation}
+\label{weight_update_PGT}
+w_{ij}^{t+1} = w_{ij}^{t} - \eta \widehat{\frac{\partial L}{\partial w_{ij}^{t}}}
+\end{equation}
+The weight update equation using PGT, can also be viewed in terms of the following
+equation,
+\begin{equation}
+\label{weight_update_final_PGT}
+w_{ij}^{t+1} = w_{ij}^{t} - \frac{\eta}{C_{PGT}} \frac{\partial L}{\partial w_{ij}^{t}}
+\end{equation}
+
+For neural networks, the exponential nature of the softmax function produces sharp
+distribution for predicted probability vector. For example, if \textbf{z} is $[20, 30,
+10]$ then \textbf{p} is equal to $[4\times 10^{-5},  0.99995, 2\times10^{-9}]$. When the
+network is at the initial phase of training, most of the training examples are
+misclassified. Under this scenario, PGT helps to update the weights in a better manner
+than traditional gradients. Suppose the label vector \textbf{q} is $[1, 0, 0]$ then it
+indicates that $z_1$ is the logit corresponding to ground truth class and $z_2$ is the
+logit where the network has assigned the highest probability. Similarly, $z_3$ is the
+logit where the network has assigned a low probability as well as the label vector is
+also $0$ for this logit. Here $\frac{\partial L}{\partial w_{1j}^{t}} = \epsilon a_j > 0
+\Rightarrow$ $ \widehat{\frac{\partial L}{\partial w_{1j}^{t}}} < \frac{\partial
+L}{\partial w_{1j}^{t}} \Rightarrow C_{PGT} > 1$ $\forall j \in (1, 2,\ldots C)$.
+Advanced gradient updation algorithms like AdaGrad \cite{duchi2011adaptive} reduces the
+gradient in the direction where the directional derivative for a weight is high. It
+reduces the learning rate for the weight so that the gradient in that direction is not
+too large in order to make the update process less volatile. The formulation of PGT also
+implicitly reduces the partial derivative in the same manner for the high derivative
+directions, the only difference with the AdaGrad is that $C_{PGT}$ depends only on the
+current iteration's gradient and not on previous iterations like AdaGrad. Similarly we
+can observe that for this example, $\frac{\partial L}{\partial w_{3j}^{t}} = \delta a_j
+\approx 0 \Rightarrow$ $  \widehat{\frac{\partial L}{\partial w_{3j}^{t}}} >
+\frac{\partial L}{\partial w_{3j}^{t}} \Rightarrow C_{PGT} < 1$ $\forall j \in (1,
+2,\ldots C)$. As directional derivative for the weights associated with third logit is
+extremely low, PGT helps to increase the derivative in these directions so that the
+training process does not become too slow.
+
+\subsection{Effect of PGT on class separation}
+\label{sec:pgt_class_separation}
+
+During final iterations of training, we analyze the iteration-wise gradient behaviour
+where the network assigns the highest probability to the correct class for a given
+training example; we deduce the effects of PGT in the next immediate training iteration.
+For example, if \textbf{z} is $[30, 20, 10]$ then \textbf{p} is $[0.99995, 4\times
+10^{-5}, 2\times10^{-9}]$ and \textbf{q} is $[1, 0, 0]$. There is no change in the
+weight update due to PGT when $a_j = 0$. However, when $i \neq 1$ and $a_j > 0$,
+$|\widehat{\frac{\partial L}{\partial w_{1j}^{t}}}| > |\frac{\partial L}{\partial
+w_{1j}^{t}}|$ and $\widehat{\frac{\partial L}{\partial w_{ij}^{t}}} > \frac{\partial
+L}{\partial w_{ij}^{t}}$. Here in this example $\frac{\partial L}{\partial z_{1}^{t}} <
+0 \Rightarrow \frac{\partial L}{\partial w_{1j}^{t}} < 0 \Rightarrow$
+$\widehat{w_{1j}^{t+1}} > w_{1j}^{t+1} > w_{1j}^{t} \forall j$. We also know that
+$z_i^{t+1} = \sum_{j=1}^{C}w_{ij}^{t+1}a_j$ and $\widehat{z_i^{t+1}} =
+\sum_{j=1}^{C}\widehat{w_{ij}^{t+1}}a_j$. It indicates that $\widehat{z_{1}^{t+1}} >
+z_{1}^{t+1} > z_{1}^{t}$ as all the weights associated with first logit is larger while
+using PGT in $(t+1)^{th}$ iteration. Similarly, for the non-true classes
+$\widehat{w_{ij}^{t+1}} < w_{ij}^{t+1} < w_{ij}^{t}$ when $i \neq 1$ $\forall j$ as
+$\frac{\partial L}{\partial z_{i}^{t}} > 0$. So in $(t+1)^{th}$ iteration,
+$\widehat{z_{i}^{t+1}} < z_{i}^{t+1} < z_{i}^{t}$ when $i \neq 1$. We observe that
+$dist(\widehat{z_{1}^{t+1}}, \widehat{z_{i}^{t+1}}) > $  $dist(z_{1}^{t+1},
+z_{i}^{t+1})$ where the first logit corresponds to the true class and all other logits
+indexed by $i$ are non-true classes. Therefore the distance between correct and
+incorrect class logits increases due to PGT and this leads to better class separation.
 
@@
 \subsection{Norm plots of weights (baseline (low batch size)) of non-BN ResNet-18}
+\label{sec:plots1}
 \begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{baseline-w_norm}
@@
 \subsection{Norm plots of weights (baseline (high batch size)) of non-BN ResNet-18}
+\label{sec:plots2}
 \begin{figure}[ht] \centering
@@
 \subsection{Norm plots of weights (PGT) of non-BN ResNet-18}
+\label{sec:plots3}
 \begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{pgt-w_norm}
@@
 \subsection{Norm plots of weights (AGC + PGT) of non-BN ResNet-18}
+\label{sec:plots4}
 \begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{agc_pgt-w_norm}
@@
 \subsection{ResNet-18 Layer Index Diagram}
+\label{sec:r18_layers}
 

### 2022-11-17 0100.tex
Mod2

@@
 \usepackage{pifont}
+\usepackage{wrapfig}
 \usepackage{graphicx}
@@
 \graphicspath{{images/}}
+\raggedbottom
 \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 \author{Bishshoy Das\inst{1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{1}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Brejesh Lall\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Shiv Dutt Joshi\inst{3}\orcidID{0000-0002-0326-0524}\and Sumantra Dutta Roy\inst{3}\orcidID{0000-0002-2141-5067}}
-\authorrunning{B. Das et al.}
+\authorrunning{Das et al.}
 \institute{Indian Institute of Technology Delhi, New Delhi - 110016, India \\ \email{\{bishshoy.das, milton.mondal, brejesh, sdjoshi, sumantra\}@ee.iitd.ac.in}}
@@
 
-Backpropagation is traditionally used to train deep neural networks
+Backpropagation is traditionally used for training deep neural networks
 \cite{lillicrap2020backpropagation}. Gradients are computed using basic calculus
@@
 several ways to achieve gradient modification in the backward pass. We discuss a few
-techniques in Fig. \ref{fig:gradient_altering}. One could modify the gradients at either
-\textbf{(I)} multiple points in the backward computation graph or \textbf{(II)} at a
-point very early in the backward graph and allow this changed gradient to also affect
-the rest of the network's gradients as it backpropagates through the backward
-computation graph.
+techniques in Fig. \ref{fig:gradient_altering}.
 
@@
 \caption{ Different ways of altering gradients in the backward pass. Blue blocks denote
-a CNN's different layers (convolutional, linear, batchnorm, softmax, etc.). Orange
-blocks indicate the backward graph with unmodified gradients. Green blocks represent
-transformation functions, while red blocks indicate transformed gradients. (a)
-\textbf{Type 0:} Backpropagation without any gradient modification. (b) \textbf{Type I:}
-Altering gradients after they are all computed. (c) \textbf{Type II:} Altering gradients
-at an earlier stage in the backward computation graph, subsequently leading to
-alteration of all other gradients. }
+a different layers. Orange blocks indicate the backward graph with unmodified gradients.
+Green blocks represent transformation functions, while red blocks indicate transformed
+gradients. }
 \label{fig:gradient_altering}
+\vspace{-0.5cm}
 \end{figure}
@@
 
-\textbf{Type I:} Gradient modification at multiple points independently: In this method,
+\textbf{Type I:} Independent gradient modification at multiple points: In this method,
 the gradients are first computed using standard procedure and then individually altered
@@
 
-Type I modification is computationally more expensive than Type II modification as it
-requires altering the gradients of each and every parameter individually. Type II
-modification recomputes gradients at each and every location through the natural flow of
-backpropagation. We propose PowerGrad Transform (PGT) which is a type II gradient
-modification method. With virtually no additional compute or memory requirement,
-PowerGrad Transform enhances classification performance of a convolutional neural
-network, by leading the network to arrive at a higher quality optima at convergence, at
-which both training and test accuracies improve.
-
-PowerGrad Transform modifies the gradients at the softmax layer, which is the earliest
-part of most convolutional neural network's backward computation graph. By changing the
-gradient at this stage, all other gradients for all other parameters in the network such
-as the linear layer, batchnorm statistics and convolutional layer's filter parameters
-are affected. We analyze and explore the behaviour of such alteration both
-mathematically and through empirical experiments conducted on networks without
-batch-normalization. Experimentally, we see that PGT improved training and test
-accuracies for both normalized and non-normalized networks. PGT improves a network's
-learning capacity and results in a better fit of the training data.
+Type I is computationally more expensive than type II as it requires altering the
+gradients of each and every parameter individually. Type II modification recomputes
+gradients at each and every location through the natural flow of backpropagation. We
+propose PowerGrad Transform (PGT), a type II method that modifies the gradients at the
+softmax layer.
 
@@
 
-\item We introduce PowerGrad Transform, which decouples the backward and the forward
-passes of neural network training and enables a considerably better fit to the dataset,
-as assessed by both training and test accuracy measures. PGT is a performance
-enhancement method that alters the gradients in the backward pass before the update step
-leading to accelerated training and a significant boost in the network's predictive
-performance.
+\item We propose PowerGrad Transform, which decouples the backward and the forward
+passes of neural network training and enables a considerably better fit to the dataset.
+PGT is a performance enhancement method that alters the gradients in the backward pass
+before the update step leading to accelerated training and a significant boost in the
+network's predictive performance.
 
 \item We perform theoretical analysis of the properties of the PowerGrad transformation
-(section \ref{sec:pgt_prop}) and explore its effect on weight parameters and gradients
-(section \ref{sec:pgt_weight_gradients}), logits (section
-\ref{sec:pgt_logit_derivative}), (section \ref{sec:Effe}), loss values (section
-\ref{sec:Effe}) and class separation (section \ref{sec:pgt_class_separation}).
-
-\item We study experimentally the degenerate behaviour of non-BN networks that is often
-seen during the normal training procedure, particularly at larger batch sizes. With PGT,
-we see improved gradient behaviour and a decreased likelihood of the weights attaining
-degenerate states.
+(section \ref{sec:pgt_prop}) and explore its effect on weight parameters and gradients,
+logits, loss values and class separation (section \ref{sec:pgt_effects}).
 
 \item We provide complete results from a variety of models (non-BN and BN ResNets,
-SE-ResNets) using the ImageNet dataset. It helps the network to improve its learning
-capabilities by locating a more optimum convergence point. Additionally, we conduct an
-ablation study and compare its impacts to those of regularization approaches such as
-label smoothing.
+SE-ResNets) using the ImageNet dataset. We empirically conclude that PGT helps a network
+to improve by locating a more optimum convergence point. Additionally, we conduct
+ablation studies and compare its effects to those of regularization methods.
 
@@
 
-We examine techniques that are related to gradient modification and emphasize the
-distinction between them and our proposed method.
-
-Gradient Clipping (GC): Gradient Clipping \cite{pascanu2013difficulty}, often used in
-natural language processing methods \cite{merity2017regularizing}, is a technique that
-involves changing or clipping the gradients with respect to a predefined threshold value
-during backward propagation through the network and updating the weights using the
-clipped gradients \cite{zhang2019gradient} \cite{smith2020generalization}. By rescaling the
-gradients, the weight updates are likewise rescaled, significantly reducing the risk of
-an overflow or underflow \cite{pascanu2012understanding}. GC can be used for training
-networks without batch-normalization.
-
-The formulation is as follows: If $G$ is a gradient vector which is $\partial L /
-\partial \theta$ (the gradient of the loss $L$ with respect to the parameters $\theta$).
-In clip by value, the gradients are modified as:
-
-\begin{equation}
-G \rightarrow
-\begin{cases}
-\lambda \frac{G}{\norm{G}} & \text{if } \norm{G} > \lambda, \\
-G & \text{otherwise}
-\end{cases}
-\end{equation}
-
-At larger batch sizes, the clipping threshold in GC becomes highly sensitive and
-requires extensive finetuning for various models, batch sizes, and learning rates. As we
-demonstrate later in our studies, GC is not as effective in improving the performance of
-non-normalized networks. AGC performs better than GC in non-normalized networks.
-However, we show that PGT outperforms both in such networks.
-
-Adaptive Gradient Clipping (AGC): Adaptive Gradient Clipping \cite{brock2021high} is
-developed to further enhance backward pass gradients than what is performed by GC. It
-takes into account the fact that the ratio of the gradient norm to the weight norm can
-provide an indication of the expected change in a single step of optimization. Here the
-normalized gradient is clipped by $\lambda$, which means that large weights can have
-large gradients and still the normalized gradient can be within $\lambda$. If it is not
-within the threshold, then their gradient is normalized by the ratio of norm of the
-gradient and norm of the weight in order to avoid gradient explosion. If $W^l$ is the
-weight matrix of the $l^\textit{th}$ layer, and $G^l$ is its gradient in the backward
-pass, then the following equation is used to modify the gradient of the $i^\textit{th}$
-filter before its update:
-
-\begin{equation}
-G_{i}^{l}\rightarrow
-\begin{cases}
-\lambda \frac{\norm{W_{i}^{l} }_{F}^{*}}{\norm{G_{i}^{l} }_{F}} G_{i}^{l} & if\ \frac{\norm{G_{i}^{l} }_{F}}{\norm{W_{i}^{l} }_{F}^{*}} \  >\ \lambda \\
-G_{i}^{l} & otherwise
-\end{cases}
-\end{equation}
-
-The hyperparameter $\lambda$ of AGC is less sensitive to changes in batch size and depth
-as observed in \cite{brock2021high}. However, when applied to vanilla residual networks
-(ResNet-18), the performance is still afar from normalized variants as demonstrated by
-our experiments. In most cases, AGC is used for training networks without
-batch-normalization. Our experiments demonstrate that PGT outperforms AGC, when used
-independently. When used together, it can further improve a network's performance
-(section \ref{sec:Empi}).
-
-Label Smoothing: Label smoothing, introduced by Szegedy et al.
+Gradient Clipping \cite{pascanu2013difficulty} is a technique that involves changing or
+clipping the gradients with respect to a predefined threshold value during backward
+propagation through the network and updating the weights using the clipped gradients
+\cite{zhang2019gradient}\cite{smith2020generalization}. By rescaling the gradients, the
+weight updates are likewise rescaled, significantly reducing the risk of an overflow or
+underflow \cite{pascanu2012understanding}. GC can be used for training networks without
+batch-normalization. At larger batch sizes, the clipping threshold in GC becomes highly
+sensitive and requires extensive finetuning for various models, batch sizes, and
+learning rates. As we demonstrate later in our studies, GC is not as effective in
+improving the performance of non-normalized networks. Adaptive Gradient Clipping
+\cite{brock2021high} is developed to further enhance backward pass gradients than what
+is performed by GC. It takes into account the fact that the ratio of the gradient norm
+to the weight norm can provide an indication of the expected change in a single step of
+optimization. Label smoothing, introduced by Szegedy et al.
 \cite{szegedy2016rethinking}, utilizes smoothing of the ground truth labels as a method
-to impose regularization on the logits and the weights of the neural network.
-
-The formulation is as follows. If $q_i$ is the value at the $i^{th}$ index of the one
-hot encoded vector of the ground truth label, then the transformed distribution of the
-labels $q_i'$ is: \begin{equation} q_i' = \begin{cases} 1-\epsilon & \text{if } i = y,
-\\ \frac{\epsilon}{K-1} & otherwise \end{cases} \end{equation} where $\epsilon$ is a
-hyperparameter with a value in $[0,1]$ and $y$ denotes the index of the correct class.
-
-Müller et al. investigates how label smoothing work in \cite{DBLP:conf/nips/MullerKH19}.
-Using visualizations of penultimate layer activations, they demonstrate that label
-smoothing calibrates the networks predictions and aligns them with the accuracies of the
-predictions. Parallel works in label smoothing include Pairwise Confusion
-\cite{dubey2018pairwise}, combatting Label Noise \cite{reed2014training}, achieving
-regularization through intentional label modification \cite{xie2016disturblabel}.
-
-PowerGrad Transform works in a much different way. PGT does not smooth the ground truth
-labels, rather it modifies the gradient by smoothing the predicted probability vector.
-As we show in section \ref{sec:Math}, our proposed transformation has the opposite
-effect of label smoothing. From our studies, we show that PGT in fact leads to larger
-generalization gap between the training and test dataset, although it does lead to an
-increase in both training and test accuracies at convergence. Coupled with PGT, label
-smoothing can narrow the generalization gap, while retaining the benefits of PGT.
-
-Knowledge Distillation: The formulation of PowerGrad Transform is akin to temperature
-based logit suppression of \cite{hinton2015distilling}. Knowledge distillation
-\cite{DBLP:conf/nips/BaC14} is a process in which two networks are trained. First a
-teacher network is trained on a given dataset and then the soft-labels (the predicted
-probabilities) of the teacher network are used to train another network, called the
-student network. In similar style of label smoothing, knowledge distillation aims to
-generate smooth gradient behaviour by forcing the network not to overpredict on any
-single image. As in case of label smoothing, the student network's weights are
-automatically penalized if the network assigns a probability value higher than the soft
-labels generated by the teacher network. Variants of knowledge distillation include
-self-distillation \cite{zhang2019your} \cite{wang2021memory}, identical student network
-distillation \cite{furlanello2018born}, channel distillation \cite{ge2019distilling},
-regularizing wrong predictions \cite{yun2020regularizing} and representation or
-embedding based knowledge distillation \cite{aguilar2020knowledge} \cite{yao2018deep}
-\cite{passalis2018unsupervised}. Distillation is applied to various tasks such as model
-compression \cite{wang2020real}, face recognition \cite{ge2018low}, enhancing ensemble
-performance \cite{zhang2018deep}, interpreting neural networks \cite{liu2018improving}.
-Efficacy of distillation is extensivly studied in \cite{cho2019efficacy}
-\cite{yuan2020revisiting}. Applications of logit suppression include: deep metric
-learning \cite{zhai2018classification}, non-parametric neighborhood component analysis
-\cite{wu2018improving}, sequence-to-sequence modelling \cite{chorowski2016towards},
-reinforcement learning \cite{he2018determining} and face recognition
-\cite{liu2017sphereface} \cite{wang2018cosface}.
-
-The key difference between PGT and distillation methods is that in the latter, the
-transformation is applied in the forward pass, while PGT is a backward pass modification
-only. Also, in distillation settings the temperature parameter is a part of the
-network's computation graph. In the case of PowerGrad Transform, we directly tamper the
-gradients without introducing any change in the forward pass. It is a more generalized
-way of arriving at altered gradients in the backward pass. Moreover we show that
-modifying the gradients in the backward pass can improve performance in many cases. The
-rise in training accuracy indicates that PGT enables the network to learn more from the
-training data.
-
-Self-knowledge distillation is used in many papers in different ways, which mainly
-focuses on improving the performance of the model by either creating multiple ensemble
-student models \cite{zhang2018deep}, using different distorted versions of same data
-with two different versions of a single network which guides one another
-\cite{xu2019data} or injecting multiple sub-modules in the network to improve its
-capacity \cite{zhang2019your}. However none of these methods use normal traditional
-training without introducing sub-modules in the architecture. PGT differs from
-self-knowledge distillation as it neither introduces any additional sub-modules nor
-creates different ensembles to improve the performance of the model. PGT follows the
-standard neural network training mechanism with a single change which is altering the
-last layer's gradients in the backward pass.
-
-Other techniques of weight update include slot machines \cite{aladago2021slot}, where
-random weights are generated and carefully selected during backpropagation in a way that
-minimizes the loss function.
+to impose regularization on the logits and the weights of the neural network. AGC
+performs better than GC in non-normalized networks. However, we show that PGT
+outperforms both in such networks.
+
+Knowledge distillation \cite{hinton2015distilling} is a process in which two networks
+are trained with hard and soft labels alternatively. Variants of knowledge distillation
+include self-distillation \cite{zhang2018deep}, identical student network distillation
+\cite{furlanello2018born}, channel distillation \cite{ge2019distilling}, regularizing
+wrong predictions \cite{yun2020regularizing} and representation or embedding based
+knowledge distillation \cite{yao2018deep}\cite{aguilar2020knowledge}. The key difference between PGT and knowledge
+distillation is that in the latter, the transformation is applied in the forward pass,
+while PGT is a backward pass modification only. Also, in distillation settings the
+temperature parameter is a part of the network's computation graph. In the case of
+PowerGrad Transform, we directly tamper the gradients without introducing any change in
+the forward pass. PGT differs from self-knowledge distillation as it neither introduces
+any additional sub-modules nor creates different ensembles to improve the performance of
+the model. PGT follows the standard neural network training mechanism with modified
+gradients.
 
@@
 \label{fig:prob_plots}
+\vspace{-0.5cm}
 \end{figure}
 
-The PowerGrad Transform technique is described in this section. It is a general
-technique and can be applied to any neural network. A neural network with parameters $W$
-generates $C$ logits denoted by $z$ for every input vector $x$. $z$ is given as $z=Wx$.
-Then a set of probability values $p_i$ are generated from the logits using a softmax
-function which is defined as $p_i=\frac{e^{z_i}}{\sum _{j=1}^C e^{z_j}}$. $p_i$ and
-$z_i$ represent the predicted probability values and the logits for the $i^{th}$ class
-respectively. Following this step, the loss function is invoked and the loss between the
-predicted probability values the ground truth labels (which is also a probability
-distribution) is calculated. If the loss function is cross-entropy loss, then the value
-of the loss is given as $L=-\sum _{i=1}^C q_i \log \left(p_i\right)$ where $q_i$ is the
-ground truth label of the $i^{th}$ class for a particular training example. By standard
-gradient update rule, we can calculate the gradient of the loss with respect to the
-logits which takes the form $\frac{\partial L}{\partial z_i}=p_i-q_i$.
+A neural network with parameters $W$ generates $C$ logits denoted by $z$ for every input
+vector $x$. $z$ is given as $z=Wx$. Then a set of probability values $p_i$ are generated
+from the logits using a softmax function which is defined as $p_i=\frac{e^{z_i}}{\sum
+_{j=1}^C e^{z_j}}$. $p_i$ and $z_i$ represent the predicted probability values and the
+logits for the $i^{th}$ class respectively. Following this step, the loss function is
+invoked and the loss between the predicted probability values the ground truth labels
+(which is also a probability distribution) is calculated. If the loss function is
+cross-entropy loss, then the value of the loss is given as $L=-\sum _{i=1}^C q_i \log
+\left(p_i\right)$ where $q_i$ is the ground truth label of the $i^{th}$ class for a
+particular training example. By standard gradient update rule, we can calculate the
+gradient of the loss with respect to the logits which takes the form $\frac{\partial
+L}{\partial z_i}=p_i-q_i$.
 
@@
 
+\vspace{-0.5cm}
 \begin{equation} \hspace*{3.5cm} p_i'=\frac{p_i^{\alpha }}{\sum _{j=1}^C
@@
 
+\vspace{-0.5cm}
 \begin{equation} \widehat{\frac{\partial L}{\partial z_i}}=p_i'-q_i
@@
 
-A code snippet of the algorithm implemented using PyTorch \cite{NEURIPS2019_9015} is
-given in the appendix. In PyTorch, the available method to modify gradients in the
-backward pass is using the \codeword{register_hook} function which is called in the
-\codeword{forward} function. We explore the effect of this change from a theoretical
-standpoint in section \ref{sec:Math}. In section \ref{sec:Expe}, we offer empirical data
-and experiments to support the proposed method.
-
-\section{Analysis of the PowerGrad transformation}
-\label{sec:Math}
-
-In this section, we first describe the properties of the  PowerGrad Transform (PGT) and
-then we highlight how these suitable properties of PGT helps the neural network to
-improve the performance. We use the same setup as described in section \ref{sec:Powe}.
-To explore the properties of PGT, we start by investigating the effect of the transform
-on the softmax probabilities.
-
-\subsection{Properties}
+\subsection{Properties of the PowerGrad transformation}
 \label{sec:pgt_prop}
+
+We use the same setup as described in section \ref{sec:Powe}. To explore the properties
+of PGT, we start by investigating the effect of the transform on the softmax
+probabilities.
 
@@
 
+\vspace{-0.5cm}
 \begin{equation} \begin{split} p_i' \geq p_i & \text{,\ \ if } p_i \leq
@@
 
+\vspace{-0.2cm}
 We call this threshold, the \textit{stationary threshold}. The stationary threshold is
@@
 
-We conclude from the analysis that the stationary threshold is a monotonic
-non-decreasing function with respect to $\alpha$. Also the derivative of PGT function
-with respect to the true probabilities is non-negative which in turn means that the
-transformation is an order-preserving map. All values greater than the threshold move
-towards the threshold after transformation and all values below the threshold also move
-towards the threshold, and the threshold itself moves monotonically towards $1/C$ as
-$\alpha$ is decreased from $1$ to $0$. This concludes that the transformation smooths
-the original distribution.
+We provide the proofs in the supplementary section and conclude that the stationary
+threshold is a monotonic non-decreasing function with respect to $\alpha$. Also the
+derivative of PGT function with respect to the true probabilities is non-negative which
+in turn means that the transformation is an order-preserving map. All values greater
+than the threshold move towards the threshold after transformation and all values below
+the threshold also move towards the threshold, and the threshold itself moves
+monotonically towards $1/C$ as $\alpha$ is decreased from $1$ to $0$. This concludes
+that the transformation smooths the original distribution.
 
 \subsection{Effects of PGT on logit, weight gradients and class separation}
+\label{sec:pgt_effects}
 
@@
 
-\subsection{Experiments on convolutional neural networks (with batch-normalization)}
-\label{sec:bn}
-
 \begin{figure}[!t]
@@
 \end{subfigure}
-\caption{ Log-linear plots of training and test accuracies and comparison with baseline
-of batch-normalized variants: \textbf{(a)} ResNet-18 ($\alpha=0.25$), \textbf{(b)}
-ResNet-50 ($\alpha=0.3$). For better visualization of accuracies obtained in the final
-iterations, we plot both axes in log scale. PowerGrad Transform improves upon the
-baseline in all cases. The corresponding accuracies are mentioned in Table
-\ref{tab:imagenet_table}. \textbf{(c)} Training speed comparison between PGT ($60$
+\caption{ Log-log plots of training and test accuracies and comparison with baseline of
+batch-normalized variants: \textbf{(a)} ResNet-18 ($\alpha=0.25$), \textbf{(b)}
+ResNet-50 ($\alpha=0.3$). \textbf{(c)} Training speed comparison between PGT ($60$
 epochs) and baseline ($100$ epochs). They both converge to the same test accuracy
@@
 \label{fig:metrics}
+\vspace{-0.5cm}
 \end{figure}
 
+We perform experiments on different variants ResNets using the ImageNet-1K dataset
+\cite{deng2009imagenet}. All models are trained on four V100 GPUs with a batch size of
+$1024$. We utilize a common set of hyperparameters for all experiments, which are as
+follows: 100 epoch budget, 5 epochs linear warmup phase beginning with a learning rate
+of $4\times 10^{-4}$ and ending with a peak learning rate of $0.4$, a momentum of $0.9$
+and weight decay of $5\times 10^{-4}$, the SGD Nesterov optimizer and mixed precision.
+In our studies, we employ either a step scheduler (dividing the learning rate by $10$ at
+the $30^{th}$, $60^{th}$, and $90^{th}$ epochs) or a cosine decay scheduler
+\cite{loshchilov2016sgdr}. We find $\alpha=0.25$ and $\alpha=0.05$ to be good choices
+for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$ also have good
+performance as well. The experimental results are shown in Table
+\ref{tab:imagenet_table}, and we mention the value of the PGT hyperparmeter ($\alpha$)
+in each experiment. We explore different values of $\alpha$ in section \ref{sec:Abla}
+and provide detailed grid plots in the supplementary section\footnote{Reproducible code,
+training recipes for the experiments, pretrained checkpoints and training logs are
+provided at: \url{ https://github.com/skalien/power-grad-transform }.}.
+
 \begin{table}[!t]
 \centering
-\caption{ Results and comparison table for CNN networks trained on ImageNet-1K (sorted
-in descending order w.r.t. the last column denoting test accuracy improvements over
-baseline). Best training and test accuracies are highlighted in red and blue
-respectively. Accuracy differences are highlighted in yellow. In all experiments, we
-observe that PGT positively improves both training and test accuracies. We also observe
-that deeper networks such as SE-ResNet-50 and ResNet-50 accrue higher gains over the
-baselines as compared to shallow networks like ResNet-18. }
+\caption{ Results and comparison table for networks trained on ImageNet-1K. Best
+training and test accuracies are highlighted in red and blue respectively. Accuracy
+differences are highlighted in yellow. }
 \label{tab:imagenet_table}
@@
 \end{tabular}
+\vspace{-0.5cm}
 \end{table}
-
-We perform experiments on different variants ResNets using the ImageNet-1K dataset
-\cite{deng2009imagenet}. All models are trained on four V100 GPUs with a batch size of
-$1024$. We utilize a common set of hyperparameters for all experiments, which are as
-follows: 100 epoch budget, 5 epochs linear warmup phase beginning with a learning rate
-of $4\times 10^{-4}$ and ending with a peak learning rate of $0.4$, a momentum of $0.9$
-and weight decay of $5\times 10^{-4}$, the SGD Nesterov optimizer and mixed precision.
-In our studies, we employ either a step scheduler (dividing the learning rate by $10$ at
-the $30^{th}$, $60^{th}$, and $90^{th}$ epochs) or a cosine decay scheduler
-\cite{loshchilov2016sgdr}. We find $\alpha=0.25$ and $\alpha=0.05$ to be good choices
-for ResNet-18 and ResNet-50, though larger values such as $\alpha=0.3$ also have good
-performance as well. The experimental results are shown in Table
-\ref{tab:imagenet_table}, and we mention the value of the PGT hyperparmeter ($\alpha$)
-in each experiment. We explore different values of $\alpha$ in section \ref{sec:Abla}
-and provide detailed grid plots in the supplementary section.
 
@@
 SE-ResNet-18 and ResNet-18 sees accuracy boosts which are smaller but nevertheless
-positive. Therefore we conclude that with PGT, the consistent improvements in training
-accuracies across all cases is because the networks train better and arrive at better
-optimas during convergence. Per epoch training and test accuracy plots of ResNet-18 and
-ResNet-50 (both with and without PGT) are shown in Fig. \ref{fig:metrics}(a, b). For
-better visualization of the top end of the plots, we scale both axes in log scale.
-Because both training and test accuracies improve, it indicates that PowerGrad Transform
-allows the network to learn better representations\footnote{Reproducible code, training
-recipes for the experiments, pretrained checkpoints and training logs are provided at:
-\url{ https://github.com/skalien/power-grad-transform }. Code is adapted from
-\cite{rw2019timm}.}.
-
-These improvements are obtained with the number of epochs being kept the same as that of
-the baseline model. Alternatively, practitioners can apply PGT and have the networks
-converge to accuracy levels comparable to the baseline with a significantly reduced
-epoch budget. In Fig. \ref{fig:metrics}(c), we illustrate one such instance in which we
-train a ResNet-50 to its cosine scheduler baseline in just $60$ epochs with PGT, and
-match its baseline performance that is acquired in $100$ epochs. This hints to the
-observation that PGT not only makes the network perform better but also speeds up
-training. Because of the acceleration in training that is made possible by PGT, the
-total amount of training time can be decreased by $40\%$.
+positive. With consistent improvements in training accuracies across all cases, we
+conclude PGT helps networks learn better representations and arrive at better optimas
+during convergence. Per epoch training and test accuracy plots of ResNet-18 and
+ResNet-50 (both with and without PGT) are shown in Fig. \ref{fig:metrics}(a, b).
+Practioners can also choose to accelerate training and save as much as $40\%$ of the
+epoch budget [Fig. \ref{fig:metrics}(c)].
 
@@
 
-We examine issues that occur in non-BatchNormalized networks and we provide insight into
-the effects of PGT on such networks. We use ResNet-18 \cite{he2016deep} as the
-foundation model trained on ImageNet-1K \cite{russakovsky2015imagenet}. We found deeper
-networks (such as ResNet-34 and ResNet-50 models) are impossible to train without Batch
-Normalization because of increased depth. So in order to conduct our empirical
-experiments, we solely focus on the ResNet-18 architecture. We designate different
-layers with their corresponding layer indices. We provide a detailed layer-wise index
-list of ResNet-18 in supplementary section \ref{sec:pgt_class_separation}.
-
-Throughout the training process, we monitor variations in the the per-filter L2-norm of
-each layer's weight tensor. Because a full plot of each layer's data would use an
-inordinate amount of space, we focus on a few critical layers' plots. The supplementary
-sections (\ref{sec:plots1} through \ref{sec:plots4}) contains full plots for all layer
-for various settings. The norm of the weights of the $11^\textit{th}$ convolutional
-layer is shown in Fig. \ref{fig:norm_plots}(a). Layer $11$ includes $256$ filters. Each
-colour on the figure represents the iteration-wise evolution of a particular filter
-throughout the course of training. The critical point to note is that some filters
-achieve a norm of zero during training. We refer to this event as `Zeroing Out', and
-occurs when a channel of a weight tensor gets fully filled with zeros. If all the
-coefficients of the filter becomes zero, only then the L2 norm of that filter becomes
-zero. If we observe zero filter norm for a filter, then that filter does not contribute
-at all to determine the input-output relationship of a dataset, as the feature tensor it
-produces is also an all-zero tensor. According to our findings, when a filter tensor
-becomes zeroed out, it does not recover with further training. We observe both effects
-\textbf{1)} zeroed out filter \textbf{2)} zeroed out feature.
+We examine issues that occur in non-normalized networks (networks without BN layers). We
+use ResNet-18 \cite{he2016deep} as the foundation model trained on ImageNet-1K
+\cite{russakovsky2015imagenet}. Deeper networks such as ResNet-34 and ResNet-50 are
+impossible to train without Batch Normalization courtesy of the increased depth and so
+we solely focus on ResNet-18. We designate different layers with their corresponding
+layer indices. Throughout the training process, we monitor variations in the the
+per-filter L2-norm of each layer's weights (layer indices are mentioned in section
+\ref{sec:r18_layers}). In Fig. \ref{fig:norm_plots}(a), some filters of layer $11$
+achieve a norm of zero during training. We refer to this event as `Zeroing Out', and it
+occurs when one of the channels (or filters) of a weight tensor gets fully filled with
+zeros and such filters do not contribute at all to determine the input-output
+relationship of a dataset, as the feature tensor it produces is also an all-zero tensor.
+According to our findings, when a filter once zeroed out does not recover with further
+training.
 
@@
 \centering
+\captionsetup{font=footnotesize}
+
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-4-2}
+\caption{Layer 11\\ \textbf{(baseline)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-4-2}
+\caption{Layer 11\\ \textbf{(PGT)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-7-2}
+\caption{Final layer\\ \textbf{(baseline)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-7-2}
+\caption{Final layer\\ \textbf{(PGT)}
+}\end{subfigure}
 \captionsetup{font=scriptsize}
-
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{baseline-w-layer-4-2}
-\caption{Norm of filters of convolutional layer 11 (without PGT)}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-f-layer-22-1}
+\caption{Norm of final layer's features\\ \textbf{(baseline)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-f-layer-22-1}
+\caption{Norm of final layer's features\\ \textbf{(PGT)}
+}\end{subfigure}
+
+\vspace{6pt}
+
+\captionsetup{font=footnotesize}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-5-3}
+\caption{Layer 15\\ \textbf{(baseline)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-5-3}
+\caption{Layer 15\\ \textbf{(PGT)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-5-3}
+\caption{Layer 15\\ \textbf{(AGC+PGT)}}
 \end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{baseline-w-layer-7-2}
-\caption{Norm of filters of the final convolutional layer (without PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{baseline-f-layer-22-1}
-\caption{Norm of final layer's output features (without PGT)}
-\end{subfigure}
-
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{pgt-w-layer-4-2}
-\caption{Norm of filters of convolutional layer 11 (with PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{pgt-w-layer-7-2}
-\caption{Norm of filters of the final convolutional layer (with PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{pgt-f-layer-22-1}
-\caption{Norm of final layer's output features (with PGT)}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-1-2}
+\caption{Layer 2\\ \textbf{(baseline)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-1-2}
+\caption{Layer 2\\ \textbf{(PGT)}
+}\end{subfigure}
+\begin{subfigure}[t]{0.16\textwidth}
+\includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-1-2}
+\caption{Layer 2\\ \textbf{(AGC+PGT)}}
 \end{subfigure}
 \captionsetup{font=normalsize}
-\caption{ Experiment demonstrating the Zeroing Out phenomena, as observed in different
-filters and feature vectors. PGT greatly reduces the zeroing out of filters and promote
-better learning. }
+\caption{ Norm vs iter. plots demonstrating layer characteristics (the zero out
+phenomena) and the efficacy of PGT and AGC+PGT over baseline. Each colour represents a
+different filter or feature vector of a particular layer of a non-BN variant of
+ResNet-18. }
 \label{fig:norm_plots}
+\vspace{-0.5cm}
 \end{figure}
 
-Similar to the plot \ref{fig:norm_plots}(a), we see the plot of the final convolutional
-layer in Fig. \ref{fig:norm_plots}(b), where we observe a number of filters / units
-completely dropping out towards the end of training leading to underutilization of the
-network's parameters. In Fig. \ref{fig:norm_plots}(c), we see the norm of the feature
-vector at the output side of the global average pooling layer of ResNet-18. The feature
-vector here directly interfaces with the fully connected layer immediately after. Any
-zeroed out regions in this tensor directly leads to permanent information loss, as it
-does not contribute to the learning of decision boundaries in the upcoming fully
-connected layer, and indeed there exists a few zeroed out features in the baseline
-model, which we discuss next. The impact of having a zeroed out filter or a feature
-might have a variety of consequences for training and optimization. If a zeroed out
-filter is present in a convolutional layer, there are two possible outcomes. If the
-layer is immediately followed by a fully connected regression layer, there is an abrupt
-loss of information. For the reason that a zeroed out filter always generates zeroed out
-features regardless of the input image, when this tensor is fed to a fully connected
-layer, it does not communicate any information that is useful for classification. In the
-event that the zeroed out filter has a skip connection linked across it, this may be
-circumvented as long as there are no overlapping zeroed out channels in the two feature
-maps. On the other hand, if a feature tensor has a zeroed out channel at the input of a
-layer, the layer creates a zeroed out feature channel at the output, regardless of the
-weights. However, if there is a zeroed out channel in the feature map just before the
-fully connected layer, then this results in a permanent loss of information as it does
-not contribute to the learning of decision boundaries. ResNet-18 suffers from the same
-kind of information loss, as the final feature tensor (which is generated after the
-global average pooling layer) has multiple channels that are zeroed out [Fig.
-\ref{fig:norm_plots}(c)]. This information collapse phenomenon occurs when a channel
-that has been zeroed out stays zeroed out for all images in the dataset.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.45\columnwidth]{baseline_high_bs-w-layer2-2}
-\caption{ Per filter norm vs. iteration plot of layer 5 of an unnormalized ResNet-18,
-trained with a very high batch size of 1024. The weight matrix of the entire layer
-zeroes out during training. }
+Fig. \ref{fig:norm_plots}(c) and \ref{fig:norm_plots}(e) are the plot of the final conv
+layer's filters and the final global average pooling layer respectively. We observe a
+number of filters completely zeroing out. The feature vector here directly interfaces
+with the fully connected layer immediately after. Therefore any zeroed out features
+leads to permanent information loss, as it does not contribute to the learning of
+decision boundaries in the fully connected layer. Also, zeroed out weights tensors lead
+to zeroed out gradients hence stopping training for all subsequent iterations leading to
+a collapse in training for the affected layers. With large batch sizes, it is possible
+that an entire layer zeroes out [Fig. \ref{fig:high_bs}]. Gradient modification methods
+such as PGT or AGC can alleviate the zeroing out phenomena. We observe that the number
+of zeroed out filters has considerably reduced [Fig. \ref{fig:norm_plots}(b, d)]. The
+final feature tensor [Fig. \ref{fig:norm_plots}(f)] with PGT enabled, does not contain
+any zeroed out regions indicating that information loss is mitigated as the features
+pass on from the feature extracting layers to the fully connected layer. When PGT and
+AGC are combined, we get the most benefit for layer $15$ [Fig \ref{fig:norm_plots}(g-i)]
+and layer $2$ [Fig \ref{fig:norm_plots}(j-l)].
+
+In Table \ref{tab:wobn_table} we find the baseline performance for high batch sizes
+($1024$) is drastically inferior to baselines for other batch sizes. PGT helps regain
+some of the lost performance by $0.682\%$ ($65.498\%$ vs. $64.816\%$). At a batch size
+of 512, invoking PGT improves the training accuracy baseline by $1.48\%$ and the test
+accuracy baseline by $0.684\%$, while at a batch size of 256, the improvement in
+training and test accuracies are $1.11\%$ and $1.018\%$ respectively. In comparison, the
+test accuracy improvements obtained by GC and AGC at batch size of 256
+is much less at $0.27\%$ and $0.5\%$, respectively. On the training accuracy front,
+since we get a significant boost ($1.48\%$ at batch size of 512 and $1.11\%$ at batch
+size of 256), it leads us to infer that when PowerGrad Transform is used, the network
+
+\begin{wrapfigure}{l}{0.4\textwidth}
+\vspace{-1cm}
+\centering
+\includegraphics[width=0.45\textwidth]{baseline_high_bs-w-layer2-2}
+\caption{}
 \label{fig:high_bs}
-\end{figure}
-
-The impact of a feature tensor or weight tensor being zeroed out is seen to be
-particularly common in training procedures involving larger batch sizes in networks
-without normalization layers. As a result, the network's parameters are underutilized
-since some of the filters are not involved in extracting any useful information that is
-important for performing classification. When working with large batch sizes, it is
-possible that an entire layer zeroes out [Fig. \ref{fig:high_bs}]. In a network without
-skip connections, this can cause a full collapse in training. For example, if a weight
-tensor completely zeroes out, and it always produces a zero feature tensor, then the
-input feature tensor to the next convolutional layer is a zero tensor and thus the
-output of the next layer is always a zero tensor irrespective of its weights, and this
-continues until the end of the neural network pipeline leading to zero logits. Also,
-zeroed out weights tensors lead to zeroed out gradients hence stopping training for all
-subsequent iterations leading to a collapse in training. Although residual networks
-bypass this difficulty via the use of the skip connection, it does not solve the layer's
-non-participation in the overall objective of classification. It is possible to solve
-this issue by using methods such as gradient clipping (GC), adaptive gradient clipping
-(AGC), and PowerGrad Transform (PGT), all of which regulate the amount of gradient that
-flows at various junctions of the network. GC and AGC modify the gradients at each and
-every filter of every layer. PGT performs this operation at only one junction, which is
-at the softmax layer.
-
-\begin{figure}[t]
-\centering
-\captionsetup{font=scriptsize}
-
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{baseline-w-layer-5-3}
-\caption{Norm of filters of layer 15 (without AGC/PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{pgt-w-layer-5-3}
-\caption{Norm of filters of layer 15 (with PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{agc_pgt-w-layer-5-3}
-\caption{Norm of filters of layer 15 (with AGC+PGT)}
-\end{subfigure}
-
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{baseline-w-layer-1-2}
-\caption{Norm of filters of layer 2 (without AGC/PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{pgt-w-layer-1-2}
-\caption{Norm of filters of layer 2 (with PGT)}
-\end{subfigure}
-\begin{subfigure}[t]{0.32\textwidth}
-\includegraphics[width=\textwidth]{agc_pgt-w-layer-1-2}
-\caption{Norm of filters of layer 2 (with AGC+PGT)}
-\end{subfigure}
-\captionsetup{font=normalsize}
-\caption{ Experiment demonstrating the efficacy of PGT and AGC+PGT, as observed in
-different layers. }
-\label{fig:agc_pgt}
-\end{figure}
-
-In a similar fashion to the study in section \ref{sec:Empi}, we examine the norm
-profiles of features and weights in the PGT and AGC. The supplementary section includes
-all the detailed graphs of all layer. In Fig. \ref{fig:norm_plots}(d), we show the
-per-filter norm plot of the $11^\textit{th}$ convolutional layer as it progresses
-through a PGT enabled training session. We observe that the number of zeroed out filters
-has considerably reduced. We observe similar benefits in plot of Fig.
-\ref{fig:norm_plots}(e), which is the plot of the per-filter norm of the final
-convolutional layer. Coming to benefits in feature norms, we observe in plot of Fig.
-\ref{fig:norm_plots}(f), the final feature tensor obtained with PGT enabled training do
-not contain any zeroed out regions, leading us to assert that information loss is
-heavily mitigated as the features pass on from the feature extracting convolutional
-layers to the fully connected stage for regression.
-
-AGC is excellent at eliminating the issue of filter and feature zeroing out. However,
-AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
-it also mitigates the zeroing out phenomenon in most layers. In Fig.
-\ref{fig:agc_pgt}(a), we see a plot of the $15^\textit{th}$ layer of the non-BN
-ResNet-18 trained without PGT. As we can see that there has been quite a few filter
-dropouts in the baseline. PGT recovers most of the filters [Fig. \ref{fig:agc_pgt}(b)].
-In some layers however, the performance of PGT is slightly weaker than the baseline. The
-$2^\textit{nd}$ convolutional layer has more zeroed out filters in PGT [Fig.
-\ref{fig:agc_pgt}(e)] than in the baseline [Fig. \ref{fig:agc_pgt}(d)]. Despite that,
-PGT achieves a higher overall accuracy (Table \ref{tab:wobn_table}). This necessitates
-the simultaneous use of AGC and PGT, for the non-BN ResNet-18.
+\vspace{-2cm}
+\end{wrapfigure}
+
+\noindent
+fits the training dataset more tightly and the convergence optima is significantly
+superior. Combining GC with PGT is not very significant. When AGC and PGT are combined,
+we see a tremendous increase in test accuracy of over $2.06\%$ over the baseline.
+Additionally, we provide comprehensive graphs of the norm profiles of the AGC + PGT
+method for all layers in the supplementary section \ref{sec:plots1}\ -\
+\ref{sec:plots4}.
 
@@
 \centering
-\caption{ Results for non-normalized ResNets on ImageNet-1K. Each experiment is
-conducted with a 100 epoch budget using the cosine decay scheduler. Best training and
-test accuracies are highlighted in red and blue respectively. Top differences in
-training and test accuracies are marked in yellow. The baselines in each segment is
-taken as the reference point for all `Diff' accuracies in that segment. The first three
-segments highlight comparison for experiments where the methods are applied
-individually, while the last segment denotes performance metrics for combinations of
-different methods (GC + PGT, AGC + PGT). We see that at any batch size, with PGT, the
-performance of non-normalized version of ResNet-18 improves significantly over the
-baseline owing to the improvements in training as observed in Fig. \ref{fig:agc_pgt}. }
+\caption{ Results for non-normalized ResNets on ImageNet-1K. Best training and test
+accuracies are highlighted in red and blue respectively. Top differences in training and
+test accuracies are marked in yellow. }
 \label{tab:wobn_table}
@@
 \end{tabular}
+\vspace{-0.5cm}
 \end{table}
-
-We provide performance metrics on the test set of ImageNet-1K for different batch sizes
-for the non-BN ResNet-18 and also experiment with different methods (GC, AGC, PGT along
-with their combinations). Table \ref{tab:wobn_table} shows the findings. To begin, we
-use a batch size of 1024. The baseline performance for this batch size is drastically
-inferior to baselines for other batch sizes. PGT helps regain some of the lost
-performance by $0.682\%$ ($65.498\%$ vs. $64.816\%$). This illustrates that in
-unnormalized networks, accuracy suffers significantly when batch sizes are large.
-Additionally, the training process is often unstable at this batch size, resulting in
-filters and sometimes entire layers frequently zeroing out (as seen in section
-\ref{sec:Empi}). At a batch size of 512, invoking PGT improves the training accuracy
-baseline by $1.48\%$ and the test accuracy baseline by $0.684\%$, while at a batch size
-of 256, the improvement in training and test accuracies are $1.11\%$ and $1.018\%$
-respectively. In comparison, the test accuracy improvements obtained by GC and AGC at
-batch size of 256 is much less, at $0.27\%$ and $0.5\%$, respectively.
-
-On the training accuracy front, since we get a significant boost ($1.48\%$ at batch size
-of 512 and $1.11\%$ at batch size of 256), it leads us to infer that when PowerGrad
-Transform is used, the network fits the training dataset more tightly and the
-convergence optima is significantly superior. We also notice that the training versus
-testing gap increases with PGT as compared to the gap at baseline even though both
-training and test accuracies improve. Therefore it allows us to conclude that the
-improvements in training and test accuracies are obtained not through regularization
-like it does in methods such as label smoothing; rather the improvements are obtained as
-PGT enables the network to utilize its learning capacity better and arrive at a higher
-performing optima at convergence, something which is inaccessible through traditional
-gradient based training.
-
-We also experiment with combinations of GC/AGC and PGT. The advantage of combining GC
-with PGT is not very significant. When AGC and PGT are combined, we see a tremendous
-increase in test accuracy of over $2.06\%$ over the baseline. As seen in Fig.
-\ref{fig:agc_pgt}(c) and \ref{fig:agc_pgt}(f), activating AGC and PGT concurrently
-results in no filter dropouts in all layers while also improving the accuracy by a
-significant margin. Additionally, we provide comprehensive graphs of the norm profiles
-of the AGC + PGT method for all layers in the supplementary section.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[!bp]
 \centering
@@
 As seen in Fig. \ref{fig:stats}(a), the logit norm increases as $\alpha$ falls from $1$
-to $0$. In the initial training phase, PGT causes the gradient in both the valid and
-wrong classes to rise as the network misclassifies majority of the examples. Due to the
-exponential nature of the softmax function, the predicted probability distribution may
-frequently be close to a delta distribution, even in instances of misclassifications.
-PGT reshapes the predicted probability distribution. It redistributes probability from
-the confident classes to the less confident ones. Gradient rises in all classes. We also
-see that the final value of the loss rises as $\alpha$ drops, and that the logit norm
-and the final value of the loss are linearly correlated Fig. \ref{fig:stats}(c). What is
-surprising is that: \textit{it is possible to achieve higher accuracies even though the
-loss values are larger during convergence}. In our studies with ResNet-50, for instance,
-if PGT is not invoked, the corresponding test accuracy is $76.56\%$ (Table
-\ref{tab:imagenet_table}) with a cross-entropy loss value of $0.852$. However, if PGT is
-activated with $\alpha=0.05$, the corresponding test accuracy is $77.216\%$ (Table
-\ref{tab:imagenet_table}) with a loss value of $2.360$. This is markedly different from
-the coupled gradient descent based training procedures. PGT enables the network to
-arrive at such optima where the loss values are high, but both training and test
-perfomance is better. From these plots of logits and losses we find that with the
-inclusion of PGT in the training process, the model can have access to such regions of
-the loss landscape that is otherwise inaccessible to traditional gradient based training
-procedures.
+to $0$. We also see that the final value of the loss rises as $\alpha$ drops, and that
+the logit norm and the final value of the loss are linearly correlated Fig.
+\ref{fig:stats}(c). What is surprising is that: \textit{it is possible to achieve higher
+accuracies even though the loss values are larger during convergence}. In our studies
+with ResNet-50, for instance, if PGT is not invoked, the corresponding test accuracy is
+$76.56\%$ [Table \ref{tab:imagenet_table}] with a cross-entropy loss value of $0.852$.
+However, if PGT is activated with $\alpha=0.05$, the corresponding test accuracy is
+$77.216\%$ [Table \ref{tab:imagenet_table}] with a loss value of $2.360$. This is
+markedly different from the coupled gradient descent based training procedures. PGT
+enables the network to arrive at such optima where the loss values are high, but both
+training and test perfomance is better. From these plots of logits and losses we find
+that with the inclusion of PGT in the training process, the model can have access to
+such regions of the loss landscape that is otherwise inaccessible to traditional
+gradient based training procedures.
 
@@
 \end{tabular}
+\vspace{-0.5cm}
 \end{table}
 
-We conduct an ablation study to investigate the effects of PowerGrad Transform for
-different values of the hyperparameter ($\alpha$). Grid plots are made available in the
-supplementary section. We have previously demonstrated from the experiments depicted in
-section \ref{sec:Expe}, that PowerGrad Transform causes the network to arrive at a
-better optima and fit the training dataset better than the baseline. In this ablation
-study, we use the ResNet-50 architecture and we combine our proposed method with other
-regularization techniques such as label smoothing and report our findings in Table
-\ref{tab:ablation_table}. We also investigate the two optimal peaks of the
-hyperparameter ($\alpha$) obtained from ResNet-50's grid search ($\alpha=0.05$ and
-$\alpha=0.3$). First, we examine the effect of PGT on the step scheduler baseline in
-order to later compare it to the cosine scheduler baseline. \textbf{Row-1)} We begin
-with the step scheduler baseline ($75.97\%$). \textbf{Row-2)} PGT improves upon the step
-scheduler baseline (test set) by a substantial margin with $0.524\%$ ($76.494\%$ as
-opposed to $75.97\%$). \textbf{Row-3)} Introducing the cosine scheduler yields a
-$0.59\%$ improvement ($76.56\%$ vs. $75.97\%$) over the step scheduler. \textbf{Row-4)}
-After introducing label smoothing, the test accuracy relative to the cosine scheduler
-baseline increases by only $0.138\%$ (from $76.56\%$ to $76.698\%$). \textbf{Row-5)}
-However, introducing PGT with $\alpha=0.3$ alone (without label smoothing) improves the
-cosine scheduler baseline by $0.326\%$ ($76.886\%$ vs. $76.56\%$). \textbf{Row-6)}
-Combining PGT ($\alpha=0.3$) with label smoothing improves the performance on the test
-set further by $0.408\%$ (from $76.56\%$ to $76.968\%$) and reduces the generalization
-gap (from $2.54\%$ to $1.5\%$). However, the impact of combining PGT with label
-smoothing can vary depending on the value of the hyperparameter ($\alpha$).
-\textbf{Row-7)} With a PGT hyperparameter value of $\alpha=0.05$, we notice the greatest
-performance improvement, $1.246\%$ over the step scheduler test baseline and $0.656\%$
-over the cosine scheduler test baseline. \textbf{Row-8)} Adding label smoothing to PGT
-($\alpha=0.05$) hurts performance even though it reduces the generalization gap.
-
-\section{Discussions}
-\label{sec:Disc}
-
-We introduced PowerGrad Transform, which decouples the backward and forward passes of
-neural network training and enables a significantly better fit to the dataset, as
-measured by training and test accuracy metrics. We show the application of PowerGrad
-Transform, a simple yet powerful technique for modifying the gradient flow behaviour. We
-investigate experimentally the degenerate behavior of non-BN networks, which is
-frequently observed during the standard training approach, especially with higher batch
-sizes. With PGT, gradient behavior is enhanced and the likelihood of weights attaining
-degenerate states is diminished. We provide a theoretical analysis of the PowerGrad
-transformation. With the use of different network topologies and datasets, we are able
-to show the potential of PGT and explore its impacts from an empirical standpoint. We
-provide comprehensive results on a number of models (non-BN and BN ResNets, SE-ResNets)
-using the ImageNet dataset. PGT helps the network to improve its learning capabilities
-by locating a more optimum convergence point and simultaneously speeds up training. In
-addition, we undertake an ablation research and compare its effects to those of
-regularization methods such label smoothing.
-
-% \end{document}
+We conduct ablation studies to investigate the effects of PowerGrad Transform for
+different values of the hyperparameter ($\alpha$), where we use the ResNet-50
+architecture and combine our proposed method with different schedulers, regularization
+techniques and different values of $\alpha$. We report our findings in Table
+\ref{tab:ablation_table}. First we examine the effect of PGT on the step scheduler
+baseline in order to later compare it to the cosine scheduler baseline. \textbf{Row-1)}
+We begin with the step scheduler baseline ($75.97\%$). \textbf{Row-2)} PGT improves upon
+the step scheduler baseline (test set) by a substantial margin with $0.524\%$
+($76.494\%$ as opposed to $75.97\%$). \textbf{Row-3)} Introducing the cosine scheduler
+yields a $0.59\%$ improvement ($76.56\%$ vs. $75.97\%$) over the step scheduler.
+\textbf{Row-4)} After introducing label smoothing, the test accuracy relative to the
+cosine scheduler baseline increases by only $0.138\%$ (from $76.56\%$ to $76.698\%$).
+\textbf{Row-5)} However, introducing PGT with $\alpha=0.3$ alone (without label
+smoothing) improves the cosine scheduler baseline by $0.326\%$ ($76.886\%$ vs.
+$76.56\%$). \textbf{Row-6)} Combining PGT ($\alpha=0.3$) with label smoothing improves
+the performance on the test set further by $0.408\%$ (from $76.56\%$ to $76.968\%$) and
+reduces the generalization gap (from $2.54\%$ to $1.5\%$). However, the impact of
+combining PGT with label smoothing can vary depending on the value of the hyperparameter
+($\alpha$). \textbf{Row-7)} With a PGT hyperparameter value of $\alpha=0.05$, we notice
+the greatest performance improvement, $1.246\%$ over the step scheduler test baseline
+and $0.656\%$ over the cosine scheduler test baseline. \textbf{Row-8)} Adding label
+smoothing to PGT ($\alpha=0.05$) hurts performance even though it reduces the
+generalization gap.
+
+\section{Conclusion}
+\label{sec:Conc}
+
+PowerGrad Transform enables a significantly better fit to the dataset as measured by
+training and test accuracy metrics. With PGT, gradient behavior is enhanced and weights
+attain better values in normalized networks and degenerate states are avoided in non-BN
+networks. We provide theoretical analyses of the transformation. With different network
+topologies and datasets, we are able to show the potential of PGT and explore its
+impacts from an empirical standpoint. PGT helps the network to improve its learning
+capabilities by locating a more optimum convergence point and simultaneously speeds up
+training.
 
@@
 \begin{figure}[ht] \hrule \lstinputlisting[language=Python]{code.py} \hrule
-\vspace{0.1cm} \caption{Python code of PowerGrad Transform based on PyTorch. }
+\vspace{0.25cm} \caption{Python code of PowerGrad Transform based on PyTorch. }
 \label{fig:code}
@@
 For the layer indices used in figures Fig. \ref{fig:norm_plots}, \ref{fig:high_bs},
-\ref{fig:agc_pgt}, we provide a block diagram depicting the different convolutional
+\ref{fig:norm_plots}, we provide a block diagram depicting the different convolutional
 layers of ResNet-18 with corresponding layer indices in Fig. \ref{fig:resnet18}. Each

### 2022-11-17 0100.tex
Mod3

@@
 \raggedbottom
+\newcommand{\forceindenta}{\parindent=1em\indent\parindent=0pt\relax}
+\newcommand{\forceindentb}{\parindent=1.65em\indent\parindent=0pt\relax}
 \renewcommand\UrlFont{\color{blue}\rmfamily}
@@
 
-\textbf{Type I:} Independent gradient modification at multiple points: In this method,
-the gradients are first computed using standard procedure and then individually altered
-as shown in Fig. \ref{fig:gradient_altering}(b). Gradient clipping
+\textbf{Type I:} Independent gradient modification at multiple points: Here the
+gradients are first computed using standard procedure and then individually altered as
+shown in Fig. \ref{fig:gradient_altering}(b). Gradient clipping
 \cite{pascanu2013difficulty} and Adaptive gradient clipping \cite{brock2021high} are
@@
 p_i' < p_i & \text{,\ \ if } p_i > \Big(\sum _{j=1}^C
-p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}} \end{split} \label{eqn:threshold}
+p_j^{\alpha}\Big)^{\frac{1}{\alpha-1}} \end{split} \label{eqn:threshold0}
 \end{equation}
@@
 \includegraphics[width=\textwidth]{trimmed/baseline-w-layer-4-2}
-\caption{Layer 11\\ \textbf{(baseline)}
-}\end{subfigure}
+\caption{Layer 11\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/pgt-w-layer-4-2}
-\caption{Layer 11\\ \textbf{(PGT)}
-}\end{subfigure}
+\caption{Layer 11\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/baseline-w-layer-7-2}
-\caption{Final layer\\ \textbf{(baseline)}
-}\end{subfigure}
+\caption{Layer 19\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/pgt-w-layer-7-2}
-\caption{Final layer\\ \textbf{(PGT)}
-}\end{subfigure}
-\captionsetup{font=scriptsize}
+\caption{Layer 19\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/baseline-f-layer-22-1}
-\caption{Norm of final layer's features\\ \textbf{(baseline)}
-}\end{subfigure}
+\caption{\scriptsize{GAP\\\forceindentb features}\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/pgt-f-layer-22-1}
-\caption{Norm of final layer's features\\ \textbf{(PGT)}
-}\end{subfigure}
+\caption{\scriptsize{GAP\\\forceindentb features}\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
 
@@
 \includegraphics[width=\textwidth]{trimmed/baseline-w-layer-5-3}
-\caption{Layer 15\\ \textbf{(baseline)}
-}\end{subfigure}
+\caption{Layer 15\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/pgt-w-layer-5-3}
-\caption{Layer 15\\ \textbf{(PGT)}
-}\end{subfigure}
+\caption{Layer 15\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-5-3}
-\caption{Layer 15\\ \textbf{(AGC+PGT)}}
+\caption{Layer 15\\ \forceindenta\textbf{AGC+PGT}}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{trimmed/baseline-w-layer-1-2}
-\caption{Layer 2\\ \textbf{(baseline)}
-}\end{subfigure}
+\caption{Layer 2\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/pgt-w-layer-1-2}
-\caption{Layer 2\\ \textbf{(PGT)}
-}\end{subfigure}
+\caption{Layer 2\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
 \begin{subfigure}[t]{0.16\textwidth}
 \includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-1-2}
-\caption{Layer 2\\ \textbf{(AGC+PGT)}}
+\caption{Layer 2\\ \forceindenta\textbf{AGC+PGT}}
 \end{subfigure}
@@
 Fig. \ref{fig:norm_plots}(c) and \ref{fig:norm_plots}(e) are the plot of the final conv
-layer's filters and the final global average pooling layer respectively. We observe a
-number of filters completely zeroing out. The feature vector here directly interfaces
-with the fully connected layer immediately after. Therefore any zeroed out features
-leads to permanent information loss, as it does not contribute to the learning of
-decision boundaries in the fully connected layer. Also, zeroed out weights tensors lead
-to zeroed out gradients hence stopping training for all subsequent iterations leading to
-a collapse in training for the affected layers. With large batch sizes, it is possible
-that an entire layer zeroes out [Fig. \ref{fig:high_bs}]. Gradient modification methods
-such as PGT or AGC can alleviate the zeroing out phenomena. We observe that the number
-of zeroed out filters has considerably reduced [Fig. \ref{fig:norm_plots}(b, d)]. The
-final feature tensor [Fig. \ref{fig:norm_plots}(f)] with PGT enabled, does not contain
-any zeroed out regions indicating that information loss is mitigated as the features
-pass on from the feature extracting layers to the fully connected layer. When PGT and
-AGC are combined, we get the most benefit for layer $15$ [Fig \ref{fig:norm_plots}(g-i)]
-and layer $2$ [Fig \ref{fig:norm_plots}(j-l)].
+layer's filters (layer $19$) and the features output of global average pooling (GAP)
+layer respectively. We observe a number of filters in the final layer completely zeroing
+out. The feature vector after GAP layer directly interfaces with the fully connected
+layer immediately after. Therefore any zeroed out features leads to permanent
+information loss, as it does not contribute to the learning of decision boundaries in
+the fully connected layer. Also, zeroed out weights tensors lead to zeroed out gradients
+hence stopping training for all subsequent iterations leading to a collapse in training
+for the affected layers. With large batch sizes, it is possible that an entire layer
+zeroes out [Fig. \ref{fig:high_bs}]. Gradient modification methods such as PGT or AGC
+can alleviate the zeroing out phenomena. We observe that the number of zeroed out
+filters has considerably reduced [Fig. \ref{fig:norm_plots}(b, d)]. The final feature
+tensor [Fig. \ref{fig:norm_plots}(f)] with PGT enabled, does not contain any zeroed out
+regions indicating that information loss is mitigated as the features pass on from the
+feature extracting layers to the fully connected layer. When PGT and AGC are combined,
+we get the most benefit for layer $15$ [Fig \ref{fig:norm_plots}(g-i)] and layer $2$
+[Fig \ref{fig:norm_plots}(j-l)].
 
@@
 
-\clearpage
-
 \subsection{PGT restricts the partial derivative of the loss w.r.t. each logit from becoming too small or too high}
@@
 
+\clearpage
+
 \subsection{Sample Code}
@@
 ImageNet dataset. Each datapoint is collected over a 50 epoch budget, with other
-hyperparameters kept the same as mentioned in section \ref{sec:bn}. }
+hyperparameters kept the same as mentioned in section \ref{sec:Expe}. }
 \label{fig:grid_searcha}
@@
 ImageNet dataset. Each datapoint is collected over a 50 epoch budget, with other
-hyperparameters kept the same as mentioned in section \ref{sec:bn}. }
+hyperparameters kept the same as mentioned in section \ref{sec:Expe}. }
 \label{fig:grid_searchb}

### 2022-11-17 0100.tex
Mod4
Mod5
Mod6

@@
 by the forward pass. As a result, the two stages are coupled. This is how most neural
-networks are trained currently. However, gradient modification in the backward pass has
-seldom been studied in the literature. In this paper we explore decoupled training,
-where we alter the gradients in the backward pass. We propose a simple yet powerful
-method called PowerGrad Transform (PGT), that alters the gradients before the weight
-update in the backward pass and significantly enhances the predictive performance of the
-neural network. PGT trains the network to arrive at a better optima at convergence. It
-is computationally extremely efficient, virtually adding no additional cost to either
-memory or compute, but results in improved final accuracies on both the training and
-test sets. PGT is easy to integrate into existing training routines, requiring just a
-few lines of code. PGT accelerates training and makes it possible for the network to
-better fit the training data. With decoupled training, PGT improves baseline accuracies
+networks are trained hitherto. Gradient modification in the backward pass has seldom
+been studied in the literature. In this paper we explore decoupled training, where we
+alter the gradients in the backward pass. We propose a simple yet powerful method called
+PowerGrad Transform (PGT), that alters the gradients before the weight update in the
+backward pass and significantly enhances the predictive performance of a convolutional
+neural network. PGT trains networks to arrive at a better optima at convergence. It is
+computationally efficient, and adds no additional cost to either memory or compute, but
+results in improved final accuracies on both the training and test datasets. PowerGrad
+Transform is easy to integrate into existing training routines, requiring just a few
+lines of code. PGT accelerates training and makes it possible for the network to better
+fit the training data. With decoupled training, our method improves baseline accuracies
 for ResNet-50 by 0.73\%, for SE-ResNet-50 by 0.66\% and by more than 1.0\% for the
 non-normalized ResNet-18 network on the ImageNet classification task.
-\keywords{backpropagation \and gradients \and neural networks \and softmax \and clipping.}
+\keywords{backpropagation \and gradients \and softmax \and clipping \and classification.}
 \end{abstract}
@@
 
-where first the gradient $\nabla_{W_D}(L)$ is transformed using the transformation
+where the gradient $\nabla_{W_D}(L)$ is first transformed using the transformation
 function `$f$' and then this transformed gradient is propagated through the rest of the
@@
 \item We perform theoretical analysis of the properties of the PowerGrad transformation
-(section \ref{sec:pgt_prop}) and explore its effect on weight parameters and gradients,
-logits, loss values and class separation (section \ref{sec:pgt_effects}).
+and explore its effect on weight parameters and gradients, logits, loss values and class
+separation (section \ref{sec:pgt_prop}).
 
@@
 
-Gradient Clipping \cite{pascanu2013difficulty} is a technique that involves changing or
-clipping the gradients with respect to a predefined threshold value during backward
-propagation through the network and updating the weights using the clipped gradients
-\cite{zhang2019gradient}\cite{smith2020generalization}. By rescaling the gradients, the
-weight updates are likewise rescaled, significantly reducing the risk of an overflow or
-underflow \cite{pascanu2012understanding}. GC can be used for training networks without
-batch-normalization. At larger batch sizes, the clipping threshold in GC becomes highly
-sensitive and requires extensive finetuning for various models, batch sizes, and
-learning rates. As we demonstrate later in our studies, GC is not as effective in
-improving the performance of non-normalized networks. Adaptive Gradient Clipping
-\cite{brock2021high} is developed to further enhance backward pass gradients than what
-is performed by GC. It takes into account the fact that the ratio of the gradient norm
-to the weight norm can provide an indication of the expected change in a single step of
-optimization. Label smoothing, introduced by Szegedy et al.
+Gradient Clipping \cite{pascanu2013difficulty} is a gradient modification method that
+involves clipping/altering the gradients with respect to a predefined threshold value
+during backward propagation through the network and updating the weights using the
+clipped gradients \cite{zhang2019gradient}\cite{smith2020generalization}. By rescaling
+the gradients, the weight updates are likewise rescaled, significantly reducing the risk
+of an overflow or underflow \cite{pascanu2012understanding}. GC can be used for training
+networks without batch-normalization. At larger batch sizes, the clipping threshold in
+GC becomes highly sensitive and requires extensive finetuning for various models, batch
+sizes, and learning rates. As we demonstrate later in our studies, GC is not as
+effective in improving the performance of non-normalized networks. Adaptive Gradient
+Clipping \cite{brock2021high} is developed to further enhance backward pass gradients
+than what is performed by GC. It takes into account the fact that the ratio of the
+gradient norm to the weight norm can provide an indication of the expected change in a
+single step of optimization. Label smoothing, introduced by Szegedy et al.
 \cite{szegedy2016rethinking}, utilizes smoothing of the ground truth labels as a method
@@
 performs better than GC in non-normalized networks. However, we show that PGT
-outperforms both in such networks.
-
-Knowledge distillation \cite{hinton2015distilling} is a process in which two networks
-are trained with hard and soft labels alternatively. Variants of knowledge distillation
-include self-distillation \cite{zhang2018deep}, identical student network distillation
-\cite{furlanello2018born}, channel distillation \cite{ge2019distilling}, regularizing
-wrong predictions \cite{yun2020regularizing} and representation or embedding based
-knowledge distillation \cite{yao2018deep}\cite{aguilar2020knowledge}. The key difference between PGT and knowledge
-distillation is that in the latter, the transformation is applied in the forward pass,
-while PGT is a backward pass modification only. Also, in distillation settings the
-temperature parameter is a part of the network's computation graph. In the case of
-PowerGrad Transform, we directly tamper the gradients without introducing any change in
-the forward pass. PGT differs from self-knowledge distillation as it neither introduces
-any additional sub-modules nor creates different ensembles to improve the performance of
-the model. PGT follows the standard neural network training mechanism with modified
-gradients.
+outperforms both in networks such as ResNets.
+
+Knowledge distillation (KD) \cite{hinton2015distilling} is a process in which two
+networks are trained with hard and soft labels alternatively. Variants of knowledge
+distillation include self-distillation \cite{zhang2018deep}, identical student network
+distillation \cite{furlanello2018born}, channel distillation \cite{ge2019distilling},
+regularizing wrong predictions \cite{yun2020regularizing} and representation or
+embedding based knowledge distillation \cite{yao2018deep}. Even though both PGT and KD
+requires probability manipulation, the key difference is that in the latter the
+transformation is applied in the forward pass, while PGT is a backward pass modification
+only. In distillation settings the temperature parameter is a part of the network's
+computation graph. In the case of PowerGrad Transform, we directly tamper the gradients
+without introducing any change in the forward pass. PGT differs from self-knowledge
+distillation as it neither introduces any additional sub-modules nor creates different
+ensembles to improve the performance of the model. PGT follows the standard neural
+network training mechanism with modified gradients.
 
@@
 
-We provide the proofs in the supplementary section and conclude that the stationary
-threshold is a monotonic non-decreasing function with respect to $\alpha$. Also the
-derivative of PGT function with respect to the true probabilities is non-negative which
-in turn means that the transformation is an order-preserving map. All values greater
-than the threshold move towards the threshold after transformation and all values below
-the threshold also move towards the threshold, and the threshold itself moves
-monotonically towards $1/C$ as $\alpha$ is decreased from $1$ to $0$. This concludes
-that the transformation smooths the original distribution.
-
-\subsection{Effects of PGT on logit, weight gradients and class separation}
-\label{sec:pgt_effects}
-
-PGT, other than having the properties mentioned above, also:
-\begin{enumerate}
-\item restricts the magnitude of the directional derivative for each logit from becoming too small or too high
-\item enhances gradient behavior of weights in a manner that improves learning
-\item leads to better class separation.
-\end{enumerate}
-
-We demonstrate each of these effects in detail in supplementary sections
-\ref{sec:pgt_logit_derivative}, \ref{sec:pgt_weight_gradients} and
-\ref{sec:pgt_class_separation} respectively.
-
-\section{Experiments}
-\label{sec:Expe}
+We provide the proofs in supplementary section \ref{sec:proofs} and conclude that the
+stationary threshold is a monotonic non-decreasing function with respect to $\alpha$.
+Also the derivative of PGT function with respect to the true probabilities is
+non-negative which in turn means that the transformation is an order-preserving map. All
+values greater than the threshold move towards the threshold after transformation and
+all values below the threshold also move towards the threshold, and the threshold itself
+moves monotonically towards $1/C$ as $\alpha$ is decreased from $1$ to $0$. This
+concludes that the transformation smooths the original distribution.
 
@@
 \end{figure}
+
+PGT, other than having the properties mentioned above, also satisfies the following. The
+transformation:
+\begin{enumerate}
+\item restricts the magnitude of the directional derivative for each logit from becoming too small or too high
+\item enhances gradient behavior of weights in a manner that improves learning
+\item leads to better class separation.
+\end{enumerate}
+
+We demonstrate each of these effects in detail in supplementary sections
+\ref{sec:pgt_logit_derivative}, \ref{sec:pgt_weight_gradients} and
+\ref{sec:pgt_class_separation} respectively.
+
+\section{Experiments}
+\label{sec:Expe}
 
@@
 during convergence. Per epoch training and test accuracy plots of ResNet-18 and
-ResNet-50 (both with and without PGT) are shown in Fig. \ref{fig:metrics}(a, b).
+ResNet-50 (both with and without PGT) are shown in Fig. \ref{fig:metrics}(a,b).
 Practioners can also choose to accelerate training and save as much as $40\%$ of the
@@
 zeros and such filters do not contribute at all to determine the input-output
-relationship of a dataset, as the feature tensor it produces is also an all-zero tensor.
-According to our findings, when a filter once zeroed out does not recover with further
-training.
+relationship of a dataset, as the feature tensor it produces is also filled with zeros
+for the corresponding filter. When a filter once zeroes out, it does not recover with
+further training, as all gradients that it receives in future iterations are all zeros.
 
@@
 
-\captionsetup{font=footnotesize}
 \begin{subfigure}[t]{0.16\textwidth}
@@
 different filter or feature vector of a particular layer of a non-BN variant of
-ResNet-18. }
+ResNet-18. Readers are requested to zoom in or refer to section \ref{sec:big_plots} for
+a zoomed-in view. }
 \label{fig:norm_plots}
@@
 
-Fig. \ref{fig:norm_plots}(c) and \ref{fig:norm_plots}(e) are the plot of the final conv
-layer's filters (layer $19$) and the features output of global average pooling (GAP)
-layer respectively. We observe a number of filters in the final layer completely zeroing
-out. The feature vector after GAP layer directly interfaces with the fully connected
-layer immediately after. Therefore any zeroed out features leads to permanent
+Fig. \ref{fig:norm_plots}(c) and \ref{fig:norm_plots}(e) are plots of the final conv
+layer's filters (layer $19$) and the features output of final global average pooling
+(GAP) layer respectively. We observe a number of filters and features in the final layer
+completely zeroing out. The feature vector after the GAP layer directly interfaces with
+the fully connected layer. Therefore any zeroed out features leads to permanent
 information loss, as it does not contribute to the learning of decision boundaries in
@@
 for the affected layers. With large batch sizes, it is possible that an entire layer
-zeroes out [Fig. \ref{fig:high_bs}]. Gradient modification methods such as PGT or AGC
-can alleviate the zeroing out phenomena. We observe that the number of zeroed out
+zeroes out as shown in the figure below. Gradient modification methods such as PGT or
+AGC can alleviate the zeroing out phenomena as we observe that the number of zeroed out
 filters has considerably reduced [Fig. \ref{fig:norm_plots}(b, d)]. The final feature
@@
 feature extracting layers to the fully connected layer. When PGT and AGC are combined,
-we get the most benefit for layer $15$ [Fig \ref{fig:norm_plots}(g-i)] and layer $2$
-[Fig \ref{fig:norm_plots}(j-l)].
-
-In Table \ref{tab:wobn_table} we find the baseline performance for high batch sizes
-($1024$) is drastically inferior to baselines for other batch sizes. PGT helps regain
-some of the lost performance by $0.682\%$ ($65.498\%$ vs. $64.816\%$). At a batch size
-of 512, invoking PGT improves the training accuracy baseline by $1.48\%$ and the test
-accuracy baseline by $0.684\%$, while at a batch size of 256, the improvement in
-training and test accuracies are $1.11\%$ and $1.018\%$ respectively. In comparison, the
-test accuracy improvements obtained by GC and AGC at batch size of 256
-is much less at $0.27\%$ and $0.5\%$, respectively. On the training accuracy front,
-since we get a significant boost ($1.48\%$ at batch size of 512 and $1.11\%$ at batch
-size of 256), it leads us to infer that when PowerGrad Transform is used, the network
-
-\begin{wrapfigure}{l}{0.4\textwidth}
-\vspace{-1cm}
-\centering
-\includegraphics[width=0.45\textwidth]{baseline_high_bs-w-layer2-2}
-\caption{}
-\label{fig:high_bs}
-\vspace{-2cm}
-\end{wrapfigure}
-
-\noindent
-fits the training dataset more tightly and the convergence optima is significantly
-superior. Combining GC with PGT is not very significant. When AGC and PGT are combined,
-we see a tremendous increase in test accuracy of over $2.06\%$ over the baseline.
-Additionally, we provide comprehensive graphs of the norm profiles of the AGC + PGT
-method for all layers in the supplementary section \ref{sec:plots1}\ -\
-\ref{sec:plots4}.
+we get the most benefit as we demonstrate for layers $15$ and layer $2$ (Fig
+\ref{fig:norm_plots}(g-i)] and Fig \ref{fig:norm_plots}(j-l) respectively).
 
@@
 
+In Table \ref{tab:wobn_table} we find that the baseline performance for high batch sizes
+($1024$) is drastically inferior to baselines for other batch sizes. PGT helps regain
+some of the lost performance by $0.682\%$ ($65.498\%$ vs. $64.816\%$). At a batch size
+of 512, invoking PGT improves the training accuracy baseline by $1.48\%$ and the test
+accuracy baseline by $0.684\%$, while at a batch size of 256, the improvement in
+training and test accuracies are $1.11\%$ and $1.018\%$ respectively. In comparison, the
+test accuracy improvements obtained by GC and AGC at batch size of 256
+
+\begin{wrapfigure}{l}{0.4\textwidth}
+\vspace{-1cm}
+\centering
+\includegraphics[width=0.45\textwidth]{baseline_high_bs-w-layer2-2}
+\vspace{-1.5cm}
+\end{wrapfigure}
+
+\noindent
+is much less at $0.27\%$ and $0.5\%$, respectively. On the training accuracy front,
+since we get a significant boost ($1.48\%$ at batch size of 512 and $1.11\%$ at batch
+size of 256), it leads us to infer that when PowerGrad Transform is used, the network
+fits the training dataset more tightly and the convergence optima is significantly
+superior. Combining GC with PGT is not very significant. When AGC and PGT are combined,
+we see a tremendous increase in test accuracy of over $2.06\%$ over the baseline.
+Additionally, we provide comprehensive graphs of the norm profiles of the AGC + PGT
+method for all layers in the supplementary section \ref{sec:plots1}\ -\
+\ref{sec:plots4}.
+
 \subsection{Effect on Loss, Logits and other metrics}
@@
 
-\begin{figure}[!bp]
+\begin{figure}[!t]
 \centering
@@
 \label{fig:stats}
+\vspace{-0.5cm}
 \end{figure}
@@
 \subsection{Proofs}
+\label{sec:proofs}
 
@@
 indices are shown. Each filter in each subplot is shown with a separate and randomly
-determined colour. } \end{figure}
+determined colour. Readers are requested to zoom in. } \end{figure}
 
@@
 of the unnormalized ResNet-18. Above each subplot, the layer indices are shown. Each
-filter in each subplot is shown with a separate and randomly determined colour. }
+filter in each subplot is shown with a separate and randomly determined colour. Readers
+are requested to zoom in. }
 \end{figure}
@@
 indices are shown. Each filter in each subplot is shown with a separate and randomly
-determined colour. } \end{figure}
+determined colour.
+Readers are requested to zoom in. } \end{figure}
 
@@
 
-For the layer indices used in figures Fig. \ref{fig:norm_plots}, \ref{fig:high_bs},
-\ref{fig:norm_plots}, we provide a block diagram depicting the different convolutional
-layers of ResNet-18 with corresponding layer indices in Fig. \ref{fig:resnet18}. Each
-layer represents a convolutional layer, with the layer index and the number of filters
-denoted alongside it. Downsampling convolutional layers in skip connections are drawn
-and their layer indices are shown on the skip connection paths.
+For the layer indices used in figures Fig. \ref{fig:norm_plots}, \ref{fig:norm_plots},
+we provide a block diagram depicting the different convolutional layers of ResNet-18
+with corresponding layer indices in Fig. \ref{fig:resnet18}. Each layer represents a
+convolutional layer, with the layer index and the number of filters denoted alongside
+it. Downsampling convolutional layers in skip connections are drawn and their layer
+indices are shown on the skip connection paths.
 
@@
 
+\clearpage
+
+\subsection{Plot of Figure \ref{fig:norm_plots}}
+\label{sec:big_plots}
+
+\begin{figure}[h]
+\centering
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-4-2}
+\caption{Layer 11\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-4-2}
+\caption{Layer 11\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-7-2}
+\caption{Layer 19\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-7-2}
+\caption{Layer 19\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-f-layer-22-1}
+\caption{GAP features\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-f-layer-22-1}
+\caption{GAP features\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
+\caption{ Norm vs iter. plots demonstrating layer characteristics (the zero out
+phenomena). Each colour represents a different filter or feature vector of a particular
+layer of a non-BN variant of ResNet-18. }
+\end{figure}
+
+\clearpage
+
+\begin{figure}[h]
+\centering
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-5-3}
+\caption{Layer 15\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-5-3}
+\caption{Layer 15\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/baseline-w-layer-1-2}
+\caption{Layer 2\\ \forceindentb\textbf{baseline}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/pgt-w-layer-1-2}
+\caption{Layer 2\\ \forceindentb\textbf{PGT}}
+\end{subfigure}
+
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-5-3}
+\caption{Layer 15\\ \forceindenta\textbf{AGC+PGT}}
+\end{subfigure}
+\begin{subfigure}[t]{0.34\textwidth}
+\includegraphics[width=\textwidth]{trimmed/agc_pgt-w-layer-1-2}
+\caption{Layer 2\\ \forceindenta\textbf{AGC+PGT}}
+\end{subfigure}
+\caption{ Norm vs iter. plots demonstrating the efficacy of PGT and AGC+PGT over
+baseline. Each colour represents a different filter or feature vector of a particular
+layer of a non-BN variant of ResNet-18. }
+\end{figure}
+
 \end{document}

### 2022-11-21 0100.tex
Mod7

@@
 \raggedbottom
+\makeatletter
+\def\blfootnote{\xdef\@thefnmark{}\@footnotetext}
+\makeatother
 \newcommand{\forceindenta}{\parindent=1em\indent\parindent=0pt\relax}
@@
 \begin{document}
-\title{Altering Backward Pass Gradients to improve Convergence}
-\author{Bishshoy Das\inst{1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{1}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Brejesh Lall\inst{3}\orcidID{xxxx-xxxx-xxxx-xxxx}\and Shiv Dutt Joshi\inst{3}\orcidID{0000-0002-0326-0524}\and Sumantra Dutta Roy\inst{3}\orcidID{0000-0002-2141-5067}}
+\title{Altering Backward Pass Gradients\\ to Improve Convergence}
+\author{Bishshoy Das\inst{\ast 1}\orcidID{0000-0002-6525-4015}\and Milton Mondal\inst{\ast 1}\orcidID{0000-0002-1832-0702}\and Brejesh Lall\inst{1}\orcidID{0000-0003-2677-3071}\and Shiv Dutt Joshi\inst{1}\orcidID{0000-0002-0326-0524}\and Sumantra Dutta Roy\inst{1}\orcidID{0000-0002-2141-5067}}
 \authorrunning{Das et al.}
@@
 \label{sec:Intr}
+
+\blfootnote{$^\ast \text{Equal contribution.}$}
 
@@
 \label{sec:plots1}
-\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{baseline-w_norm}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.57\textwidth]{compressed/baseline-w_norm}
 \caption{ Method: baseline, batch size = 256. This is a plot depicting the
@@
 indices are shown. Each filter in each subplot is shown with a separate and randomly
-determined colour. Readers are requested to zoom in. } \end{figure}
+determined colour. Readers are requested to zoom in. } 
+\end{figure}
 
@@
 \label{sec:plots2}
-\begin{figure}[ht] \centering
-\includegraphics[width=0.57\textwidth]{baseline_high_bs-w_norm} \caption{ Method:
-baseline (high batch size), batch size = 1024. This is a plot depicting the
-iteration-wise evolution of the norm of each filter in each layer. Each subplot
-corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the layer
-indices are shown. Each filter in each subplot is shown with a separate and randomly
-determined colour. Readers are requested to zoom in. } \end{figure}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.57\textwidth]{compressed/baseline_high_bs-w_norm}
+\caption{ Method: baseline (high batch size), batch size = 1024. This is a plot
+depicting the iteration-wise evolution of the norm of each filter in each layer. Each
+subplot corresponds to a layer of the unnormalized ResNet-18. Above each subplot, the
+layer indices are shown. Each filter in each subplot is shown with a separate and
+randomly determined colour. Readers are requested to zoom in. } \end{figure}
 
@@
 \label{sec:plots3}
-\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{pgt-w_norm}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.57\textwidth]{compressed/pgt-w_norm}
 \caption{ Method: PGT, batch size = 256. This is a plot depicting the iteration-wise
@@
 \label{sec:plots4}
-\begin{figure}[ht] \centering \includegraphics[width=0.57\textwidth]{agc_pgt-w_norm}
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.57\textwidth]{compressed/agc_pgt-w_norm}
 \caption{ Method: AGC + PGT, batch size = 256. This is a plot depicting the
@@
 determined colour.
-Readers are requested to zoom in. } \end{figure}
+Readers are requested to zoom in. }
+\end{figure}
 
